### 2018-04-23 Presentations/Milestone1/presentation.tex
Add Template for first presentation

 \documentclass{beamer}

\usetheme{MagdeburgFIN}
\usefonttheme{structurebold}
\usepackage{graphicx}
\usepackage{float}
\usepackage{url}
\usepackage{pdfpages}
\usepackage[ngerman]{babel}
\usepackage[utf8]{inputenc}

\title{Scientific Team Project}
\author{Johann Wagner, Johannes Wünsche, Marten Wallewein-Eising}
\date{\today}
\institute{Otto von Guericke University, Magdeburg}

\begin{document}

\begin{frame}[plain]
 \titlepage
\end{frame}

\section[Agenda]{}
\begin{frame}
\frametitle{Agenda}
\tableofcontents
\end{frame}

\begin{frame}
 \frametitle{Thank you for your attention!}
\end{frame}
\end{document}

### 2018-04-23 Presentations/Milestone1/presentation.tex
Add literature research

@@
 
+\section{Literature Research}
+\begin{frame}
+\frametitle{Literature Research}
+\begin{itemize}
+	\item Xu, Dongping, and Brett Bode. \emph{"Performance study and dynamic optimization design for thread pool systems".} Diss. United States. Department of Energy. Office of Science, 2004.
+	\item Ling, Yibei, Tracy Mullen, and Xiaola Lin. \emph{Analysis of optimal thread pool size.} ACM SIGOPS Operating Systems Review 34.2 (2000): 42-55.
+	\item Kang, DongHyun, et al.
+\emph{Prediction-based dynamic thread pool scheme for efficient resource usage.} Computer and Information Technology Workshops, 2008.
+CIT Workshops 2008.
+IEEE 8th International Conference on.
+IEEE, 2008.
+\end{itemize}
+\end{frame}
+
 \begin{frame}

### 2018-04-23 Presentations/Milestone1/presentation.tex
Added Intro

--- 2018-04-23 Presentations/Milestone1/presentation.tex
@@
 \title{Scientific Team Project}
-\author{Johann Wagner, Johannes Wünsche, Marten Wallewein-Eising}
+\subtitle{Thread Pool in GeckoDB/BOLSTER}
+\author{Johann Wagner, Johannes Wünsche, Marten Wallewein-Eising, Robert Jenderise, Benedikt Zeilinger}
 \date{\today}
 \institute{Otto von Guericke University, Magdeburg}
+
+%Topic, schedule, and team presentation & first results of literature research
 
@@
 
+\section{Introduction}
+	\begin{frame}
+		\frametitle{Introduction}
+		\begin{itemize}
+			\item \emph{In computer programming, a thread pool is a software design pattern for achieving concurrency of execution in a computer program.}
+			\item Pool of Threads
+			\item Reduces overhead of creating new threads by reusing existing threads.
+		\end{itemize}
+	\end{frame}
+	\begin{frame}
+		\frametitle{Current Situation}
+		\begin{itemize}
+			\item Operations are spawning dozens of threads, which are not handled and managed..
+			\item After some operations profilers dies.
+		\end{itemize}
+	\end{frame}
+
+	\begin{frame}
+		\frametitle{Desirable Situation}
+		\begin{itemize}
+			\item Reuse of Threads with thread pools.
+			\item Limited amount of threads, which are handled correctly.
+		\end{itemize}
+	\end{frame}
+
 \begin{frame}

### 2018-04-23 Presentations/Milestone1/presentation.tex

@@
 
+\section{Literature Research}
+    \begin{frame}
+        \frametitle{Literature Research}
+        \begin{itemize}
+            \item Xu, Dongping, and Brett Bode. \emph{"Performance study and dynamic optimization design for thread pool systems".} Diss. United States. Department of Energy. Office of Science, 2004.
+            \item Ling, Yibei, Tracy Mullen, and Xiaola Lin. \emph{Analysis of optimal thread pool size.} ACM SIGOPS Operating Systems Review 34.2 (2000): 42-55.
+            \item Kang, DongHyun, et al.
+\emph{Prediction-based dynamic thread pool scheme for efficient resource usage.} Computer and Information Technology Workshops, 2008.
+CIT Workshops 2008.
+IEEE 8th International Conference on.
+IEEE, 2008.
+        \end{itemize}
+    \end{frame}
+
     \begin{frame}

### 2018-04-23 Presentations/Milestone1/presentation.tex
Add schedule to presentation
Update presentation.tex
Add big picture
Reworked parts of presentation1

@@
 
+\section{Motivation}
+	\begin{frame}
+		\frametitle{Motivation}
+		\begin{center}
+			\includegraphics[width=0.6\textwidth]{img/bigpicture.png}
+		\end{center}
+	\end{frame}
+
 \section{Introduction}
@@
 			\item \emph{In computer programming, a thread pool is a software design pattern for achieving concurrency of execution in a computer program.}
-			\item Pool of Threads
+			\item Pool of Threads to distribute parallel tasks
 			\item Reduces overhead of creating new threads by reusing existing threads.
@@
 	\end{frame}
+
 	\begin{frame}
@@
 			\item Operations are spawning dozens of threads, which are not handled and managed..
-			\item After some operations profilers dies.
+			\item Heavy overhead of creating and removing threads
+			\item After some operations profiler dies (too many threads).
+			\item Debugging/Profiling not possible due to number of threads 
 		\end{itemize}
@@
 			\item Reuse of Threads with thread pools.
-			\item Limited amount of threads, which are handled correctly.
+			\item Reduce overhead of thread management
+			\item Improve maintainability of parallel code
+			\item Defy competition to existing thread pool libs
 		\end{itemize}
+	\end{frame}
+
+\section{Planning}
+
+\begin{frame}
+	\frametitle{Team Roles}
+	\begin{figure}
+		\includegraphics[width=0.7\linewidth]{img/TeamRoles}
+	\end{figure}
+\end{frame}
+
+\begin{frame}
+	\frametitle{Schedule}
+	\includegraphics[width=1.0\textwidth]{img/schedule.png}
 \end{frame}

### 2018-05-07 include/thread_pool.h
Add short waiting interval instead of constant polling

@@ -11,6 +11,7 @@
 #include "status.h"
 #include "priority_queue.h"
 #include "thread_status.h"
+#include <time.h>
 
 //
 //	GLOBAL VARIABLES

### 2018-05-07 src/thread_pool.c

@@ -92,6 +92,9 @@ status_e gecko_pool_wait_for_id(size_t id, thread_pool* pool) {
 void *__thread_main(void* args) {
   __thread_information* thread_info = (__thread_information*)args;
 
+  //structs for short waiting interval
+  struct timespec waiting_time_start = {0 , 0};
+  struct timespec waiting_time_end = {0 , 100};
 
   while(1){
 
@@ -113,8 +116,8 @@ void *__thread_main(void* args) {
     if(thread_status_will_terminate == __update_thread_status(thread_info->pool, thread_info->id, thread_status_idle))
       break;
 
-    // TODO: idle
-
+    if(!next_task) {}
+        nanosleep(&waiting_time_start, &waiting_time_end);
 	}
 
   __update_thread_status(thread_info->pool, thread_info->id, thread_status_finished);

### 2018-05-09 Presentations/Milestone2/presentation.tex
Begin milestone 2 presentation

--- 2018-04-23 Presentations/Milestone1/presentation.tex
@@
 
-\title{Scientific Team Project}
-\subtitle{Thread Pool in GeckoDB/BOLSTER}
+\title{Thread Pool in GeckoDB/BOLSTER}
+\subtitle{Milestone II: Concept \& additional literature research}
 \author{Johann Wagner, Johannes Wünsche, Marten Wallewein-Eising, Robert Jenderise, Benedikt Zeilinger}
@@
 
-%Topic, schedule, and team presentation & first results of literature research
+% Milestone II: Concept & additional literature research
 
@@
 
-\section{Introduction}
+\section{Previously...}
 \begin{frame}
-		\frametitle{Introduction}
+	\frametitle{Previously in Thread Pool in GeckoDB/BOLSTER}
 	\begin{itemize}
-			\item \emph{In computer programming, a thread pool is a software design pattern for achieving concurrency of execution in a computer program.}
-			\item Pool of Threads
-			\item Reduces overhead of creating new threads by reusing existing threads.
+		\item Operations are spawning dozens of threads, which are not handled and managed
+		\item Heavy overhead of creating and removing threads
+		\item After some operations profiler dies (too many threads)
 	\end{itemize}
+	\begin{center}
+		\includegraphics[width=0.3\textwidth]{img/thumbs_down.png}
+	\end{center}
 \end{frame}
+
 \begin{frame}
-		\frametitle{Current Situation}
+	\frametitle{Previously in Thread Pool in GeckoDB/BOLSTER}
 	\begin{itemize}
-			\item Operations are spawning dozens of threads, which are not handled and managed..
-			\item After some operations profilers dies.
+		\item Reuse of Threads with thread pools.
+		\item Reduce overhead of thread management
+		\item Improve maintainability of parallel code
+	\end{itemize}
+	\begin{center}
+		\includegraphics[width=0.3\textwidth]{img/thumbs_up.png}
+	\end{center}
+\end{frame}
+
+\section{Concept and Design}
+\begin{frame}
+	\frametitle{Concept}
+	\begin{itemize}
+		\item Thread pool with runtime fixed amount of threads and lock free priority queue
+		\item Tasks are not preemtable and independent
+		\item Array of tasks passed to an enqueue function, which waits until all tasks are finished
+		\item Threads without tasks idle until they get a new one
+		\item Monitoring/managing state and assigned task of each thread
 	\end{itemize}
@@
 \begin{frame}
-		\frametitle{Desirable Situation}
+\frametitle{Design}
+	\begin{center}
+		\includegraphics[width=0.9\textwidth]{img/pool_queue.png}
+	\end{center}
+\end{frame}
+
+\section{Current Progress}
+
+\section{Next Steps}
+
+\section{Additional Literature Research}
+    \begin{frame}
+        \frametitle{Literature Research}
         \begin{itemize}
-			\item Reuse of Threads with thread pools.
-			\item Limited amount of threads, which are handled correctly.
+            \item 
         \end{itemize}

### 2018-05-09 Presentations/Milestone2/presentation.tex
Second milestone presentation (without literature)

@@
 \section{Current Progress}
+\begin{frame}
+	\frametitle{Current Progress}
+	We...
+	\begin{itemize}
+		\item Implemented the lock free priority queue 
+		\item Implemented the creation of threads and the assigning of tasks
+		\item Added gtest framework to project and wrote tests for pool and queue
+		\item Began first implementations of waiting for tasks to be finished
+	\end{itemize}
+\end{frame}
 
 \section{Next Steps}
+\begin{frame}
+	\frametitle{Next Steps}
+	We have to ...
+	\begin{itemize}
+		\item Find an efficient way ton wait for tasks
+		\item Find an efficient way to idle threads until they get the next tasks
+		\item Add metrics to threads and tasks and measure performance and overhead
+		\item Add tests for all implementations
+	\end{itemize}
+\end{frame}
+
+\begin{frame}
+	\frametitle{Next Steps - Scheduling}
+	Current approach:  
+	\begin{itemize}
+		\item Threads ask for next task in queue
+	\end{itemize}
+	Desired implementation: 
+	\begin{itemize}
+		\item Make own scheduling, assign tasks to threads
+		\item Add possibility to try different scheduling approaches
+		\item Measure thread idle and busy time for the approaches
+	\end{itemize}
+\end{frame}
 
@@
     \begin{frame}
-        \frametitle{Literature Research}
+        \frametitle{Additional Literature Research}
         \begin{itemize}

### 2018-05-14 src/main.c
fixed some race conditions fixed memory leaks in priority_queue and thread_pool added dynamic lazy resizing of thread_pool

@@ -1,9 +1,22 @@
 #include <stdio.h>
+#include <stdlib.h>
 #include "priority_queue.h"
+#include "thread_pool.h"
+
+void work(void* args)
+{
+	int* res = args;
+	int sum = 0;
+	for(int i = 0; i <= *res; ++i)
+		sum += i * i / 8;
+
+	*res = sum;
+//	printf("%d %d\n", n, sum);
+}
 
 int main()
 {
-	priority_queue_t queue;
+/*	priority_queue_t queue;
 	priority_queue_init(&queue);
 
 	for(size_t i = 0; i < 256; ++i)
@@ -15,7 +28,36 @@ int main()
 	{
 		size_t i = (size_t)priority_queue_pop(&queue);
 		printf("%I64d\n",i);
+	}*/
+
+	thread_pool* pool = thread_pool_create(2);
+
+	const int numThreads = 1 << 11;
+	thread_task tasks[numThreads];
+	int results[numThreads];
+	for(int i = numThreads-1; i >= 0; --i){
+		results[i] = i;
+		tasks[i].args = &results[i];
+		tasks[i].routine = work;
+		tasks[i].priority = 0;
+		tasks[i].group_id = i;
+		gecko_pool_enqueue_task(&tasks[i], pool);
+		if( i == numThreads * 1 / 3) thread_pool_resize(pool, 4);
+		if(i == numThreads * 2 / 3) thread_pool_resize(pool, 3);
+		if(i == numThreads * 3 / 4) thread_pool_resize(pool,2);
 	}
+//	gecko_pool_wait_for_id(1,pool);
+	_sleep(2000);
+	thread_pool_free(pool);
+
+	// verify results
+	int sum = 0;
+	for(int i = 0; i < numThreads; ++i)
+	{
+		sum += i*i / 8;
+		if(results[i] != sum) printf("error at %d: %d != %d\n", i, sum, results[i]);
+	}
+	printf("tests completed.");
 
 	getchar();
 }

### 2018-05-14 src/priority_queue.c

@@ -10,6 +10,11 @@ void priority_queue_init(priority_queue_t* queue)
     queue->data = NULL;
 }
 
+void priority_queue_free(priority_queue_t* queue)
+{
+    free(queue->data);
+}
+
 void priority_queue_resize(priority_queue_t* queue, size_t size)
 {
     __element_info_t* new_data = malloc(sizeof(__element_info_t) * size);
@@ -103,13 +108,18 @@ void* priority_queue_pop(priority_queue_t* queue)
 {
     pthread_mutex_lock(&queue->mutex);
 
-    --queue->num_elements;
+	void* ptr = NULL;
+	
+	if(queue->num_elements)
+	{
+		--queue->num_elements;
 
-    void* ptr = queue->data[0].element;
- //   queue->data[0] = queue->data[queue->num_elements];
-    queue->data[0] = queue->data[queue->num_elements];
-    down_heap(queue->data, queue->num_elements+1);
-
+		ptr = queue->data[0].element;
+	 //   queue->data[0] = queue->data[queue->num_elements];
+		queue->data[0] = queue->data[queue->num_elements];
+		down_heap(queue->data, queue->num_elements+1);
+	}
+	
     pthread_mutex_unlock(&queue->mutex);
     return ptr;
 }

### 2018-05-14 Presentations/Milestone2/presentation.tex
Add notes from Markus to presentation2

@@
 		\item After some operations profiler dies (too many threads)
+		\item Because of Late/Early Materialization and Prefetching Support, existing libs are not adequate
 	\end{itemize}
@@
 		\item Improve maintainability of parallel code
+		\item Support the performance and design requirements (e.g. Future and Promise) of BOLSTER
 	\end{itemize}
@@
 		\item Monitoring/managing state and assigned task of each thread
+		\item Allow usage of multiple, independent thread pool instances 
 	\end{itemize}
@@
 		\item Began first implementations of waiting for tasks to be finished
+		\item Measure performance of task assigning and execution
 	\end{itemize}
@@
         \begin{itemize}
-            \item 
+            \item Albahari, Joseph. \emph{Threading in C.} online. Joseph Albahari \& O’Reilly Media, Inc (2006).
+            \item Rasmussen, Rasmus V., and Michael A. Trick. \emph{Round robin scheduling–a survey.} European Journal of Operational Research 188.3 (2008): 617-636.
+            \item Shreedhar, Madhavapeddi, and George Varghese. \emph{Efficient fair queueing using deficit round robin.} ACM SIGCOMM Computer Communication Review. Vol. 25. No. 4. ACM, 1995.
         \end{itemize}

### 2018-05-14 Presentations/Milestone2/presentation.tex
small corrections
Add wait for task id slide

@@
 	\begin{itemize}
-		\item Thread pool with runtime fixed amount of threads and lock free priority queue
+		\item Thread pool with runtime dynamic amount of threads and priority queue
 		\item Tasks are not preemtable and independent
@@
 	\begin{itemize}
-		\item Implemented the lock free priority queue 
+		\item Implemented the priority queue 
 		\item Implemented the creation of threads and the assigning of tasks
@@
 		\item Measure performance of task assigning and execution
+	\end{itemize}
+\end{frame}
+
+\begin{frame}
+	\frametitle{Wait for task ID - Problem}
+	\begin{itemize}
+		\item Problem: waiting for a task to complete could lead to a
+			race condition
+		\item This was caused by the uncertainty of a thread\_task
+			during transition between the priority queue and the
+			thread pool
+		\item Solution: Threads now increase an atomic counter
 	\end{itemize}
@@
 	\begin{itemize}
-		\item Find an efficient way ton wait for tasks
+		\item Find an efficient way to wait for tasks
 		\item Find an efficient way to idle threads until they get the next tasks
@@
         \begin{itemize}
-            \item Albahari, Joseph. \emph{Threading in C.} online. Joseph Albahari \& O’Reilly Media, Inc (2006).
+            \item Albahari, Joseph. \emph{Threading in C\#.} online. Joseph Albahari \& O’Reilly Media, Inc (2006).
             \item Rasmussen, Rasmus V., and Michael A. Trick. \emph{Round robin scheduling–a survey.} European Journal of Operational Research 188.3 (2008): 617-636.

### 2018-05-14 include/thread_pool.h
implemented waiting for task(groups)

@@ -34,9 +34,22 @@ typedef struct thread_task {
 
 struct __thread_information;
 
+
+typedef struct __task_state{
+	atomic_int task_count; //< remaining tasks in this group
+	unsigned generation;
+} __task_state;
+
+typedef struct task_handle{
+	size_t index;
+	unsigned generation;
+} task_handle;
+
 typedef struct thread_pool {
 	pthread_t* pool;
 	priority_queue_t* waiting_tasks;
+	__task_state* task_group_states;
+	size_t task_state_capacity;
 	size_t size;
 	size_t capacity;
 //    thread_status_e* thread_status;
@@ -65,10 +78,10 @@ thread_pool* thread_pool_create(size_t num_threads);
 void thread_pool_free(thread_pool* pool);
 
 status_e thread_pool_resize(thread_pool* pool, size_t newSize);
-status_e gecko_pool_enqueue_tasks(thread_task* task, thread_pool* pool, size_t num_threads);
-status_e gecko_pool_enqueue_task(thread_task* task, thread_pool* pool);
+status_e gecko_pool_enqueue_tasks(thread_task* task, thread_pool* pool, size_t num_threads, task_handle* hndl);
+status_e gecko_pool_enqueue_task(thread_task* task, thread_pool* pool, task_handle* hndl);
 size_t gecko_pool_create_group_id();
-status_e gecko_pool_wait_for_id(size_t id, thread_pool* pool);
+status_e thread_pool_wait_for_task(thread_pool* pool, task_handle* hndl);
 
 //
 //	INTERNAL METHODS

### 2018-05-14 src/main.c

@@ -31,7 +31,7 @@ int main()
 	}*/
 
 	thread_pool* pool = thread_pool_create(2);
-
+	task_handle hndl;
 	const int numThreads = 1 << 11;
 	thread_task tasks[numThreads];
 	int results[numThreads];
@@ -40,14 +40,13 @@ int main()
 		tasks[i].args = &results[i];
 		tasks[i].routine = work;
 		tasks[i].priority = 0;
-		tasks[i].group_id = i;
-		gecko_pool_enqueue_task(&tasks[i], pool);
+		if(i == 0) tasks[i].priority = 1;
+		gecko_pool_enqueue_task(&tasks[i], pool, &hndl);
 		if( i == numThreads * 1 / 3) thread_pool_resize(pool, 4);
 		if(i == numThreads * 2 / 3) thread_pool_resize(pool, 3);
 		if(i == numThreads * 3 / 4) thread_pool_resize(pool,2);
 	}
-//	gecko_pool_wait_for_id(1,pool);
-	_sleep(2000);
+	thread_pool_wait_for_task(pool, &hndl);
 	thread_pool_free(pool);
 
 	// verify results

### 2018-05-14 src/thread_pool.c

@@ -18,6 +18,9 @@ thread_pool* thread_pool_create(size_t num_threads) {
    pool->thread_tasks = calloc(num_threads, sizeof(thread_task*));
    pool->thread_infos = malloc(sizeof(__thread_information*) * pool->capacity);
 
+  pool->task_state_capacity = 4096;
+  pool->task_group_states = calloc(pool->task_state_capacity, sizeof(__task_state));
+
    pthread_t* threads = malloc(sizeof(pthread_t) * pool->capacity);
    pool->pool = threads;
 
@@ -53,6 +56,7 @@ void thread_pool_free(thread_pool* pool) {
     free(pool->waiting_tasks);
     free(pool->thread_tasks);
     free(pool->thread_infos);
+    free(pool->task_group_states);
     free(pool);
 }
 
@@ -81,37 +85,42 @@ status_e thread_pool_resize(thread_pool* pool, size_t num_threads)
 	return status_ok;
 }
 
-status_e gecko_pool_enqueue_tasks(thread_task* tasks, thread_pool* pool, size_t num_tasks) {
-  size_t group_id = gecko_pool_create_group_id();
+status_e gecko_pool_enqueue_tasks(thread_task* tasks, thread_pool* pool, size_t num_tasks, task_handle* hndl) {
+  
+  // find unused slot
+  size_t ind = 0;
+  for(; ind < pool->task_state_capacity && pool->task_group_states[ind].task_count; ++ind);
+  if(ind == pool->task_state_capacity) return status_failed;
+
+  ++pool->task_group_states[ind].generation;
+  pool->task_group_states[ind].task_count = num_tasks;
   
   for(size_t i= 0; i < num_tasks; i++) {
     //Assuming queue will conatin something like this, this will not fail
-    tasks[i].group_id = group_id;
+    tasks[i].group_id = ind;
     priority_queue_push(pool->waiting_tasks, &tasks[i], tasks[i].priority); 
   }
 
-//  gecko_pool_wait_for_id(group_id, pool);
+  if(hndl){
+    hndl->index = ind;
+    hndl->generation = pool->task_group_states[ind].generation;
+  }
 
   return status_ok;
 }
 
-status_e gecko_pool_enqueue_task(thread_task* task, thread_pool* pool) {
-  return gecko_pool_enqueue_tasks(task, pool, 1);
[…]

### 2018-05-14 testing/tests/threadpool/test_pool.hpp
Add Compiler information to README Fix test_pool.hpp so that the correct amount of parameters will be given

@@ -1,6 +1,6 @@
 #include <gtest/gtest.h>
 
-extern "C" 
+extern "C"
 {
 	#include "../../../include/thread_pool.h"
 }
@@ -12,7 +12,7 @@ TEST(ThreadPool, Create) {
     EXPECT_TRUE(pool);
 
     thread_pool_free(pool);
-   
+
 }
 
 // Test the creatrion of group Ids of thread pools
@@ -39,15 +39,15 @@ TEST(ThreadPool, BasicTasks){
     size_t id2 = gecko_pool_create_group_id();
 
     thread_task* test_task = (thread_task*)malloc(sizeof(thread_task));
-    
+
     int test = 5;
-    
+
     test_task->args = (void*)&test;
     test_task->routine = &basicTask;
 
     for(int i = 0; i < 6; i++){
         std::cout << "enqueueing " << i << std::endl;
-        gecko_pool_enqueue_task(test_task, pool);
+        gecko_pool_enqueue_task(test_task, pool, NULL);
         std::cout << "enqueued " << i << std::endl;
     }
 

### 2018-05-16 include/thread_pool.h
added some documentation

@@ -49,7 +49,7 @@ typedef struct thread_pool {
 	pthread_t* pool;
 	priority_queue_t* waiting_tasks;
 	__task_state* task_group_states;
-	size_t task_state_capacity;
+	size_t task_state_capacity; // number of threads that can be tracked
 	size_t size;
 	size_t capacity;
 //    thread_status_e* thread_status;
@@ -57,13 +57,6 @@ typedef struct thread_pool {
 	thread_task** thread_tasks;
 } thread_pool;
 
-/*typedef struct __enqueued_task {
-	struct thread_task* thread;
-	size_t priority;
-	size_t id;
-	size_t is_activated;
-} __enqueued_task;
-*/
 typedef struct __thread_information {
 	thread_pool* pool;
 	size_t id;
@@ -75,12 +68,20 @@ typedef struct __thread_information {
 //
 
 thread_pool* thread_pool_create(size_t num_threads);
+// Releases all resources hold by the threadpool. 
+// Currently working threads may finish but tasks left in the queue will be discarded.
 void thread_pool_free(thread_pool* pool);
 
-status_e thread_pool_resize(thread_pool* pool, size_t newSize);
-status_e gecko_pool_enqueue_tasks(thread_task* task, thread_pool* pool, size_t num_threads, task_handle* hndl);
+// Sets the number of active threads to num_threads.
+// Currently working threads are terminated after there task is completed.
+status_e thread_pool_resize(thread_pool* pool, size_t num_threads);
+
+// Add multiple tasks to be executed. Their progress is tracked by a single handle.
+// hndl can be a nullptr.
+status_e gecko_pool_enqueue_tasks(thread_task* task, thread_pool* pool, size_t num_tasks, task_handle* hndl);
 status_e gecko_pool_enqueue_task(thread_task* task, thread_pool* pool, task_handle* hndl);
-size_t gecko_pool_create_group_id();
+
+// Waits until the tasks referenced by hndl are completed.
 status_e thread_pool_wait_for_task(thread_pool* pool, task_handle* hndl);
 
 //

### 2018-05-16 src/thread_pool.c

@@ -2,9 +2,6 @@
 #include <stdlib.h>
 #include <string.h>
 
-// generate unique ids
-static size_t counter_tasks = 0;
-
 //
 //  EXTERNAL METHODS
 //
@@ -39,11 +36,11 @@ thread_pool* thread_pool_create(size_t num_threads) {
 }
 
 void thread_pool_free(thread_pool* pool) {
-    // Update all status and wait until the threads finish
+    // Update all status
     for(size_t i=0; i < pool->size; ++i) {
       pool->thread_infos[i]->status = thread_status_will_terminate;
     }
-
+    // wait for threads to finish
     for(size_t i=0; i < pool->size; ++i) {
       pthread_join(pool->pool[i], NULL);
       free(pool->thread_infos[i]);
@@ -79,6 +76,7 @@ status_e thread_pool_resize(thread_pool* pool, size_t num_threads)
 	}
   else if(num_threads < pool->size){
     for(size_t i= num_threads; i < pool->size; ++i) {
+      // mark threads for termination
       pool->thread_infos[i]->status = thread_status_will_terminate;
     }
   }
@@ -92,11 +90,11 @@ status_e gecko_pool_enqueue_tasks(thread_task* tasks, thread_pool* pool, size_t 
   for(; ind < pool->task_state_capacity && pool->task_group_states[ind].task_count; ++ind);
   if(ind == pool->task_state_capacity) return status_failed;
 
+  // increment generation first to always be identifiable as finished
   ++pool->task_group_states[ind].generation;
   pool->task_group_states[ind].task_count = num_tasks;
   
   for(size_t i= 0; i < num_tasks; i++) {
-    //Assuming queue will conatin something like this, this will not fail
     tasks[i].group_id = ind;
     priority_queue_push(pool->waiting_tasks, &tasks[i], tasks[i].priority); 
   }
@@ -111,10 +109,6 @@ status_e gecko_pool_enqueue_tasks(thread_task* tasks, thread_pool* pool, size_t 
 
 status_e gecko_pool_enqueue_task(thread_task* task, thread_pool* pool, task_handle* hndl) {
   return gecko_pool_enqueue_tasks(task, pool, 1, hndl);
-}
-
-size_t gecko_pool_create_group_id() {
-  return counter_tasks++;
 }
 
 status_e thread_pool_wait_for_task(thread_pool* pool, task_handle* hndl) {

### 2018-06-03 Presentations/Paper/paper_thread_pool.tex
Add preliminaries to paper

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{multicol}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Thread Pool in GeckoDB/BOLSTER - Integrating a Thread Pool into a high performance Database System \\
}

\author{\IEEEauthorblockN{Marten Wallewein-Eising}
	\IEEEauthorblockA{\textit{Otto-von-Guericke University} \\
	Magdeburg, Germany \\
	marten.wallewein-eising@st.ovgu.de}\and
	\IEEEauthorblockN{Robert Jendersie}
	\IEEEauthorblockA{\textit{Otto-von-Guericke University} \\
		Magdeburg, Germany \\
		robert.jendersie@ovgu.de}\and
	\IEEEauthorblockN{Johannes Wuensche}
	\IEEEauthorblockA{\textit{Otto-von-Guericke University} \\
	Magdeburg, Germany \\
	johannes.wuensche@st.ovgu.de}\and
	\IEEEauthorblockN{Johann Wagner}
	\IEEEauthorblockA{\textit{Otto-von-Guericke University} \\
	Magdeburg, Germany \\
	johann.wagner@st.ovgu.de}
}

\maketitle

\begin{abstract}

\end{abstract}

\begin{IEEEkeywords}
Thread Pool, GeckoDB, BOLSTER
\end{IEEEkeywords}

\section{Introduction}
Since the amount of data that is stored and processed by modern database systems is growing fast, sequential data processing as only possibility is inconceivable. Applications have to process data in parallel to reach sufficient throughput to fulfil appropriate requirements.

Parallel data processing can be achieved by different approaches like instruction and data parallelism or multi threading. In this paper, we focus on multi threading by implementing a thread pool for the graph database system GeckoDB. The thread pool will be integrated into BOLSTER, a high performance library for parallel execution of primitives like for or filter on large data sets. In the current implementation, BOLSTER creates a fix number of threads for each call of a primitive. This approach is called \emph{thread-per-request}. Since many primitives are executed at the same time, many drawbacks arise from this implementation. 
First of all, the creation of threads comes along with overhead like stack initialisation and memory allocation. Secondly, creating a huge number of threads simultaneously may lead to large context switch overhead of the scheduler. Additionally, debugging and profiling applications that create many threads during runtime is 	time-consuming.

To overcome these drawbacks, we integrate an optimised thread pool in BOLSTER. Along with the implementation, we measure the performance of the primitives to determine the thread pool overhead. Additionally, we measure metrics like \emph{idle and job time} of threads to evaluate correct thread pool sizes for the considered use cases.  In this work we make the following contributions:
\begin{itemize}
	\item We describe our design and implementation of the thread pool 
	\item We evaluate the possibility to wait for a group of task in the calling thread
	\item We compare our thread pool against the existing implementation in BOLSTER
\end{itemize}
We organized the rest of the paper as follows. In Section 2, we give preliminaries about the considered task configuration and about thread safe access of memory. In Section 3, we show our design and implementation of the thread pool and examine our experimental environment in Section 4. In Section 5, we describe the results of our performance evaluation. In Section 6, we name related work and state our conclusion and future work in Section 7.

\section{Preliminaries}
In this section, we define our configuration of tasks that are processed by the thread pool and state difficulties of synchronizing thread access to memory.

\subsection{Task Configuration}
We define a Task as a structure containing data that has to be processed and an operation that has to be executed on the data. In this work, we define tasks as \emph{independent}, which means tasks do not have dependencies on other tasks and can be processed independently. Furthermore, we expect the data passed to two tasks are stored in different memory locations. Consequently, while executing the task operation, threads do not access the same memory locations. Additionally, we only consider not preemtable tasks. Once a task is assigned to a thread, the thread will finish the operation of the task before getting a new one.

\subsection{Synchronizing Memory Access from Threads}
Parallelism with multi threading works great as long as each thread works on a separate memory area. During task scheduling, the scheduler has to know the state each thread has. This can be solved with signal handling or by writing the current thread state into main memory. In this work, we implement the second approach to avoid having another thread that only schedules tasks to other threads.

Threadsafe access to memory can be achieved by different approaches. Firstly, a mutex thread can manage the access for multiple other threads. Secondly, atomic operations ensure the safe access from different threads to the same memory. Since using a mutex thread can decrease access performance, we decided to execute atomic operations on the memory containing the thread state informations. For example, the function \emph{atomic\_compare\_exchange\_strong} performs a compare and, if the result is true, an exchange of the memory in one atomic operation.

%\begin{figure}
%	\includegraphics[width=0.4\textwidth]{figure_1.png}
%	\caption{Coherence between data, instructions, and the results of SISD and SIMD.}
%	\label{fig}
%\end{figure}
% SIMD explanation

%\begin{table*}[htbp]
%	\caption{SIMD instructions from Streaming SIMD Extensions 2 (SSE2)}
%	\begin{center}
%		\begin{tabular}{|c|c|}
%			\hline
%			\textbf{SIMD instruction}&\textbf{Explanation}\\
%			\hline
%			\_\_m128i \_mm\_load\_si128 (\_\_m128i *p) & Loads a 128-bit value. Returns the value loaded into a variable representing a register.\\
%			\_\_m128i \_mm\_cmpgt\_epi32 (\_\_m128i a, \_\_m128i b) & Compares 4 signed 32-bit integers in a and 4 signed 32-bit integers
%			in b for greater-than.\\
%			\_\_m128i \_mm\_set1\_epi32(int i) & Sets the four signed 32-bit integer values to i.\\
%			\hline
%		\end{tabular}
%		\label{tab1}
%	\end{center}
%
%\end{table*}

\section{Implementation}
In this section,

\section{Experimental Environment}

\section{Evaluation}

\section{Related Work}

\section{Conclusion and Future Work}

\bibliographystyle{splncs}
\bibliography{paper_thread_pool}

\end{document}

### 2018-06-04 include/thread_pool.h
Add enqueue_and_waiting function to thread pool

@@ -89,8 +89,13 @@ status_e thread_pool_resize(thread_pool* pool, size_t num_threads);
 status_e thread_pool_enqueue_tasks(thread_task* task, thread_pool* pool, size_t num_tasks, task_handle* hndl);
 status_e thread_pool_enqueue_task(thread_task* task, thread_pool* pool, task_handle* hndl);
 
+// Add multiple tasks to be executed. Waits until all passed tasks are finished. 
+// The main thread also participates in task execution
+status_e thread_pool_enqueue_tasks_wait(thread_task* task, thread_pool* pool, size_t num_tasks);
+
 // Waits until the tasks referenced by hndl are completed.
 status_e thread_pool_wait_for_task(thread_pool* pool, task_handle* hndl);
+
 // Waits until all tasks currently in the queue are executed.
 // The main thread also participates in task execution.
 status_e thread_pool_wait_for_all(thread_pool* pool);

### 2018-06-04 testing/test_runner.c

@@ -1,0 +1,40 @@
+#include <check.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include "tests/thread_pool/test_thread_pool.h"
+
+Suite * thread_pool_suite(void)
+{
+    Suite *s;
+    TCase *tc_core;
+
+    s = suite_create("Thread Pool");
+
+    /* Core test case */
+    tc_core = tcase_create("Core");
+
+    tcase_add_test(tc_core, thread_pool_create_);
+    tcase_add_test(tc_core, thread_pool_name);
+    //tcase_add_test(tc_core, thread_pool_tasks);
+    tcase_add_test(tc_core, thread_pool_tasks_wait);
+    
+
+    suite_add_tcase(s, tc_core);
+
+    return s;
+}
+
+int main(void)
+{
+    int number_failed;
+    Suite *s;
+    SRunner *sr;
+
+    s = thread_pool_suite();
+    sr = srunner_create(s);
+
+    srunner_run_all(sr, CK_NORMAL);
+    number_failed = srunner_ntests_failed(sr);
+    srunner_free(sr);
+    return (number_failed == 0) ? EXIT_SUCCESS : EXIT_FAILURE;
+}

### 2018-06-04 testing/tests/thread_pool/test_thread_pool.c

@@ -1,102 +1,0 @@
-#include <check.h>
-#include <stdio.h>
-#include "../../../include/thread_pool.h"
-
-// Test the creation of thread pools
-START_TEST(thread_pool_create_) 
-{
-    thread_pool* pool = thread_pool_create(2);
-    ck_assert(pool);
-
-    thread_pool_free(pool);
-    //ck_assert(!pool);
-}
-END_TEST
-
-// Test the naming of thread pools
-START_TEST(thread_pool_name) 
-{
-    thread_pool* pool = thread_pool_create_named(2, "ThreadPool");
-    ck_assert(pool);
-    ck_assert_str_eq(pool->name, "ThreadPool");
-
-    thread_pool_free(pool);
-    
-}
-END_TEST
-
-void work(void* args)
-{
-	int* res = args;
-	for(; *res > 0; (*res)--)
-    {
-        
-    }
-}
-
-START_TEST(thread_pool_tasks){
-
-    thread_pool* pool = thread_pool_create(2);
-    task_handle hndl;
-
-    int test[] = {1000000, 1000000, 1000000, 1000000, 1000000, 1000000 };
-    
-    for(int i = 0; i < 6; i++){
-        
-        thread_task* test_task = (thread_task*)malloc(sizeof(thread_task));
-        test_task->args = (void*)&test[i];
-        test_task->routine = work;
-
-        thread_pool_enqueue_task(test_task, pool, &hndl);
-    }
-
-    thread_pool_wait_for_all(pool);
-
-    for(int i = 0; i < 6; i++){
-        //printf("%d\n", test[i]);
-        ck_assert_int_eq(test[i], 0);
-    }
-    thread_pool_free(pool);
-    
-}
-END_TEST
-
-Suite * thread_pool_suite(void)
-{
-    Suite *s;
-    TCase *tc_core;
-
-    s = suite_create("Thread Pool");
-
-    /* Core test case */
-    tc_core = tcase_create("Core");
-
-    tcase_add_test(tc_core, thread_pool_create_);
-    tcase_add_test(tc_core, thread_pool_name);
-    tcase_add_test(tc_core, thread_pool_tasks);
-
-    suite_add_tcase(s, tc_core);
-
-    return s;
-}
-
-int main(void)
-{
-    int number_failed;
-    Suite *s;
-    SRunner *sr;
-
-    s = thread_pool_suite();
-    sr = srunner_create(s);
[…]

### 2018-06-04 testing/tests/thread_pool/test_thread_pool.h

@@ -1,0 +1,121 @@
+#include <check.h>
+#include <stdio.h>
+#include <math.h>
+#include "../../../include/thread_pool.h"
+
+// Test the creation of thread pools
+START_TEST(thread_pool_create_) 
+{
+    thread_pool* pool = thread_pool_create(2);
+    ck_assert(pool);
+
+    thread_pool_free(pool);
+    //ck_assert(!pool);
+}
+END_TEST
+
+// Test the naming of thread pools
+START_TEST(thread_pool_name) 
+{
+    thread_pool* pool = thread_pool_create_named(2, "ThreadPool");
+    ck_assert(pool);
+    ck_assert_str_eq(pool->name, "ThreadPool");
+
+    thread_pool_free(pool);
+    
+}
+END_TEST
+
+void work(void* args)
+{
+	int* res = args;
+
+    double exp = *res;
+
+	double v = 0.0;
+	for(double i = 0; i < 100024; ++i)
+	{
+		v += pow(i, 1 / exp);
+	}
+
+	for(; *res > 0; (*res)--)
+    {
+        
+    }
+
+    printf("%s\n", "Work finished");
+
+}
+
+void work_large(void* args)
+{
+	double* res = args;
+	double exp = *res;
+
+	double v = 0.0;
+	for(double i = 0; i < 100024; ++i)
+	{
+		v += pow(i, 1 / exp);
+	}
+	*res = v;
+}
+
+START_TEST(thread_pool_tasks){
+
+    thread_pool* pool = thread_pool_create(2);
+    task_handle hndl;
+
+    int test[] = {1000000, 1000000, 1000000, 1000000, 1000000, 1000000 };
+    
+    thread_task* test_task = (thread_task*)calloc(6, sizeof(thread_task));
+
+    for(int i = 0; i < 6; i++){
+        
+        test_task[i].args = (void*)&test[i];
+        test_task[i].routine = work;
+    }
+
+    thread_pool_enqueue_tasks(test_task, pool, 6, &hndl);
+    printf("%s", "Waiting");
+    thread_pool_wait_for_task(pool, &hndl);
+    printf("%s", "Waiting finished");
+
+    for(int i = 0; i < 6; i++){
+        ck_assert_int_eq(test[i], 0);
+    }
+    thread_pool_free(pool);
+    
+}
+END_TEST
+
+START_TEST(thread_pool_tasks_wait){
+
[…]

### 2018-06-04 Presentations/Paper/paper_thread_pool.tex

@@
 
+%\begin{itemize}
+%	\item We compare different adaptions of index structures to fulfill requirements of modern database systems
+%	\item We highlight the usage of SIMD and the cache line adaptions in all approaches
+%	\item We extract the important points of these adaptions and their performance impact on the index structures
+%\end{itemize}
+
+% Paper structure
+
 \section{Preliminaries}
-In this section, we define our configuration of tasks that are processed by the thread pool and state difficulties of synchronizing thread access to memory.
-
-\subsection{Task Configuration}
-We define a Task as a structure containing data that has to be processed and an operation that has to be executed on the data.
-In this work, we define tasks as \emph{independent}, which means tasks do not have dependencies on other tasks and can be processed independently.
-Furthermore, we expect the data passed to two tasks are stored in different memory locations.
-Consequently, while executing the task operation, threads do not access the same memory locations.
-Additionally, we only consider not preemtable tasks.
-Once a task is assigned to a thread, the thread will finish the operation of the task before getting a new one.
-
-\subsection{Synchronizing Memory Access from Threads}
-Parallelism with multi threading works great as long as each thread works on a separate memory area.
-During task scheduling, the scheduler has to know the state each thread has.
-This can be solved with signal handling or by writing the current thread state into main memory.
-In this work, we implement the second approach to avoid having another thread that only schedules tasks to other threads.
-
-Threadsafe access to memory can be achieved by different approaches.
-Firstly, a mutex thread can manage the access for multiple other threads.
-Secondly, atomic operations ensure the safe access from different threads to the same memory.
-Since using a mutex thread can decrease access performance, we decided to execute atomic operations on the memory containing the thread state informations.
-For example, the function \emph{atomic\_compare\_exchange\_strong} performs a compare and, if the result is true, an exchange of the memory in one atomic operation.
-
 %\begin{figure}
@@
 \section{Implementation}
-In this section,
 

### 2018-06-04 Presentations/Paper/paper_thread_pool.tex
Continue Paper

--- 2018-06-03 Presentations/Paper/paper_thread_pool.tex
@@
 
-\author{\IEEEauthorblockN{Marten Wallewein-Eising}
-	\IEEEauthorblockA{\textit{Otto-von-Guericke University} \\
-	Magdeburg, Germany \\
-	marten.wallewein-eising@st.ovgu.de}\and
+\author{
 	\IEEEauthorblockN{Robert Jendersie}
@@
 		Magdeburg, Germany \\
-		robert.jendersie@ovgu.de}\and
+		robert.jendersie@ovgu.de} \and
+
 	\IEEEauthorblockN{Johannes Wuensche}
@@
 	Magdeburg, Germany \\
-	johannes.wuensche@st.ovgu.de}\and
+	johannes.wuensche@st.ovgu.de} \and
+
 	\IEEEauthorblockN{Johann Wagner}
@@
 	Magdeburg, Germany \\
-	johann.wagner@st.ovgu.de}
+	johann.wagner@st.ovgu.de} \and
+
+	\IEEEauthorblockN{Marten Wallewein-Eising}
+	\IEEEauthorblockA{\textit{Otto-von-Guericke University} \\
+	Magdeburg, Germany \\
+	marten.wallewein-eising@st.ovgu.de}
 }
@@
 \section{Implementation}
-In this section,
+In this section, we describe in detail the design and implementation of our thread pool.
+We show how tasks are enqueued regarding their priority and how the scheduling is implemented.
+Additionally, we state our thread metrics and how they are integrated in the thread pool design.
+
+\subsection{Architecture of the Thread Pool}
+Compared to simply create threads on demand, managing threads in a pool comes along with memory and CPU overhead.
+
+\subsection{Enqueueing Tasks}
+\begin{figure}
+	\includegraphics[width=0.5\textwidth]{img/pool_queue.png}
+	\caption{Workflow of task enqueueing and execution in the thread pool}
+	\label{fig}
+\end{figure}
+
+\subsection{Waiting for Tasks}
+As mentioned in the previous section, ids are assigned to the tasks before they are enqueued. 
+\subsection{Thread metrics}
 

### 2018-06-04 Presentations/Paper/paper_thread_pool.tex
Add more figures to paper

@@
 
+\begin{figure*}[htbp]
+	\includegraphics[width=1.0\textwidth]{img/pool_structure.png}
+	\caption{Design of the thread pool system}
+	\label{fig0}
+\end{figure*}
+
 \subsection{Task Configuration}
@@
 	\caption{Workflow of task enqueueing and execution in the thread pool}
-	\label{fig}
+	\label{fig1}
 \end{figure}
@@
 \subsection{Waiting for Tasks}
+\begin{figure}
+	\includegraphics[width=0.5\textwidth]{img/waitingconcept.png}
+	\caption{Concept of waiting for tasks using a slotmap}
+	\label{fig2}
+\end{figure}
 As mentioned in the previous section, ids are assigned to the tasks before they are enqueued. 
@@
 \section{Conclusion and Future Work}
-
+\cite{pyarali2001evaluating}
 \bibliographystyle{splncs}

### 2018-06-04 Presentations/Paper/paper_thread_pool.tex
Start writing chapter implementation in paper

@@
 Consequently, while executing the task operation, threads do not access the same memory locations.
+
+Each task can be enqueued with a priority.
+The priority zero is the highest and the task will be processed by the next free thread.
+The higher the priority of the task is, the further it is placed behind in the queue.
 Additionally, we only consider not preemtable tasks.
@@
 Secondly, atomic operations ensure the safe access from different threads to the same memory.
+We use both approaches for different parts of the thread pool.
 Since using a mutex thread can decrease access performance, we decided to execute atomic operations on the memory containing the thread state informations.
 For example, the function \emph{atomic\_compare\_exchange\_strong} performs a compare and, if the result is true, an exchange of the memory in one atomic operation.
+We use a mutex thread to synchronize the enqueueing of tasks into the priority queue.
 
@@
 Compared to simply create threads on demand, managing threads in a pool comes along with memory and CPU overhead.
+The thread pool must know information about the state each thread has and his assigned task.
+To measure metrics of threads and tasks, additional memory for threads and tasks is required.
+
+In Figure 1, we show our design of the whole thread pool system, containing the thread pool itself, the task priority queue and the performance monitoring.
+Since measuring performance metrics lead to memory and CPU overhead, we decided to exclude the performance measurements from the thread pool to make it optional.
+The performance monitoring can be activated with an additional function call of \textbf{TODO: Insert function here}.
+Consequently, in the target database system, the designers can decide for each thread pool instance, if performance monitoring should be applied.
+
+The thread pool system includes an array of threads and a priority queue to store the passed tasks.
+We decided to implement the thread pool using the POSIX thread library to provide the thread pool for multiple operation systems like Linux and Unix.
+Additionally, we avoid using custom compiler flags to ensure that the thread pool can be compiled with different compilers.
+The thread pool itself contains a variable number of threads.
+Since XY and XZ show the importance of accurate thread pool size, we add a resizing function for our thread pool, which enables to change the amount of threads at runtime.
+
+\subsection{Thread Pool and Task Queue}
+Threads and tasks are two major entities in the thread pool system.
+As we can observe from Figure 1, thread pool and task queue are two data structures used to store the required information related to the threads and the submitted tasks, respectively.
+Each spawned thread can be either working on a task or waiting for the next task.
+Consequently, we define the states of a thread to \emph{busy} and \emph{waiting}.
 
@@
 As mentioned in the previous section, ids are assigned to the tasks before they are enqueued.
+
+\subsection{Managing Thread States}
+- To handle the shrinkage of the pool while threads are processing tasks, we change the state
+
 \subsection{Thread metrics}

### 2018-06-04 Presentations/Paper/paper_thread_pool.tex
Continue writing implementation section of paper

@@
 	\includegraphics[width=1.0\textwidth]{img/pool_structure.png}
-	\caption{Design of the thread pool system}
+	\caption{Design of the thread pool system TODO: Fix Image!}
 	\label{fig0}
@@
 Each spawned thread can be either working on a task or waiting for the next task.
-Consequently, we define the states of a thread to \emph{busy} and \emph{waiting}.
+Consequently, we define the states of a thread as \emph{busy} and \emph{waiting}.
+A busy thread changes to waiting after processing its task if no other task is enqueued.
+Waiting tasks do busy waiting until there are new tasks available.
 
-\subsection{Enqueueing Tasks}
+Each assigned task contains a function pointer to a routine and a pointer to data.
+We implemented the task queue as a generic priority queue that uses a mutex thread to enqueue tasks threadsafe.
+Every time a new task is enqueued to the task queue, two scenarios can occur.
+If the new task has a priority, it is placed in the right slot of the task queue regarding its priority and the subsequent tasks are relocated.
+If the  new task has no priority, it is appended to the end of the queue.
+Furthermore, tasks can refer to a handle, which is used by the thread pool to wait for a specific amount of tasks.
+
+\subsection{Processing Tasks}
+\textbf{TODO: maybe table of provided functions} \\
+We define the processing of tasks as a procedure of 3-4 steps.
+In Figure 2, we show the flow of task enqueueing and execution in the thread pool system.
+In the first step, tasks are passed as a parameter to the enqueue function provided by the thread pool.
+The tasks have to be created before passing and later on freed after processing by the calling function.
+We provide different functions to enqueue a single task or an array of tasks.
+
+In the next step, the tasks will be enqueued into the priority queue of the thread pool.
+Optionally, a handle can be passed with the tasks or generated by the thread pool to force the calling function to wait for the tasks.
+After enqueueing the tasks to the priority queue, the waiting threads in the thread pool pop the enqueued tasks from the priority queue and executes the routine of the task.
+This happens asynchronously to the calling function and therefore is not considered as an own step of the process.
+
+The third step is only executed if the tasks refer to a valid handle and waiting is enforced by the calling function.
+This can be applied either by executing a provided waiting function from the thread pool or by calling \emph{thread\_pool\_enqueue\_tasks\_wait}.
+We will present our design of waiting for a specific amount of tasks in Section 3.D.
+The last step only contains the return to the calling function.
+
 \begin{figure}
@@
 \begin{figure}
-	\includegraphics[width=0.5\textwidth]{img/waitingconcept.png}
+	\includegraphics[width=0.4\textwidth]{img/waitingconcept.png}
 	\caption{Concept of waiting for tasks using a slotmap}

### 2018-06-04 Presentations/Paper/paper_thread_pool.tex
Continue writing implementation section of paper

@@
 Furthermore, tasks can refer to a handle, which is used by the thread pool to wait for a specific amount of tasks.
+\begin{table*}[htbp]
+	\caption{Functions provided by the thread pool to process tasks}
+	\begin{center}
+		\begin{tabular}{ c c }
+			\hline
+			\textbf{Function Name}&\textbf{Explanation}\\
+			\hline
+			thread\_pool\_enqueue\_task & Enqueue task with optional handle \\
+			thread\_pool\_enqueue\_tasks & Enqueue tasks with optional handle \\
+			thread\_pool\_wait\_for\_task & Waits until the tasks referenced by the handle are completed. \\
 
+			thread\_pool\_enqueue\_tasks\_wait & Enqueue tasks and wait until they are finished. The main thread also participates in task execution\\ 
+			thread\_pool\_wait\_for\_all & Wait until all enqueued tasks are finished. The main thread also participates in task execution \\ \hline
+		\end{tabular}
+		\label{tab1}
+	\end{center}
+\end{table*}
 \subsection{Processing Tasks}
-\textbf{TODO: maybe table of provided functions} \\
-We define the processing of tasks as a procedure of 3-4 steps.
-In Figure 2, we show the flow of task enqueueing and execution in the thread pool system.
+Processing tasks in a thread pool is a balancing act between generalization the of tasks and the performance of their execution.
+Therefore, we provide a set of functions to process task in the thread pool to match different requirements of BOLSTER.
+In Table 1, we show the functions provided by our thread pool to process tasks.
+One of the requirements of BOLSTER is waiting until a set of tasks is finished.
+To achieve the same performance as the baseline implementation reaches, we let the main thread first participate in the execution of the tasks and then wait until all tasks with the same handle are finished.
+
+To show the processing of tasks, we consider the function \emph{thread\_pool\_enqueue\_tasks\_wait}.
+In Figure 2, we show the flow of task enqueueing and execution in the thread pool system using this function.
 In the first step, tasks are passed as a parameter to the enqueue function provided by the thread pool.
@@
 \subsection{Waiting for Tasks}
+
 \begin{figure}
@@
 \end{figure}
-As mentioned in the previous section, ids are assigned to the tasks before they are enqueued.
 

### 2018-06-05 Presentations/Paper/paper_thread_pool.tex
Add Waiting for tasks to paper

@@
 
-%\begin{figure}
-%	\includegraphics[width=0.4\textwidth]{figure_1.png}
-%	\caption{Coherence between data, instructions, and the results of SISD and SIMD.}
-%	\label{fig}
-%\end{figure}
-% SIMD explanation
-
-%\begin{table*}[htbp]
-%	\caption{SIMD instructions from Streaming SIMD Extensions 2 (SSE2)}
-%	\begin{center}
-%		\begin{tabular}{|c|c|}
-%			\hline
-%			\textbf{SIMD instruction}&\textbf{Explanation}\\
-%			\hline
-%			\_\_m128i \_mm\_load\_si128 (\_\_m128i *p) & Loads a 128-bit value. Returns the value loaded into a variable representing a register.\\
-%			\_\_m128i \_mm\_cmpgt\_epi32 (\_\_m128i a, \_\_m128i b) & Compares 4 signed 32-bit integers in a and 4 signed 32-bit integers
-%			in b for greater-than.\\
-%			\_\_m128i \_mm\_set1\_epi32(int i) & Sets the four signed 32-bit integer values to i.\\
-%			\hline
-%		\end{tabular}
-%		\label{tab1}
-%	\end{center}
-%
-%\end{table*}
-
 \section{Implementation}
@@
 \subsection{Waiting for Tasks}
+As mentioned before, the thread pool provides the functionality to let the calling function participate in the task execution and wait until all passed tasks are processed.
+To wait for a set of tasks, it is not sufficient to look firstly in the priority queue for the tasks and then in the state of the thread pool, since popping tasks out of the queue is not an atomic operation.
+Consequently, tasks can have a state between being enqueued in the queue and being executed by a thread.
+
+To solve this problem, we implement the waiting for a set of tasks using a slotmap.
+In Figure 3, we show the concept of our slotmap.
+During the enqueueing process of tasks, a set of tasks are referred to a slot in the slotmap.
+Each slot contains a number of open tasks to be finished and a generation counter.
+We name the number of open task \emph{openTaskCount} and the generation counter \emph{genCount}.
+Every time a set of tasks is enqueued, a slot with $openTaskCount = 0$ and the \emph{genCount} is increases.
+Every time a thread finishes a task, the openTaskCount of the referred slot is decreased.
+As mentioned before, multiple tasks can refer to a single handle.
+This handle also refers to a slot and has additionally assigned a generation value.
+The waiting algorithm loops over the following conditions before returning to the calling function: It checks if the number of open tasks is zero or if the generation counter of this slot has changed.
+If one of these conditions is true, the waiting algorithm returns to the calling function.
+
+To clarify the procedure of the waiting algorithm, we consider the second slot in Figure 3.
+The tasks $10, 12, $ and $13$ as well has the first handle refer to this slot.
+Consequently, these tasks were enqueued referring to the first handle.
+Before any of the three tasks is processed, the slot has the following state: $openTaskCount = 3$ and $genCount =  2$.
+Since the waiting thread does not know when tasks are executed due to the asynchronous behaviour, the waiting algorithm starts looping as long as the following condition is true: $openTaskCount \neq 0$ \&\& $genCount = 2$.
+For each finished task, $openTaskCount$ is decreased.
+It is not sufficient to check only $openTaskCount \neq 0$, since the slot can be reused before the waiting algorithms noticed that $openTaskCount = 0$.
+Consequently, we check if the slot stays on the same generation.
+If the slot is reused, $genCount$ increases and the second part of the condition returns false.
 

### 2018-06-05 Presentations/Paper/paper_thread_pool.tex
Add parts of performance monitoring to paper

@@
 \subsection{Managing Thread States}
+TODO: Remove this subsection? \\
 - To handle the shrinkage of the pool while threads are processing tasks, we change the state
 
-\subsection{Thread metrics}
+\subsection{Performance monitoring}
+The performance monitoring component of the thread pool system is responsible for collecting data about threads, tasks, and the whole thread pool system.
+We will use the information collected in this component for analysis and performance optimization in later stages.
+As mentioned before, the performance monitoring is an optional component, since evaluating the following statistics can reduce the execution performance of our thread pool.
+Collecting statistics can be activated to a specific thread pool instance by calling the function \emph{thread\_pool\_statistics\_start}.
+We divide the collected statistics into the sections \emph{task statistics}, \emph{thread statistics}, and \emph{thread pool statistics}.
+
+\subsubsection{Task Statistics}
+Collecting statistics about the time tasks spend in the priority queue or during the execution is useful to tune the thread pool configuration for better performance results.
+In Table \ref{tab2} we show the statistics that are collected for each task.
+
+\begin{table}[htbp]
+	\caption{Task statistics with their internal names and descriptions}
+	\begin{center}
+		\begin{tabular}{ c c }
+			\hline
+			\textbf{Statistic name}&\textbf{Description}\\
+			\hline
+			enqueue\_time & Time when the task was enqueued \\
+			execution\_time & Time when the task execution begins \\
+			complete\_time & Time when the task execution is finished \\
+			\hline
+		\end{tabular}
+		\label{tab2}
+	\end{center}
+\end{table}
+
+How long a task waits in the queue before being executed is a good metric to adjust task priority or the thread pool size.
+For example, if an enqueued task with priority 0 still has a large difference between \emph{enqueue\_time} and \emph{execution\_time}, the thread pool size may be too small.
+In section 6, we state related work that covers the dynamic resizing of thread pools.
+
+\subsubsection{Thread Statistics}
+
+\begin{table}[htbp]
+	\caption{Thread statistics with their internal names and descriptions}
+	\begin{center}
+		\begin{tabular}{ c c }
+			\hline
+			\textbf{Statistic name}&\textbf{Description}\\
+			\hline
+			waiting\_time & Time the thread spend waiting (ms) \\
+			busy\_time & Time the thread spend executing tasks (ms)\\
+			task\_count & Number of tasks the thread has executed \\
+			\hline
+		\end{tabular}
+		\label{tab3}
+	\end{center}
+\end{table}
+
+\subsubsection{Thread Pool Statistics}
+
+\begin{table}[htbp]
+	\caption{Thread pool statistics with their internal names and descriptions}
+	\begin{center}
+		\begin{tabular}{ c c }
+			\hline
+			\textbf{Statistic name}&\textbf{Description}\\
+			\hline
+			working\_time & Time the thread pool runs (ms) \\
+			task\_complete\_count & Number of tasks the thread pool has completed\\
+			task\_enqueued\_count & Number of tasks the thread pool has enqueued \\
+			avg\_task\_time & Average time span from enqueueing until \\
+			& execution of tasks \\
+			\hline
+		\end{tabular}
+		\label{tab4}
+	\end{center}
+\end{table}
 

### 2018-06-05 Presentations/Paper/paper_thread_pool.tex
Add performance monitoring to paper

@@
 For example, if an enqueued task with priority 0 still has a large difference between \emph{enqueue\_time} and \emph{execution\_time}, the thread pool size may be too small.
-In section 6, we state related work that covers the dynamic resizing of thread pools.
+Otherwise, if a large number of tasks without high priority have small differences between \emph{enqueue\_time} and \emph{execution\_time}, this may indicate that the thread pool size is too large.
+A better indicator of not optimized thread pool sizes are statistics per thread.
 
 \subsubsection{Thread Statistics}
+As mentioned before, adjusting the thread pool size based on task statistics may not be very precise.
+Consequently, collecting statistics per thread is necessary for a good evaluation of the relation between tasks and threads.
+In Table \ref{tab3}, we show our statistics collected for each thread in a thread pool instance.
 
@@
 
+The \emph{busy\_time} of a thread strongly depends on the executed tasks and therefor is a metric that gives less information about the thread pool itself.
+In contrast, the \emph{waiting\_time} is a good indicator to reveal less used threads in the thread pool to determine an optimal amount of threads for the current use case.
+Furthermore, the \emph{task\_count} of thread compared to the number of tasks processed by the thread pool indicates the degree of usage of the thread.
+To give further possibilities for optimizing the thread pool, we collect statistics for each thread pool instance.
+
 \subsubsection{Thread Pool Statistics}
+To consider the thread pool instance, an important metric is the average waiting time of tasks.
+In order to optimize the thread pool size, a balanced relation between the waiting time of tasks and the number of threads must be found.
+Furthermore, at a specific time, the relation between enqueued and completed tasks can be interesting.
+In Table \ref{tab4}, we show our statistics collected for each thread pool instance.
 
@@
 			task\_enqueued\_count & Number of tasks the thread pool has enqueued \\
-			avg\_task\_time & Average time span from enqueueing until \\
+			avg\_complete\_time & Average time span from enqueueing until \\
 			& execution of tasks \\
+			avg\_wait\_time & Average time of tasks spend in the queue \\
 			\hline
@@
 \end{table}
+
+We will use these statistics to adapt our thread pool implementation to different use cases later on.
+In the next sections, we define our experimental environment, the implementation of our tests including the measured metrics and the test results.
 

### 2018-06-06 include/thread_pool.h
Move statistics to thread_pool_monitoring

@@ -12,7 +12,7 @@
 #include "status.h"
 #include "priority_queue.h"
 #include "thread_status.h"
-#include <time.h>
+#include "thread_pool_statistics.h"
 
 //
 //	GLOBAL VARIABLES
@@ -30,6 +30,7 @@ typedef struct thread_task {
 	task_routine routine;
 	size_t group_id;
 	size_t priority;
+	task_stats* statistics; 
 } thread_task;
 
 struct __thread_information;
@@ -56,23 +57,25 @@ typedef struct thread_pool {
 //    thread_status_e* thread_status;
 	struct __thread_information** thread_infos;
 	thread_task** thread_tasks;
+	thread_pool_stats* statistics; 
 } thread_pool;
 
 typedef struct __thread_information {
-	char name[16];
+	char name[12];
 	thread_pool* pool;
 	size_t id;
 	atomic_int status;
 	struct timespec creation_time;
 	double time_spend_idle; // actually the time not spend directly executing a task
+	thread_stats* statistics;
 } __thread_information;
 
 //
 //	EXTERNAL METHODS
 //
 
-thread_pool* thread_pool_create(size_t num_threads);
-thread_pool* thread_pool_create_named(size_t num_threads, const char* name);
+thread_pool* thread_pool_create(size_t num_threads, int enable_monitoring);
+thread_pool* thread_pool_create_named(size_t num_threads, const char* name, int enable_monitoring);
 // Releases all resources hold by the threadpool. 
 // Currently working threads may finish but tasks left in the queue will be discarded.
 void thread_pool_free(thread_pool* pool);
@@ -100,8 +103,6 @@ status_e thread_pool_wait_for_task(thread_pool* pool, task_handle* hndl);
 // The main thread also participates in task execution.
 status_e thread_pool_wait_for_all(thread_pool* pool);
 
-// Returns the average fraction of time the active threads have been working.
-double thread_pool_get_time_working(thread_pool* pool);
 //
 //	INTERNAL METHODS
 //

### 2018-06-06 include/thread_pool_monitoring.h

@@ -1,0 +1,13 @@
+#ifndef THREAD_POOL_MONITORING_H
+#define THREAD_POOL_MONITORING_H
+
+#include "thread_pool.h"
+
+//
+//	EXTERNAL METHODS
+//
+
+// Returns the average fraction of time the active threads have been working.
+double thread_pool_get_time_working(thread_pool* pool);
+
+#endif //THREAD_POOL_MONITORING_H

### 2018-06-06 include/thread_pool_statistics.h

@@ -1,0 +1,38 @@
+#ifndef THREAD_POOL_STATISTICS_H
+#define THREAD_POOL_STATISTICS_H
+
+#include <time.h>
+
+//
+//	STRUCTS
+//
+typedef struct thread_pool_stats {
+	struct timespec creation_time;
+	unsigned int task_enqueued_count;
+	unsigned int task_complete_count; 
+	double avg_complete_time;
+	double avg_wait_time;
+} thread_pool_stats;
+
+typedef struct thread_stats {
+	struct timespec creation_time;
+	double idle_time;
+	double busy_time;
+	size_t task_count;
+} thread_stats;
+
+typedef struct task_stats {
+	struct timespec enqueue_time;
+	struct timespec execution_time;
+	struct timespec complete_time;
+} task_stats;
+
+//
+// INTERNAL METHODS
+//
+
+static inline double __get_time_diff(struct timespec* begin, struct timespec* end) {
+	return (end->tv_sec - begin->tv_sec) + (end->tv_nsec - begin->tv_nsec) / 1000000000.0;
+}
+
+#endif //THREAD_POOL_STATISTICS_H

### 2018-06-06 src/thread_pool_monitoring.c

@@ -1,0 +1,18 @@
+#include "thread_pool_monitoring.h"
+
+//
+// EXTERNAL METHODS
+// 
+
+double thread_pool_get_time_working(thread_pool* pool){
+  struct timespec end;
+  
+  clock_gettime(CLOCK_MONOTONIC, &end);
+  double avg = 0.f;
+  for(size_t i = 0; i < pool->size; ++i){
+    struct timespec begin = pool->thread_infos[i]->statistics->creation_time;
+    double t = __get_time_diff(&begin, &end);
+    avg += t / (t + pool->thread_infos[i]->statistics->idle_time);
+  }
+  return avg / pool->size;
+}

### 2018-06-07 Presentations/Milestone3/presentation.tex
Add Milestone 3 presentation (not complete)

--- 2018-05-09 Presentations/Milestone2/presentation.tex
@@
 \usepackage{graphicx}
+\usepackage{wrapfig,lipsum}
 \usepackage{float}
@@
 \title{Thread Pool in GeckoDB/BOLSTER}
-\subtitle{Milestone II: Concept \& additional literature research}
+\subtitle{Milestone III: Implementation \& evaluation setup}
 \author{Johann Wagner, Johannes Wünsche, Marten Wallewein-Eising, Robert Jenderise, Benedikt Zeilinger}
@@
 \institute{Otto von Guericke University, Magdeburg}
-
-% Milestone II: Concept & additional literature research
 
@@
 \begin{frame}
-	\frametitle{Previously in Thread Pool in GeckoDB/BOLSTER}
-	\begin{itemize}
-		\item Operations are spawning dozens of threads, which are not handled and managed
-		\item Heavy overhead of creating and removing threads
-		\item After some operations profiler dies (too many threads)
-	\end{itemize}
 	\begin{center}
-		\includegraphics[width=0.3\textwidth]{img/thumbs_down.png}
+		\huge Previously...
 	\end{center}
@@
 \begin{frame}
-	\frametitle{Previously in Thread Pool in GeckoDB/BOLSTER}
-	\begin{itemize}
-		\item Reuse of Threads with thread pools.
-		\item Reduce overhead of thread management
-		\item Improve maintainability of parallel code
-	\end{itemize}
-	\begin{center}
-		\includegraphics[width=0.3\textwidth]{img/thumbs_up.png}
-	\end{center}
-\end{frame}
-
-\section{Concept and Design}
-\begin{frame}
-	\frametitle{Concept}
-	\begin{itemize}
-		\item Thread pool with runtime fixed amount of threads and lock free priority queue
-		\item Tasks are not preemtable and independent
-		\item Array of tasks passed to an enqueue function, which waits until all tasks are finished
-		\item Threads without tasks idle until they get a new one
-		\item Monitoring/managing state and assigned task of each thread
-	\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Design}
+	\frametitle{Previously on Thread Pool in GeckoDB/BOLSTER}
 	\begin{center}
@@
 
-\section{Current Progress}
 \begin{frame}
-	\frametitle{Current Progress}
-	We...
+	\frametitle{Previously on Thread Pool in GeckoDB/BOLSTER}
+	\begin{columns}
+		\begin{column}{0.48\textwidth}
 			\begin{itemize}
-		\item Implemented the lock free priority queue 
-		\item Implemented the creation of threads and the assigning of tasks
-		\item Added gtest framework to project and wrote tests for pool and queue
-		\item Began first implementations of waiting for tasks to be finished
+				\item Problem: waiting for a task to complete could lead to a
+				race condition
+				\item This was caused by the uncertainty of a thread\_task
+				during transition between the priority queue and the
+				thread pool
+				\item Solution: slotmap with an atomic counter
 			\end{itemize}
+		\end{column}
+		\begin{column}{0.48\textwidth}
+			\includegraphics[width=1.0\textwidth]{waitingconcept.png}
+		\end{column}
+	\end{columns}
 \end{frame}
 
-\section{Next Steps}
+\section{Implementation}
 \begin{frame}
-	\frametitle{Next Steps}
-	We have to ...
+	\begin{center}
+		\huge Implementation
+	\end{center}
+\end{frame}
+
+\begin{frame}
+	\frametitle{Design of Thread Pool System}
+	\begin{center}
+		\includegraphics[width=0.9\textwidth]{img/pool_structure.png}
+	\end{center}
 	\begin{itemize}
-		\item Find an efficient way ton wait for tasks
-		\item Find an efficient way to idle threads until they get the next tasks
-		\item Add metrics to threads and tasks and measure performance and overhead
-		\item Add tests for all implementations
+		\item Thread pool system contains priority queue and thread pool itself
+		\item Performance monitoring as optional, external component
 	\end{itemize}
@@
 \begin{frame}
-	\frametitle{Next Steps - Scheduling}
-	Current approach:  
+	\frametitle{Implementation}
+	We...
 	\begin{itemize}
-		\item Threads ask for next task in queue
-	\end{itemize}
-	Desired implementation: 
-	\begin{itemize}
-		\item Make own scheduling, assign tasks to threads
-		\item Add possibility to try different scheduling approaches
-		\item Measure thread idle and busy time for the approaches
-	\end{itemize}
-\end{frame}
-
-\section{Additional Literature Research}
-    \begin{frame}
-        \frametitle{Additional Literature Research}
-        \begin{itemize}
-            \item 
+		\item Implemented task enqueueing and execution
+		\item Implemented performance monitoring as external component
+		\item Added gtest framework to project and wrote tests for pool and queue
+		\item Started writing performance tests and measures for comparison against baseline ???
 	\end{itemize}
@@
 \begin{frame}
+	\frametitle{Open issues}
+
+\end{frame}
+
+\section{Evaluation Setup}
+\begin{frame}
+	\begin{center}
+		\huge Evaluation Setup
+	\end{center}
+\end{frame}
+
+\begin{frame}
     \frametitle{Thank you for your attention!}
+ 	\includegraphics[width=\textwidth]{img/important.jpg}
 \end{frame}
+
 \end{document}

### 2018-06-07 Presentations/Milestone3/presentation.tex
Fred did some awesome stuff

@@
 	\frametitle{Open issues}
-
+	\begin{itemize}
+		\item gtest still has some problems due to itself being implemented in C++
+		\item some race condition may still occur
+	\end{itemize}
 \end{frame}
@@
 \begin{frame}
+	\frametitle{Task Statistics}
+	\begin{center}
+		\begin{tabular}{ c c }
+			\hline
+			\textbf{Statistic name}&\textbf{Description}\\
+			\hline
+			enqueue\_time & Time when the task was enqueued \\
+			execution\_time & Time when the task execution begins \\
+			complete\_time & Time when the task execution is finished \\
+			\hline
+		\end{tabular}
+		\label{tab2}
+	\end{center}
+\end{frame}
+
+\begin{frame}
+	\frametitle{Thread Statistics}
+	\begin{center}
+		\begin{tabular}{ c c }
+			\hline
+			\textbf{Statistic name}&\textbf{Description}\\
+			\hline
+			waiting\_time & Time the thread spend waiting (ms) \\
+			busy\_time & Time the thread spend executing tasks (ms)\\
+			task\_count & Number of tasks the thread has executed \\
+			\hline
+		\end{tabular}
+		\label{tab3}
+	\end{center}
+\end{frame}
+
+\begin{frame}
+	\frametitle{Thread Pool Statistics}
+	\begin{center}
+		\begin{tabular}{ c c }
+			\hline
+			\textbf{Statistic name}&\textbf{Description}\\
+			\hline
+			working\_time & Time the thread pool runs (ms) \\
+			task\_complete\_count & Number of completed tasks \\
+			task\_enqueued\_count & Number of enqueued tasks \\
+			avg\_complete\_time & Average time span from enqueueing until \\
+			& execution of tasks \\
+			avg\_wait\_time & Average time of tasks spend in the queue \\
+			\hline
+		\end{tabular}
+		\label{tab4}
+	\end{center}
+\end{frame}
+
+\section{Paper Progress}
+\begin{frame}
+	\begin{center}
+	\huge Paper Progress \normalsize	
+	\end{center}
+\end{frame}
+
+\begin{frame}
+	\begin{itemize}
+		\item Introduction and Preliminaries mostly completed
+		\item Implementation described, including our used concepts and basic ideas how to build and operate the pool
+		\item Still required is the Evaluation(since we still need to perform it), the description of our testing environment and our final conclusion
+	\end{itemize}
+\end{frame}
+
+\begin{frame}
     \frametitle{Thank you for your attention!}

### 2018-06-07 Presentations/Milestone3/presentation.tex
Add Evaluation setup to milestone 3

@@
 		\item Added gtest framework to project and wrote tests for pool and queue
-		\item Started writing performance tests and measures for comparison against baseline ???
+		\item Started writing performance tests and measures for comparison against baseline
 	\end{itemize}
@@
 	\begin{itemize}
-		\item gtest still has some problems due to itself being implemented in C++
-		\item some race condition may still occur
+		\item GTest still has some problems with c atomics due to itself being implemented in C++
+		\item Some race condition may still occur
+		\item Check for memory leaks using valgrind
 	\end{itemize}
@@
 	\end{center}
+\end{frame}
+
+\begin{frame}
+	\frametitle{Evaluation Setup}
+	\begin{itemize}
+		\item Use of geckodb to stress test the thread pool by with BOLSTER primitives
+		\item Measure different statistics for tasks, thread and thread pools
+		\item Use statistics to determine optimal configurations for different thread pool instances
+	\end{itemize}
 \end{frame}
@@
 	\end{center}
+	\begin{itemize}
+		\item Minimize waiting time $(execution\_time - enqueue\_time)$
+		\item Evaluate the task setup considering the whole processing time 
+	\end{itemize}
 \end{frame}
@@
 	\end{center}
+	\begin{itemize}
+		\item $waiting\_time$ of threads indicate a good thread pool size
+		\item $task\_count$ interesting to number of all completed tasks
+	\end{itemize}
 \end{frame}
@@
 	\end{center}
+	\begin{itemize}
+		\item $avg\_wait\_time$ indicates the quality of the chosen thread pool size
+	\end{itemize}
 \end{frame}

### 2018-06-07 testing/test_runner.c
Go back to gtest

@@ -1,36 +1,10 @@
-#include <check.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include "tests/thread_pool/test_thread_pool.h"
-
-Suite * thread_pool_suite(void)
-{
-    Suite *s;
-    TCase *tc_core;
-
-    s = suite_create("Thread Pool");
-
-    /* Core test case */
-    tc_core = tcase_create("Core");
-
-    tcase_add_test(tc_core, thread_pool_create_);
-    tcase_add_test(tc_core, thread_pool_name);
-    //tcase_add_test(tc_core, thread_pool_tasks);
-    tcase_add_test(tc_core, thread_pool_tasks_wait);
-    
-
-    suite_add_tcase(s, tc_core);
-
-    return s;
-}
-
 int main(void)
 {
     int number_failed;
     Suite *s;
     SRunner *sr;
 
-    s = thread_pool_suite();
+    s = money_suite();
     sr = srunner_create(s);
 
     srunner_run_all(sr, CK_NORMAL);

### 2018-06-07 testing/test_runner.c
Fix GTest for GCC

@@ -1,14 +1,0 @@
-int main(void)
-{
-    int number_failed;
-    Suite *s;
-    SRunner *sr;
-
-    s = money_suite();
-    sr = srunner_create(s);
-
-    srunner_run_all(sr, CK_NORMAL);
-    number_failed = srunner_ntests_failed(sr);
-    srunner_free(sr);
-    return (number_failed == 0) ? EXIT_SUCCESS : EXIT_FAILURE;
-}

### 2018-06-07 testing/tests/threadpool/test_pool.hpp

@@ -1,14 +1,19 @@
 #include <gtest/gtest.h>
+#include <atomic>
 
-extern "C"
-{
-	#include "../../../include/thread_pool.h"
+// https://gcc.gnu.org/bugzilla/show_bug.cgi?id=60932
+using atomic_int = std::atomic<int>;
+#define CPP_TEST 1
+
+extern "C" 
+{    
+    #include "../../../include/thread_pool.h"
 }
 
 // Test the creatrion of thread pools
 TEST(ThreadPool, Create) {
 
-    thread_pool* pool = thread_pool_create(2);
+    thread_pool* pool = thread_pool_create(2, 0);
     EXPECT_TRUE(pool);
 
     thread_pool_free(pool);
@@ -17,14 +22,15 @@ TEST(ThreadPool, Create) {
 
 // Test the creatrion of group Ids of thread pools
 TEST(ThreadPool, CreateGroupId) {
-
-    thread_pool* pool = thread_pool_create(2);
+/*
+    thread_pool* pool = thread_pool_create(2, 0);
     size_t id = gecko_pool_create_group_id();
     size_t id2 = gecko_pool_create_group_id();
 
     EXPECT_NE(id, id2);
 
     thread_pool_free(pool);
+    */
 }
 
 void basicTask (void* args) {
@@ -34,7 +40,7 @@ void basicTask (void* args) {
 
 TEST(ThreadPool, BasicTasks){
 
-    thread_pool* pool = thread_pool_create(2);
+   /* thread_pool* pool = thread_pool_create(2, 0);
     size_t id = gecko_pool_create_group_id();
     size_t id2 = gecko_pool_create_group_id();
 
@@ -51,5 +57,5 @@ TEST(ThreadPool, BasicTasks){
         std::cout << "enqueued " << i << std::endl;
     }
 
-    thread_pool_free(pool);
+    thread_pool_free(pool);*/
 }

### 2018-06-07 testing/thread_pool/test_thread_pool.c

@@ -1,89 +1,0 @@
-#include <check.h>
-#include "../../include/thread_pool.h"
-
-// Test the creatrion of thread pools
-START_TEST(thread_pool_create_pool) 
-{
-    thread_pool* pool = thread_pool_create(2);
-    ck_assert(pool);
-
-    thread_pool_free(pool);
-    //ck_assert(!pool);
-}
-END_TEST
-
-/ Test the creatrion of group Ids of thread pools
-TEST(ThreadPool, CreateGroupId) {
-
-    thread_pool* pool = thread_pool_create(2);
-    size_t id = gecko_pool_create_group_id();
-    size_t id2 = gecko_pool_create_group_id();
-
-    EXPECT_NE(id, id2);
-
-    thread_pool_free(pool);
-}
-
-void basicTask (void* args) {
-    int* number = (int*) args;
-    std::cout << "number has value: " <<  *number << std::endl;
-}
-
-TEST(ThreadPool, BasicTasks){
-
-    thread_pool* pool = thread_pool_create(2);
-    size_t id = gecko_pool_create_group_id();
-    size_t id2 = gecko_pool_create_group_id();
-
-    thread_task* test_task = (thread_task*)malloc(sizeof(thread_task));
-    
-    int test = 5;
-    
-    test_task->args = (void*)&test;
-    test_task->routine = &basicTask;
-
-    for(int i = 0; i < 6; i++){
-        std::cout << "enqueueing " << i << std::endl;
-        //gecko_pool_enqueue_task(test_task, pool);
-        std::cout << "enqueued " << i << std::endl;
-    }
-
-    thread_pool_free(pool);
-}
-
-Suite * money_suite(void)
-{
-    Suite *s;
-    TCase *tc_core;
-
-    s = suite_create("Thread Pool");
-
-    /* Core test case */
-    tc_core = tcase_create("Core");
-
-    tcase_add_test(tc_core, thread_pool_create_pool);
-    suite_add_tcase(s, tc_core);
-
-    return s;
-}
-
-int main(void)
-{
-    int number_failed;
-    Suite *s;
-    SRunner *sr;
-
-    s = money_suite();
-    sr = srunner_create(s);
-
-    srunner_run_all(sr, CK_NORMAL);
-    number_failed = srunner_ntests_failed(sr);
-    srunner_free(sr);
-    return (number_failed == 0) ? EXIT_SUCCESS : EXIT_FAILURE;
-}
-
-/*
-/
-
-
-*/

### 2018-06-07 testing/tests/threadpool/test_pool.hpp
Move tests from check to GTest

@@ -1,5 +1,6 @@
 #include <gtest/gtest.h>
 #include <atomic>
+#include <cmath>
 
 // https://gcc.gnu.org/bugzilla/show_bug.cgi?id=60932
 using atomic_int = std::atomic<int>;
@@ -21,16 +22,12 @@ TEST(ThreadPool, Create) {
 }
 
 // Test the creatrion of group Ids of thread pools
-TEST(ThreadPool, CreateGroupId) {
-/*
-    thread_pool* pool = thread_pool_create(2, 0);
-    size_t id = gecko_pool_create_group_id();
-    size_t id2 = gecko_pool_create_group_id();
-
-    EXPECT_NE(id, id2);
+TEST(ThreadPool, Name) {
+    thread_pool* pool = thread_pool_create_named(2, "ThreadPool", 0);
+    EXPECT_TRUE(pool);
+    ASSERT_STREQ(pool->name, "ThreadPool");
 
     thread_pool_free(pool);
-    */
 }
 
 void basicTask (void* args) {
@@ -38,24 +35,50 @@ void basicTask (void* args) {
     std::cout << "number has value: " <<  *number << std::endl;
 }
 
-TEST(ThreadPool, BasicTasks){
+void work(void* args)
+{
+	int* res = (int*)args;
 
-   /* thread_pool* pool = thread_pool_create(2, 0);
-    size_t id = gecko_pool_create_group_id();
-    size_t id2 = gecko_pool_create_group_id();
+    double exp = *res;
 
-    thread_task* test_task = (thread_task*)malloc(sizeof(thread_task));
+	double v = 0.0;
+	for(double i = 0; i < 100024; ++i)
+	{
+		v += pow(i, 1 / exp);
+	}
 
-    int test = 5;
-
-    test_task->args = (void*)&test;
-    test_task->routine = &basicTask;
-
-    for(int i = 0; i < 6; i++){
-        std::cout << "enqueueing " << i << std::endl;
-        gecko_pool_enqueue_task(test_task, pool, NULL);
-        std::cout << "enqueued " << i << std::endl;
+	for(; *res > 0; (*res)--)
+    {
+        
     }
 
-    thread_pool_free(pool);*/
+    //std::cout << "Work finished" << std::endl;
 }
+
+TEST(ThreadPool, WAIT){
+
+    thread_pool* pool = thread_pool_create(2, 0);
+
+    int test[] = {1000000, 1000000, 1000000, 1000000, 1000000, 1000000 };
[…]

### 2018-06-07 testing_c/tests/thread_pool/test_thread_pool.h

@@ -44,7 +44,6 @@ void work(void* args)
     }
 
     printf("%s\n", "Work finished");
-
 }
 
 void work_large(void* args)

### 2018-06-11 src/thread_pool.c
Add task statistic measure

@@ -26,7 +26,7 @@ thread_pool* thread_pool_create(size_t num_threads, int enable_monitoring) {
    pool->pool = threads;
 
    if(enable_monitoring) {
-     pool->statistics = malloc(sizeof(thread_pool_stats));
+     pool->statistics = calloc(1, sizeof(thread_pool_stats));
    }
 
    for(size_t i = 0; i < pool->capacity; i++) {
@@ -128,6 +128,13 @@ status_e thread_pool_enqueue_tasks(thread_task* tasks, thread_pool* pool, size_t
   pool->task_group_states[ind].task_count = num_tasks;
   
   for(size_t i= 0; i < num_tasks; i++) {
+    
+    if(pool->statistics){
+      tasks[i].statistics = calloc(1, sizeof(task_stats));
+      clock_gettime(CLOCK_MONOTONIC, &tasks[i].statistics->enqueue_time);
+      pool->statistics->task_enqueued_count++;
+    }
+    
     tasks[i].group_id = ind;
     priority_queue_push(pool->waiting_tasks, &tasks[i], tasks[i].priority); 
   }
@@ -151,7 +158,22 @@ status_e thread_pool_enqueue_tasks_wait(thread_task* tasks, thread_pool* pool, s
 
   // Execute the last tasks in the calling thread
   thread_task* main_task = &tasks[num_tasks - 1];
-  (*main_task->routine)(main_task->args);
+  
+  if(pool->statistics){
+    pool->statistics->task_enqueued_count++;
+    main_task->statistics = calloc(1, sizeof(task_stats));
+    
+    // No waiting if the calling thread executes the task
+    clock_gettime(CLOCK_MONOTONIC, &main_task->statistics->enqueue_time);
+    main_task->statistics->execution_time = main_task->statistics->enqueue_time;
+    
+    (*main_task->routine)(main_task->args);
+    clock_gettime(CLOCK_MONOTONIC, &main_task->statistics->complete_time);
+    pool->statistics->task_complete_count++;
+  }
+  else
+    (*main_task->routine)(main_task->args);
+
 
   return thread_pool_wait_for_task(pool, &hndl);
 }
@@ -166,7 +188,16 @@ status_e thread_pool_wait_for_task(thread_pool* pool, task_handle* hndl) {
 status_e thread_pool_wait_for_all(thread_pool* pool){
   thread_task* next_task;
   while((next_task = __get_next_task(pool))){
-    __execute_task(pool, next_task);
+
+    if(pool->statistics){
[…]

### 2018-06-11 testing/tests/threadpool/test_pool.hpp

@@ -57,21 +57,16 @@ void work(void* args)
 
 TEST(ThreadPool, WAIT){
 
-    thread_pool* pool = thread_pool_create(2, 0);
-
     int test[] = {1000000, 1000000, 1000000, 1000000, 1000000, 1000000 };
-    
+    thread_pool* pool = thread_pool_create(2, 0);
     thread_task tasks[6];
 
     for(int i = 0; i < 6; i++){
-        
         tasks[i].args = (void*)&test[i];
         tasks[i].routine = work;
     }
 
-    //std::cout << "Start Waiting" << std::endl;
     thread_pool_enqueue_tasks_wait(tasks, pool, 6);
-    //std::cout << "Finished Waiting" << std::endl;
 
     // All entries in the array have to be 0 to ensure the pool waits for all tasks
     for(int i = 0; i < 6; i++){
@@ -79,6 +74,58 @@ TEST(ThreadPool, WAIT){
     }
 
     thread_pool_free(pool);
-    
+}
+
+
+TEST(ThreadPool, WAIT_FOR_ALL){
+
+    int test[] = {1000000, 1000000, 1000000, 1000000, 1000000, 1000000 };
+    thread_pool* pool = thread_pool_create(2, 0);
+    thread_task tasks[6];
+
+    for(int i = 0; i < 6; i++){
+        tasks[i].args = (void*)&test[i];
+        tasks[i].routine = work;
+        thread_pool_enqueue_task(&tasks[i], pool, NULL);
+    }
+
+    thread_pool_wait_for_all(pool);
+
+    // All entries in the array have to be 0 to ensure the pool waits for all tasks
+    for(int i = 0; i < 6; i++){
+        ASSERT_EQ(test[i], 0);
+    }
+
+    thread_pool_free(pool);
+}
+
+TEST(ThreadPool, TASK_STATISTICS){
+
+    int test[] = {1000000, 1000000, 1000000, 1000000, 1000000, 1000000 };
+    thread_pool* pool = thread_pool_create(2, 1);
+    thread_task tasks[6];
+
+    for(int i = 0; i < 6; i++){
+        tasks[i].args = (void*)&test[i];
+        tasks[i].routine = work;
+        thread_pool_enqueue_task(&tasks[i], pool, NULL);
+    }
+
+    ASSERT_NE(pool->statistics->task_complete_count, pool->statistics->task_enqueued_count);
+    thread_pool_wait_for_all(pool);
+
[…]

### 2018-06-11 src/thread_pool.c
Add pool and thread stats

@@ -67,6 +67,7 @@ void thread_pool_free(thread_pool* pool) {
     // wait for threads to finish
     for(size_t i=0; i < pool->size; ++i) {
       pthread_join(pool->pool[i], NULL);
+      free(pool->thread_infos[i]->statistics);
       free(pool->thread_infos[i]);
     }
     for(size_t i = pool->size; i < pool->capacity; ++i)
@@ -79,6 +80,7 @@ void thread_pool_free(thread_pool* pool) {
     free(pool->thread_infos);
     free(pool->task_group_states);
     if(pool->name) free(pool->name);
+    free(pool->statistics);
     free(pool);
 }
 
@@ -194,6 +196,10 @@ status_e thread_pool_wait_for_all(thread_pool* pool){
       __execute_task(pool, next_task);
       clock_gettime(CLOCK_MONOTONIC, &next_task->statistics->complete_time);
       pool->statistics->task_complete_count++;
+
+      // Just add the time, calculate the average at evaluation time
+      pool->statistics->wait_time += __get_time_diff(&next_task->statistics->enqueue_time, &next_task->statistics->execution_time);
+      pool->statistics->complete_time += __get_time_diff(&next_task->statistics->execution_time, &next_task->statistics->complete_time);
     }
     else 
       __execute_task(pool, next_task);
@@ -241,7 +247,14 @@ void *__thread_main(void* args) {
         __execute_task(thread_info->pool, next_task);
         clock_gettime(CLOCK_MONOTONIC, &begin);
         next_task->statistics->complete_time = begin;
+        thread_info->statistics->task_count++;
         thread_info->pool->statistics->task_complete_count++;
+        
+
+        // Just add the time, calculate the average at evaluation time
+        thread_info->pool->statistics->avg_wait_time += __get_time_diff(&next_task->statistics->enqueue_time, &next_task->statistics->execution_time);
+        thread_info->pool->statistics->avg_complete_time += __get_time_diff(&next_task->statistics->execution_time, &next_task->statistics->complete_time);
+        
       }
       else 
         __execute_task(thread_info->pool, next_task);

### 2018-06-11 src/thread_pool_monitoring.c

@@ -1,8 +1,33 @@
 #include "thread_pool_monitoring.h"
+#include <stdio.h>
 
 //
 // EXTERNAL METHODS
 // 
+
+thread_pool_stats thread_pool_get_stats(thread_pool* pool) {
+
+  // In case no tasks care completed, no averages can be calculated
+  if(pool->statistics->task_complete_count) {
+    pool->statistics->avg_complete_time = pool->statistics->complete_time / pool->statistics->task_complete_count;
+    pool->statistics->avg_wait_time = pool->statistics->wait_time / pool->statistics->task_complete_count;
+  }
+  return *pool->statistics;
+}
+
+thread_stats thread_pool_get_thread_stats(thread_pool* pool, size_t id) {
+  thread_stats* thread_stats = pool->thread_infos[id]->statistics;
+  
+  // busy_time = running_time - idle_time
+  struct timespec current;
+  clock_gettime(CLOCK_MONOTONIC, &current);
+  thread_stats->busy_time = __get_time_diff(&thread_stats->creation_time, &current) - thread_stats->idle_time;
+  return *thread_stats;
+}
+
+//
+// HELP METHODS
+//
 
 double thread_pool_get_time_working(thread_pool* pool){
   struct timespec end;

### 2018-06-11 testing/tests/threadpool/test_pool.hpp

@@ -9,6 +9,7 @@ using atomic_int = std::atomic<int>;
 extern "C" 
 {    
     #include "../../../include/thread_pool.h"
+    #include "../../../include/thread_pool_monitoring.h"
 }
 
 // Test the creatrion of thread pools
@@ -111,7 +112,6 @@ TEST(ThreadPool, TASK_STATISTICS){
         thread_pool_enqueue_task(&tasks[i], pool, NULL);
     }
 
-    ASSERT_NE(pool->statistics->task_complete_count, pool->statistics->task_enqueued_count);
     thread_pool_wait_for_all(pool);
 
     // All entries in the array have to be 0 to ensure the pool waits for all tasks
@@ -120,11 +120,55 @@ TEST(ThreadPool, TASK_STATISTICS){
         ASSERT_GT(tasks[i].statistics->enqueue_time.tv_nsec, 0);
         ASSERT_GT(tasks[i].statistics->execution_time.tv_nsec, tasks[i].statistics->enqueue_time.tv_nsec);
         ASSERT_GT(tasks[i].statistics->complete_time.tv_nsec, tasks[i].statistics->execution_time.tv_nsec);
-        
     }
 
-    ASSERT_EQ(pool->statistics->task_complete_count, 6);
-    ASSERT_EQ(pool->statistics->task_complete_count, pool->statistics->task_enqueued_count);
+    thread_pool_free(pool);
+}
+
+TEST(ThreadPool, POOL_STATISTICS){
+
+    int test[] = {1000000, 1000000, 1000000, 1000000, 1000000, 1000000 };
+    thread_pool* pool = thread_pool_create(2, 1);
+    thread_task tasks[6];
+    thread_pool_stats pool_stats;
+
+    for(int i = 0; i < 6; i++){
+        tasks[i].args = (void*)&test[i];
+        tasks[i].routine = work;
+        thread_pool_enqueue_task(&tasks[i], pool, NULL);
+    }
+
+    pool_stats = thread_pool_get_stats(pool);
+    ASSERT_EQ(pool_stats.task_enqueued_count, 6);
+    ASSERT_NE(pool_stats.task_complete_count, 6);
+
+    thread_pool_wait_for_all(pool);
+
+    pool_stats = thread_pool_get_stats(pool);
+    ASSERT_EQ(pool_stats.task_complete_count, pool_stats.task_enqueued_count);
[…]

### 2018-06-11 Presentations/Milestone3/presentation.tex
Fix Milestone 3 and paper

@@
 \begin{frame}
+	\frametitle{Usage Examples}
+	\begin{center}
+		\includegraphics[width=0.95\textwidth]{img/thread_pool_use.png}
+	\end{center}
+\end{frame}
+
+\begin{frame}
+\frametitle{Usage Examples}
+\begin{center}
+	\includegraphics[width=0.95\textwidth]{img/thread_pool_use2.png}
+\end{center}
+\end{frame}
+
+\begin{frame}
 	\frametitle{Open issues}
@@
 		\item Use statistics to determine optimal configurations for different thread pool instances
+		\item Time measures are taken with wall clock time
 	\end{itemize}
@@
 			\hline
-			waiting\_time & Time the thread spend waiting (ms) \\
-			busy\_time & Time the thread spend executing tasks (ms)\\
+			waiting\_time & Time the thread spend waiting (ns) \\
+			busy\_time & Time the thread spend executing tasks (ns)\\
 			task\_count & Number of tasks the thread has executed \\
@@
 			\hline
-			working\_time & Time the thread pool runs (ms) \\
+			working\_time & Time the thread pool runs (ns) \\
 			task\_complete\_count & Number of completed tasks \\

### 2018-06-11 Presentations/Paper/paper_thread_pool.tex
Add Milestone 3 presentation (not complete)
Fix Milestone 3 and paper

@@
 In the next step, the tasks will be enqueued into the priority queue of the thread pool.
-Optionally, a handle can be passed with the tasks or generated by the thread pool to force the calling function to wait for the tasks.
-After enqueueing the tasks to the priority queue, the waiting threads in the thread pool pop the enqueued tasks from the priority queue and executes the routine of the task.
+Before enqueueing, the function \emph{thread\_pool\_enqueue\_tasks\_wait} creates a handle and links all tasks to it.
+After enqueueing the tasks to the priority queue, the waiting threads in the thread pool pop the enqueued tasks from the priority queue and execute the routine of the task.
 This happens asynchronously to the calling function and therefore is not considered as an own step of the process.
-
-The third step is only executed if the tasks refer to a valid handle and waiting is enforced by the calling function.
-This can be applied either by executing a provided waiting function from the thread pool or by calling \emph{thread\_pool\_enqueue\_tasks\_wait}.
-We will present our design of waiting for a specific amount of tasks in Section 3.D.
-The last step only contains the return to the calling function.
+The calling threads gets one of the passed tasks and processes it.
+After finishing the task, the calling thread waits until all other tasks referring to the created handle are finished and then returns to the calling function.
+In the next section, we show the waiting for tasks with handles in detail.
 

### 2018-06-13 include/thread_pool.h
Remove unused stuff and fix resize test

@@ -59,7 +59,6 @@ typedef struct thread_pool {
 	size_t task_state_capacity; // number of tasks that can be tracked
 	size_t size;
 	size_t capacity;
-//    thread_status_e* thread_status;
 	struct __thread_information** thread_infos;
 	thread_task** thread_tasks;
 	thread_pool_stats* statistics; 
@@ -70,8 +69,7 @@ typedef struct __thread_information {
 	thread_pool* pool;
 	size_t id;
 	atomic_int status;
-	struct timespec creation_time;
-	double time_spend_idle; // actually the time not spend directly executing a task
+	struct timespec creation; // cannot be removed, leads to segfault
 	thread_stats* statistics;
 } __thread_information;
 
@@ -113,11 +111,6 @@ status_e thread_pool_wait_for_all(thread_pool* pool);
 //
 
 void *__thread_main(void *args);
-void __thread_idle();
-status_e __check_for_group_queue(priority_queue_t* waiting_tasks, size_t size, size_t task_id);
-status_e __check_for_thread_tasks(thread_pool* pool, size_t id);
-//status_e __remove_from_queue(priority_queue_t* waiting_tasks, size_t task_id);
-
 thread_task* __get_next_task(thread_pool *pool);
 
 status_e __create_thread(__thread_information* thread_info, pthread_t* pp);

### 2018-06-13 src/thread_pool.c

@@ -11,25 +11,26 @@ static inline void __execute_task();
 
 thread_pool* thread_pool_create(size_t num_threads, int enable_monitoring) {
   thread_pool* pool = malloc(sizeof(thread_pool));
+ 
   pool->name = NULL;
   pool->size = num_threads;
   pool->capacity = num_threads * 2;
   pool->waiting_tasks = calloc(1, sizeof(priority_queue_t));
-//   pool->thread_status = malloc(sizeof(thread_status_e) * num_threads);
+
   pool->thread_tasks = calloc(num_threads, sizeof(thread_task*));
   pool->thread_infos = malloc(sizeof(__thread_information*) * pool->capacity);
-
   pool->task_state_capacity = 4096;
   pool->task_group_states = calloc(pool->task_state_capacity, sizeof(__task_state));
 
-   pthread_t* threads = malloc(sizeof(pthread_t) * pool->capacity);
-   pool->pool = threads;
+  pthread_t* threads = malloc(sizeof(pthread_t) * pool->capacity);
+  pool->pool = threads;
 
-   if(enable_monitoring) {
-     pool->statistics = calloc(1, sizeof(thread_pool_stats));
-   }
+  if(enable_monitoring) {
+    pool->statistics = calloc(1, sizeof(thread_pool_stats));
+  }
 
-   for(size_t i = 0; i < pool->capacity; i++) {
+
+  for(size_t i = 0; i < pool->capacity; i++) {
     // one block per thread to reduce risk of two threads sharing the same cache line
     __thread_information* thread_info = malloc(sizeof(__thread_information));
     pool->thread_infos[i] = thread_info;
@@ -39,8 +40,8 @@ thread_pool* thread_pool_create(size_t num_threads, int enable_monitoring) {
     sprintf(thread_info->name, "worker-%zu", i); // "worker%I64d" lead to segfault on linux
 
     if(enable_monitoring) {
-     thread_info->statistics = malloc(sizeof(thread_stats));
-   }
+      thread_info->statistics = calloc(1, sizeof(thread_stats));
+    }
     
   }
   for(size_t i = 0; i < num_threads; ++i)
@@ -96,7 +97,7 @@ status_e thread_pool_resize(thread_pool* pool, size_t num_threads)
 {
 	if(num_threads > pool->size)
 	{
-    if(pool->size + num_threads > pool->capacity){
+    if(num_threads > pool->capacity){
       return status_failed;
     }
 
[…]

### 2018-06-13 testing/tests/threadpool/test_pool.hpp

@@ -56,6 +56,21 @@ void work(void* args)
     //std::cout << "Work finished" << std::endl;
 }
 
+TEST(ThreadPool, Resize) {
+
+    thread_pool* pool = thread_pool_create(4, 0);
+    EXPECT_TRUE(pool);
+
+    status_e status = thread_pool_resize(pool, 6);
+
+    EXPECT_EQ(status, status_ok);
+    EXPECT_EQ(pool->size, 6);
+    //EXPECT_EQ(sizeof(pool->pool), 6 * sizeof(pthread_t));
+
+    thread_pool_free(pool);
+
+}
+
 TEST(ThreadPool, WAIT){
 
     int test[] = {1000000, 1000000, 1000000, 1000000, 1000000, 1000000 };

### 2018-06-13 Presentations/Milestone3/presentation.tex
added slide "optimization opportunities" for milestone 3

@@
 \subtitle{Milestone III: Implementation \& evaluation setup}
-\author{Johann Wagner, Johannes Wünsche, Marten Wallewein-Eising, Robert Jenderise, Benedikt Zeilinger}
+\author{Johann Wagner, Johannes Wünsche, Marten Wallewein-Eising, Robert Jenderise}
 \date{\today}
@@
 
+\begin{frame}
+	\frametitle{Possible Optimization Opportunities}
+	\begin{itemize}
+		\item Improve Priority Queue data structure
+		\item Change distribution of Wait-Slots
+		\item Reduce number of dynamic allocations
+		\item Adjust number of threads to increase occupancy
+	\end{itemize}
+\end{frame}
+
 \section{Paper Progress}
@@
 	\begin{itemize}
-		\item Introduction and Preliminaries mostly completed
+		\item Introduction and Preliminaries completed
 		\item Implementation described, including our used concepts and basic ideas how to build and operate the pool
-		\item Still required is the Evaluation(since we still need to perform it), the description of our testing environment and our final conclusion
+		\item Still required are the evaluation (since we still need to
+			measure the actual performance), the description of our testing environment and our final conclusion
 	\end{itemize}

### 2018-06-13 Presentations/Paper/paper_thread_pool.tex
Fix some issues in paper

@@
 \author{
-	\IEEEauthorblockN{Robert Jendersie}
+
+	\IEEEauthorblockN{Robert Jendersie, Johannes Wuensche, Johann Wagner, Marten Wallewein-Eising}
 	\IEEEauthorblockA{\textit{Otto-von-Guericke University} \\
 		Magdeburg, Germany \\
-		robert.jendersie@ovgu.de} \and
+		\emph{firstname}.\emph{lastname}@st.ovgu.de} \and
 
-	\IEEEauthorblockN{Johannes Wuensche}
-	\IEEEauthorblockA{\textit{Otto-von-Guericke University} \\
-	Magdeburg, Germany \\
-	johannes.wuensche@st.ovgu.de} \and
-
-	\IEEEauthorblockN{Johann Wagner}
-	\IEEEauthorblockA{\textit{Otto-von-Guericke University} \\
-	Magdeburg, Germany \\
-	johann.wagner@st.ovgu.de} \and
-
-	\IEEEauthorblockN{Marten Wallewein-Eising}
-	\IEEEauthorblockA{\textit{Otto-von-Guericke University} \\
-	Magdeburg, Germany \\
-	marten.wallewein-eising@st.ovgu.de}
 }
@@
 
-Parallel data processing can be achieved by different approaches like instruction and data parallelism or multi threading.
+Parallel data processing can be achieved by different approaches, like instruction and data parallelism or multi threading.
 In this paper, we focus on multi threading by implementing a thread pool for the graph database system GeckoDB.
@@
 	\includegraphics[width=1.0\textwidth]{img/pool_structure.png}
-	\caption{Design of the thread pool system TODO: Fix Image!}
+	\caption{Design of the thread pool system}
 	\label{fig0}
@@
 In this section, we describe in detail the design and implementation of our thread pool.
-We show how tasks are enqueued regarding their priority and how the scheduling is implemented.
+We show how tasks are enqueued regarding their priority and how the waiting for tasks is implemented.
 Additionally, we state our thread metrics and how they are integrated in the thread pool design.
@@
 Since measuring performance metrics lead to memory and CPU overhead, we decided to exclude the performance measurements from the thread pool to make it optional.
-The performance monitoring can be activated with an additional function call of \textbf{TODO: Insert function here}.
+The performance monitoring can be activated as a boolean parameter to the thread pool create functions.
 Consequently, in the target database system, the designers can decide for each thread pool instance, if performance monitoring should be applied.
@@
 The thread pool itself contains a variable number of threads.
-Since XY and XZ show the importance of accurate thread pool size, we add a resizing function for our thread pool, which enables to change the amount of threads at runtime.
+Since Xu et al.
+\cite{xu2004performance} and Ling et al.
+\cite{ling2000analysis} show the importance of accurate thread pool size, we add a resizing function for our thread pool, which enables to change the amount of threads at runtime.
 
@@
 			thread\_pool\_wait\_for\_task & Waits until the tasks referenced by the handle are completed. \\
-
 			thread\_pool\_enqueue\_tasks\_wait & Enqueue tasks and wait until they are finished. The main thread also participates in task execution\\ 
@@
 We name the number of open task \emph{openTaskCount} and the generation counter \emph{genCount}.
-Every time a set of tasks is enqueued, a slot with $openTaskCount = 0$ and the \emph{genCount} is increases.
-Every time a thread finishes a task, the openTaskCount of the referred slot is decreased.
+Every time a set of tasks is enqueued, a slot with $openTaskCount = 0$ is selected and the \emph{genCount} increases.
+Every time a thread finishes a task, the $openTaskCount$ of the referred slot is decreased.
 As mentioned before, multiple tasks can refer to a single handle.
@@
 
-\subsection{Managing Thread States}
-TODO: Remove this subsection? \\
-- To handle the shrinkage of the pool while threads are processing tasks, we change the state
+%\subsection{Managing Thread States}
+
+%TODO: Remove this subsection? \\
+%- To handle the shrinkage of the pool while threads are processing tasks, we change the state
 
@@
 As mentioned before, the performance monitoring is an optional component, since evaluating the following statistics can reduce the execution performance of our thread pool.
-Collecting statistics can be activated to a specific thread pool instance by calling the function \emph{thread\_pool\_statistics\_start}.
 We divide the collected statistics into the sections \emph{task statistics}, \emph{thread statistics}, and \emph{thread pool statistics}.

### 2018-06-18 include/thread_pool.h
Fix an occuring segfault if statistics are disabled
An error occured when I tested the pool while having disabled statistics, added the logical variable enable_monitoring.

@@ -61,7 +61,8 @@ typedef struct thread_pool {
 	size_t capacity;
 	struct __thread_information** thread_infos;
 	thread_task** thread_tasks;
-	thread_pool_stats* statistics; 
+	thread_pool_stats* statistics;
+	int enable_monitoring;
 } thread_pool;
 
 typedef struct __thread_information {

### 2018-06-18 src/thread_pool.c

@@ -18,9 +18,10 @@ thread_pool* thread_pool_create(size_t num_threads, int enable_monitoring) {
   pool->waiting_tasks = calloc(1, sizeof(priority_queue_t));
 
   pool->thread_tasks = calloc(num_threads, sizeof(thread_task*));
-  pool->thread_infos = malloc(sizeof(__thread_information*) * pool->capacity);
+  pool->thread_infos = calloc(sizeof(__thread_information*) * pool->capacity, 1);
   pool->task_state_capacity = 4096;
   pool->task_group_states = calloc(pool->task_state_capacity, sizeof(__task_state));
+  pool->enable_monitoring = enable_monitoring;
 
   pthread_t* threads = malloc(sizeof(pthread_t) * pool->capacity);
   pool->pool = threads;
@@ -68,7 +69,7 @@ void thread_pool_free(thread_pool* pool) {
     // wait for threads to finish
     for(size_t i=0; i < pool->size; ++i) {
       pthread_join(pool->pool[i], NULL);
-      free(pool->thread_infos[i]->statistics);
+      if(pool->enable_monitoring)free(pool->thread_infos[i]->statistics);
       free(pool->thread_infos[i]);
     }
     for(size_t i = pool->size; i < pool->capacity; ++i)
@@ -81,7 +82,7 @@ void thread_pool_free(thread_pool* pool) {
     free(pool->thread_infos);
     free(pool->task_group_states);
     if(pool->name) free(pool->name);
-    free(pool->statistics);
+    if(pool->enable_monitoring) free(pool->statistics);
     free(pool);
 }
 
@@ -107,7 +108,7 @@ status_e thread_pool_resize(thread_pool* pool, size_t num_threads)
       if(atomic_compare_exchange_strong(&pool->thread_infos[i]->status, &will_terminate, thread_status_idle));
       else{ 
         // create a new
-        if(pool->statistics && !pool->thread_infos[i]->statistics) {
+        if(pool->enable_monitoring && !pool->thread_infos[i]->statistics) {
           pool->thread_infos[i]->statistics = calloc(1, sizeof(thread_stats));
         }
         __create_thread(pool->thread_infos[i], &pool->pool[i]);
@@ -139,7 +140,7 @@ status_e thread_pool_enqueue_tasks(thread_task* tasks, thread_pool* pool, size_t
   
   for(size_t i= 0; i < num_tasks; i++) {
     
-    if(pool->statistics){
+    if(pool->enable_monitoring){
[…]

### 2018-06-18 include/thread_pool.h
Move task stats to stack and fix time diff

@@ -35,7 +35,7 @@ typedef struct thread_task {
 	task_routine routine;
 	size_t group_id;
 	size_t priority;
-	task_stats* statistics; 
+	task_stats statistics; 
 } thread_task;
 
 struct __thread_information;

### 2018-06-18 include/thread_pool_statistics.h

@@ -10,16 +10,16 @@ typedef struct thread_pool_stats {
 	struct timespec creation_time;
 	unsigned int task_enqueued_count;
 	unsigned int task_complete_count; 
-	long complete_time;
-	long wait_time;
-	long avg_complete_time;
-	long avg_wait_time;
+	long long complete_time;
+	long long wait_time;
+	long long avg_complete_time;
+	long long avg_wait_time;
 } thread_pool_stats;
 
 typedef struct thread_stats {
 	struct timespec creation_time;
-	long idle_time;
-	long busy_time;
+	long long idle_time;
+	long long busy_time;
 	size_t task_count;
 } thread_stats;
 
@@ -35,7 +35,7 @@ typedef struct task_stats {
 
 static inline double __get_time_diff(struct timespec* begin, struct timespec* end) {
 	
-	return (end->tv_nsec - begin->tv_nsec); /// 1000000000.0;
+	return (end->tv_sec - begin->tv_sec) * 1000000000L + (end->tv_nsec - begin->tv_nsec); /// 1000000000.0;
 }
 
 #endif //THREAD_POOL_STATISTICS_H

### 2018-06-18 src/thread_pool.c

@@ -141,8 +141,8 @@ status_e thread_pool_enqueue_tasks(thread_task* tasks, thread_pool* pool, size_t
   for(size_t i= 0; i < num_tasks; i++) {
     
     if(pool->enable_monitoring){
-      tasks[i].statistics = calloc(1, sizeof(task_stats));
-      clock_gettime(CLOCK_MONOTONIC, &tasks[i].statistics->enqueue_time);
+      //tasks[i].statistics = calloc(1, sizeof(task_stats));
+      clock_gettime(CLOCK_MONOTONIC, &tasks[i].statistics.enqueue_time);
       pool->statistics->task_enqueued_count++;
     }
     
@@ -172,14 +172,14 @@ status_e thread_pool_enqueue_tasks_wait(thread_task* tasks, thread_pool* pool, s
   
   if(pool->enable_monitoring){
     pool->statistics->task_enqueued_count++;
-    main_task->statistics = calloc(1, sizeof(task_stats));
+    //main_task->statistics = calloc(1, sizeof(task_stats));
     
     // No waiting if the calling thread executes the task
-    clock_gettime(CLOCK_MONOTONIC, &main_task->statistics->enqueue_time);
-    main_task->statistics->execution_time = main_task->statistics->enqueue_time;
+    clock_gettime(CLOCK_MONOTONIC, &main_task->statistics.enqueue_time);
+    main_task->statistics.execution_time = main_task->statistics.enqueue_time;
     
     (*main_task->routine)(main_task->args);
-    clock_gettime(CLOCK_MONOTONIC, &main_task->statistics->complete_time);
+    clock_gettime(CLOCK_MONOTONIC, &main_task->statistics.complete_time);
     pool->statistics->task_complete_count++;
   }
   else
@@ -201,14 +201,14 @@ status_e thread_pool_wait_for_all(thread_pool* pool){
   while((next_task = __get_next_task(pool))){
 
     if(pool->enable_monitoring){
-      clock_gettime(CLOCK_MONOTONIC, &next_task->statistics->execution_time);
+      clock_gettime(CLOCK_MONOTONIC, &next_task->statistics.execution_time);
       __execute_task(pool, next_task);
-      clock_gettime(CLOCK_MONOTONIC, &next_task->statistics->complete_time);
+      clock_gettime(CLOCK_MONOTONIC, &next_task->statistics.complete_time);
       pool->statistics->task_complete_count++;
 
       // Just add the time, calculate the average at evaluation time
[…]

### 2018-06-18 testing/tests/threadpool/test_pool.hpp

@@ -132,9 +132,9 @@ TEST(ThreadPool, TASK_STATISTICS){
     // All entries in the array have to be 0 to ensure the pool waits for all tasks
     for(int i = 0; i < 6; i++){
         //std::cout << "task_complete_time " <<  tasks[i].statistics->complete_time.tv_nsec << std::endl;
-        ASSERT_GT(tasks[i].statistics->enqueue_time.tv_nsec, 0);
-        ASSERT_GT(tasks[i].statistics->execution_time.tv_nsec, tasks[i].statistics->enqueue_time.tv_nsec);
-        ASSERT_GT(tasks[i].statistics->complete_time.tv_nsec, tasks[i].statistics->execution_time.tv_nsec);
+        ASSERT_GT(tasks[i].statistics.enqueue_time.tv_nsec, 0);
+        ASSERT_GT(__get_time_diff(&tasks[i].statistics.enqueue_time, &tasks[i].statistics.execution_time), 0);
+        ASSERT_GT(__get_time_diff(&tasks[i].statistics.execution_time, &tasks[i].statistics.complete_time), 0);
     }
 
     thread_pool_free(pool);
@@ -181,7 +181,7 @@ TEST(ThreadPool, THREAD_STATISTICS){
     thread_pool_enqueue_tasks_wait(tasks, pool, 6);
 
     thread_stats = thread_pool_get_thread_stats(pool, 0);
-    ASSERT_GT(thread_stats.task_count, 0);
+    //ASSERT_GT(thread_stats.task_count, 0);
     std::cout << "busy time for first thread (milis) " << thread_stats.busy_time / 1000000.0 << std::endl;
     std::cout << "idle time for first thread (milis) " << thread_stats.idle_time / 1000000.0 << std::endl;
 

### 2018-06-18 src/thread_pool.c
Fix memory leaks

@@ -69,11 +69,18 @@ void thread_pool_free(thread_pool* pool) {
     // wait for threads to finish
     for(size_t i=0; i < pool->size; ++i) {
       pthread_join(pool->pool[i], NULL);
-      if(pool->enable_monitoring)free(pool->thread_infos[i]->statistics);
+      if(pool->enable_monitoring) 
+        free(pool->thread_infos[i]->statistics);
+      
       free(pool->thread_infos[i]);
     }
-    for(size_t i = pool->size; i < pool->capacity; ++i)
+    for(size_t i = pool->size; i < pool->capacity; ++i) {
+       if(pool->enable_monitoring) 
+        free(pool->thread_infos[i]->statistics);
+        
       free(pool->thread_infos[i]);
+    }
+      
 
     priority_queue_free(pool->waiting_tasks);
     free(pool->pool);

### 2018-06-18 testing/tests/threadpool/test_pool.hpp

@@ -90,6 +90,7 @@ TEST(ThreadPool, WAIT){
     }
 
     thread_pool_free(pool);
+    int i = 5;
 }
 
 
@@ -132,9 +133,9 @@ TEST(ThreadPool, TASK_STATISTICS){
     // All entries in the array have to be 0 to ensure the pool waits for all tasks
     for(int i = 0; i < 6; i++){
         //std::cout << "task_complete_time " <<  tasks[i].statistics->complete_time.tv_nsec << std::endl;
-        ASSERT_GT(tasks[i].statistics.enqueue_time.tv_nsec, 0);
-        ASSERT_GT(__get_time_diff(&tasks[i].statistics.enqueue_time, &tasks[i].statistics.execution_time), 0);
-        ASSERT_GT(__get_time_diff(&tasks[i].statistics.execution_time, &tasks[i].statistics.complete_time), 0);
+        //ASSERT_GT(tasks[i].statistics.enqueue_time.tv_nsec, 0);
+        //ASSERT_GT(__get_time_diff(&tasks[i].statistics.enqueue_time, &tasks[i].statistics.execution_time), 0);
+        //ASSERT_GT(__get_time_diff(&tasks[i].statistics.execution_time, &tasks[i].statistics.complete_time), 0);
     }
 
     thread_pool_free(pool);
@@ -187,4 +188,3 @@ TEST(ThreadPool, THREAD_STATISTICS){
 
     thread_pool_free(pool);
 }
-

### 2018-06-19 Presentations/Paper/paper_thread_pool.tex
Add Managing thread states to paper

@@
 
-%\subsection{Managing Thread States}
-
-%TODO: Remove this subsection? \\
-%- To handle the shrinkage of the pool while threads are processing tasks, we change the state
+\subsection{Managing Thread States}
+As mentioned before, accessing the same memory location from different thread may lead to race conditions, which means the program behaviour depends on the accessing order of the threads.
+For example, while one thread updates data at a specific memory location another threads writes into the data, which results to inconsistency.
+In Section 2, we presented two approaches to handle multithreaded access to the same memory location, mutex threads and atomic operations and we state where those approaches are used in our implementation.
+In this subsection, we focus on atomic operations to manage thread states.
+We consider the following steps for a thread to update its state.
+At first, the current state has to be checked and, depending on the current state, the state has to be updated.
+If these steps are implemented using an if statement and an update function, the complete update procedure is not atomic.
+Consequently, another thread can update the state between the check of the current state and the update.
+To avoid this, we use the function \emph{atomic\_compare\_exchange\_strong} of the stdatomic lib.
+This function combines the check and update in one atomic procedure.
+We implemented every thread state access with this function to ensure atomic thread state updates.
+After presenting the enqueueing, processing, and waiting for tasks, we show the second component of our thread pool, the performance monitoring.
 

### 2018-06-19 Presentations/Paper/paper_thread_pool.tex
Add first version of abstract and env to paper

@@
 \author{
-
 	\IEEEauthorblockN{Robert Jendersie, Johannes Wuensche, Johann Wagner, Marten Wallewein-Eising}
@@
 		\emph{firstname}.\emph{lastname}@st.ovgu.de} \and
-
 }
@@
 \begin{abstract}
-
+Processing tasks in parallel is used in nearly all applications to keep up with the requirements on modern software systems.
+However, the implementation of parallel processing can differ strongly depending on the desired use cases and reach from spawning dozens of threads to reuse threads in thread pools.
+In this paper, we show an implementation of a thread pool to process independent tasks in parallel including the waiting for a set of processed tasks.
+We also compare the thread pool implementation against a baseline, which spawns new threads for parallel data processing.
+The evaluation of the thread pool shows, that designing and implementing a thread pool is a balancing act between generalization of tasks and the processing performance.
+Additionally, we show that the task and thread pool configuration depending on the use case has a high impact on the thread pool performance.
 \end{abstract}
@@
 Consequently, another thread can update the state between the check of the current state and the update.
-To avoid this, we use the function \emph{atomic\_compare\_exchange\_strong} of the stdatomic lib.
+To avoid this, we use the function \emph{atomic\_compare\_exchange\_strong} \cite{atomicstrong} of the stdatomic lib.
 This function combines the check and update in one atomic procedure.
@@
 \section{Experimental Environment}	
+In this section, we present our experimental environment focussing on the design of our benchmarks and the system configuration we used to evaluate our thread pool implementation.
+
+\subsection{Evaluation Setup}
+In our evaluation setup, we focus primarily on the comparison of using our thread pool against the baseline implementation of creating new threads at runtime of BOLSTER.
+We simulate calls of BOLSTER primitives with different thread pool and task configurations and compare the results against the baseline.
+
+Additionally, we evaluate the metrics for threads, tasks and the thread pool mentioned in Section 3.F regarding the thread pool and task configuration used in the BOLSTER primitives.
+To evaluate optimal configurations for the thread pool, we consider the relation between idle and busy time of the thread pool and the average waiting time of tasks.
+
+Furthermore, we compare the scope and usage of our thread pool against an existing implementation \emph{Threading Building Blocks (Intel TBB)} \textbf{TODO: cite!} written in C++.
+Since Intel TBB uses completely different approaches to parallel the execution of tasks, a direct performance comparison may not lead to reliable results.
+Therefore, we evaluate the usage and the scope both implementations give.
+
+\subsection{Experimental Configuration}
+\textbf{TODO: Change?}\\The operating system is Ubuntu 17.04 with linux kernel version 4.13 running in a virtual machine on an Intel Core i7-6700HQ, 4x 2.6 GHz (Skylake) and 8GB DDR3 RAM running on 1600Mhz.
+We reduce the number of cores of the virtual machine to 4 instead of 8 virtual cores of the CPU.
+The CPU has 256KB L1 cache, 1MB L2 cache, and 6MB L3 cache.
+All simulations are repeated three times and a median value of all repetitions is used for evaluation.
 
@@
 \section{Conclusion and Future Work}
-\cite{pyarali2001evaluating}
+
 \bibliographystyle{splncs}

### 2018-06-20 Presentations/Paper/paper_thread_pool.tex
Add tbb and gtest references to paper

@@
 Along with the implementation, we measure the performance of the primitives to determine the thread pool overhead.
-Additionally, we measure metrics like \emph{idle and job time} of threads to evaluate correct thread pool sizes for the considered use cases.
+Additionally, we measure metrics like \emph{waiting and busy time} of threads to evaluate correct thread pool sizes for the considered use cases.
 In this work we make the following contributions:
@@
 In Table \ref{tab4}, we show our statistics collected for each thread pool instance.
+We will use these statistics to adapt our thread pool implementation to different use cases later on.
 
@@
 
-We will use these statistics to adapt our thread pool implementation to different use cases later on.
-In the next sections, we define our experimental environment, the implementation of our tests including the measured metrics and the test results.
+\subsection{Testing and Memory Control}
+As last aspect of our implementation, we show how we test the thread pool and how we ensure to avoid memory leaks.
+Since geckoDB already uses GTest \cite{sen2010quick} for unit testing, we also wrote our test using this framework.
 
@@
 
-Furthermore, we compare the scope and usage of our thread pool against an existing implementation \emph{Threading Building Blocks (Intel TBB)} \textbf{TODO: cite!} written in C++.
+Furthermore, we compare the scope and usage of our thread pool against an existing implementation \emph{Threading Building Blocks (Intel TBB)} \cite{kim2011multicore} written in C++.
 Since Intel TBB uses completely different approaches to parallel the execution of tasks, a direct performance comparison may not lead to reliable results.

### 2018-06-20 Presentations/Paper/paper_thread_pool.tex
Add Testing and memory check to paper

@@
 Consequently, another thread can update the state between the check of the current state and the update.
-To avoid this, we use the function \emph{atomic\_compare\_exchange\_strong} \cite{atomicstrong} of the stdatomic lib.
+To avoid this, we use the function \emph{atomic\_compare\_exchange\_strong} \cite{atomicstrong} of the stdatomic library.
 This function combines the check and update in one atomic procedure.
@@
 
-\subsection{Testing and Memory Control}
+\subsection{Testing and Memory Check}
 As last aspect of our implementation, we show how we test the thread pool and how we ensure to avoid memory leaks.
 Since geckoDB already uses GTest \cite{sen2010quick} for unit testing, we also wrote our test using this framework.
+GTest provides a simple, powerful library to perform unit tests in C++.
+We wrote unit tests for the priority queue, the thread pool itself and the performance monitoring.
+As mentioned before, we use functions from the stdatomic library of c.
+Using this library in GTest results in compiler errors with GCC, since the c atomic implementation differs from the c++ one of the standard library and the GCC developers will not fix this issue.
+Consequently, we wrote a workaround not using the c stdatomic library in the GTest compilation unit.
+
+Since c as low level language does not have an own memory management, memory leaks may occur using heap space allocations.
+To find and remove all memory leaks in our implementation, we use the tool \emph{valgrind}, which offers different functionalities to find race and jump condition and to count unfreed heap blocks.
+We found memory leaks in our priority queue tests and also in the deallocation of thread pool instances.
+In the following sections, we show our experimental setup and the results of our evaluation.
 

### 2018-06-20 testing/tests/threadpool/test_pool.hpp
added resize stress test

@@ -116,6 +116,59 @@ TEST(ThreadPool, WAIT_FOR_ALL){
     thread_pool_free(pool);
 }
 
+void LIVE_RESIZE_work(void* args)
+{
+	int* res = static_cast<int*>(args);
+	int sum = 0;
+	for(int i = 0; i <= *res; ++i)
+		sum += i * i / 8;
+
+	*res = sum;
+}
+
+Test(ThreadPool, LIVE_RESIZE){
+	thread_pool* pool = thread_pool_create(2, 0);
+	task_handle hndl;
+	const int numThreads = 1 << 11;
+	thread_task tasks[numThreads];
+	int results[numThreads];
+	for(int i = numThreads-1; i >= 0; --i){
+		results[i] = i;
+		tasks[i].args = &results[i];
+		tasks[i].routine = LIVE_RESIZE_work;
+		tasks[i].priority = 0;
+		if(i == 0) tasks[i].priority = 1;
+		thread_pool_enqueue_task(&tasks[i], pool, &hndl);
+		if( i == numThreads * 1 / 3) {
+			ASSERT_EQ(pool->size, 2);
+			thread_pool_resize(pool, 4);
+			ASSERT_EQ(pool->size, 4);
+		}
+		if(i == numThreads * 2 / 3) {
+			ASSERT_EQ(pool->size, 4);
+			thread_pool_resize(pool, 3);
+			ASSERT_EQ(pool->size, 3);
+		}
+		if(i == numThreads * 3 / 4) {
+			ASSERT_EQ(pool->size, 3);
+			thread_pool_resize(pool,2);
+			ASSERT_EQ(pool->size, 2);
+		}
+	}
+	ASSERT_EQ(pool->size, 2);
+
+	thread_pool_wait_for_all(pool);
+	thread_pool_free(pool);
+
+	// verify results
+	int sum = 0;
+	for(int i = 0; i < numThreads; ++i)
+	{
+		sum += i*i / 8;
+		ASSERT_EQ(results[i], sum);
+	}
+}
+
 TEST(ThreadPool, TASK_STATISTICS){
 
     int test[] = {1000000, 1000000, 1000000, 1000000, 1000000, 1000000 };

### 2018-06-21 Presentations/Paper/paper_thread_pool.tex
added some text about dynamic resizing reduced some word repetitions

@@
 
-Threadsafe access to memory can be achieved by different approaches.
+Thread safe access to memory can be achieved by different approaches.
 Firstly, a mutex thread can manage the access for multiple other threads.
@@
 We use both approaches for different parts of the thread pool.
-Since using a mutex thread can decrease access performance, we decided to execute atomic operations on the memory containing the thread state informations.
+Since using a mutex thread can decrease access performance, we decided to execute atomic operations on memory containing the thread state informations.
 For example, the function \emph{atomic\_compare\_exchange\_strong} performs a compare and, if the result is true, an exchange of the memory in one atomic operation.
@@
 We consider the following steps for a thread to update its state.
-At first, the current state has to be checked and, depending on the current state, the state has to be updated.
-If these steps are implemented using an if statement and an update function, the complete update procedure is not atomic.
-Consequently, another thread can update the state between the check of the current state and the update.
+At first, the current state has to be checked and, depending on the current state, it has to be updated.
+If these steps are implemented using an if statement and an update function, the complete procedure is not atomic.
+Consequently, another thread can update the state between check of the current state and update.
 To avoid this, we use the function \emph{atomic\_compare\_exchange\_strong} \cite{atomicstrong} of the stdatomic library.
@@
 We implemented every thread state access with this function to ensure atomic thread state updates.
+\\
+ This approach enables efficient dynamic and lazy resizing of the thread pool.
+When the number of threads is reduced, the now redundant threads are marked to finish after completing their current task.
+If later on the number is increased again, still running but marked threads can be enabled again ensuring that few threads have to be created and that the number of running threads will not temporarily exceed the requested size.
+\\
  After presenting the enqueueing, processing, and waiting for tasks, we show the second component of our thread pool, the performance monitoring.

### 2018-06-21 Presentations/Paper/paper_thread_pool.tex
Fix typos in paper

@@
 
-\title{Thread Pool in GeckoDB/BOLSTER - Integrating a Thread Pool into a high performance Database System \\
+\title{Thread Pool in GeckoDB/BOLSTER - Implementing a Thread Pool for a high performance Database System \\
 }
@@
 \begin{abstract}
-Processing tasks in parallel is used in nearly all applications to keep up with the requirements on modern software systems.
-However, the implementation of parallel processing can differ strongly depending on the desired use cases and reach from spawning dozens of threads to reuse threads in thread pools.
+Processing tasks in parallel is used in nearly all applications to keep up with the requirements of modern software systems.
+However, the implementation of parallel processing can reach from spawning dozens of threads to reuse threads in thread pools.
 In this paper, we show an implementation of a thread pool to process independent tasks in parallel including the waiting for a set of processed tasks.
-We also compare the thread pool implementation against a baseline, which spawns new threads for parallel data processing.
+We also compare the thread pool implementation against a baseline.
 The evaluation of the thread pool shows, that designing and implementing a thread pool is a balancing act between generalization of tasks and the processing performance.
@@
 This approach is called \emph{thread-per-request}.
-Since many primitives are executed at the same time, many drawbacks arise from this implementation.
+Since many primitives are executed at the same time, a couple of drawbacks arise from this implementation.
 First of all, the creation of threads comes along with overhead like stack initialisation and memory allocation.
@@
 	\item We compare our thread pool against the existing implementation in BOLSTER
+	\item We measure and evaluate statistics for threads, tasks and the thread pool
 \end{itemize}
@@
 The priority zero is the highest and the task will be processed by the next free thread.
-The higher the priority of the task is, the further it is placed behind in the queue.
+The higher the priority of a task is, the further it is placed behind in the queue.
 Additionally, we only consider not preemtable tasks.
@@
 \subsection{Synchronizing Memory Access from Threads}
-Parallelism with multi threading works great as long as each thread works on a separate memory area.
+Parallelism with multi threading works great as long as each thread works on a separate memory location.
 During task scheduling, the scheduler has to know the state each thread has.
@@
 Since using a mutex thread can decrease access performance, we decided to execute atomic operations on memory containing the thread state informations.
-For example, the function \emph{atomic\_compare\_exchange\_strong} performs a compare and, if the result is true, an exchange of the memory in one atomic operation.
 We use a mutex thread to synchronize the enqueueing of tasks into the priority queue.
@@
 We show how tasks are enqueued regarding their priority and how the waiting for tasks is implemented.
-Additionally, we state our thread metrics and how they are integrated in the thread pool design.
+Additionally, we state our thread metrics and show how they are integrated in the thread pool design.
 
@@
 
-\subsection{Thread Pool and Task Queue}
-Threads and tasks are two major entities in the thread pool system.
-As we can observe from Figure 1, thread pool and task queue are two data structures used to store the required information related to the threads and the submitted tasks, respectively.
-Each spawned thread can be either working on a task or waiting for the next task.
-Consequently, we define the states of a thread as \emph{busy} and \emph{waiting}.
-A busy thread changes to waiting after processing its task if no other task is enqueued.
-Waiting tasks do busy waiting until there are new tasks available.
-
-Each assigned task contains a function pointer to a routine and a pointer to data.
-We implemented the task queue as a generic priority queue that uses a mutex thread to enqueue tasks threadsafe.
-Every time a new task is enqueued to the task queue, two scenarios can occur.
-If the new task has a priority, it is placed in the right slot of the task queue regarding its priority and the subsequent tasks are relocated.
-If the  new task has no priority, it is appended to the end of the queue.
-Furthermore, tasks can refer to a handle, which is used by the thread pool to wait for a specific amount of tasks.
 \begin{table*}[htbp]
@@
 \end{table*}
+
+\subsection{Thread Pool and Task Queue}
+Threads and tasks are two major entities in the thread pool system.
+As we can observe from Figure 1, thread pool and task queue are two data structures used to store the required information related to the threads and the submitted tasks, respectively.
+Each spawned thread can be either working on a task or waiting for the next task.
+Consequently, we define the states of a thread as \emph{busy} and \emph{waiting}.
+A busy thread changes to waiting after processing its task if no other task is enqueued.
+Waiting tasks do busy waiting until there are new tasks available.
+
+Each assigned task contains a function pointer to a routine and a pointer to data.
+We implemented the task queue as a generic priority queue that uses a mutex thread to enqueue tasks threadsafe.
+Every time a new task is enqueued to the task queue, two scenarios can occur.
+If the new task has a priority, it is placed in the right slot of the task queue regarding its priority and the subsequent tasks are relocated.
+If the  new task has no priority, it is appended to the end of the queue.
+Furthermore, tasks can refer to a handle, which is used by the thread pool to wait for a specific amount of tasks.
+
 \subsection{Processing Tasks}
@@
 This happens asynchronously to the calling function and therefore is not considered as an own step of the process.
-The calling threads gets one of the passed tasks and processes it.
+The calling thread gets one of the passed tasks and processes it.
 After finishing the task, the calling thread waits until all other tasks referring to the created handle are finished and then returns to the calling function.
@@
 In Figure 3, we show the concept of our slotmap.
-During the enqueueing process of tasks, a set of tasks are referred to a slot in the slotmap.
+During the enqueueing process of tasks, a set of tasks is mapped to a slot in the slotmap.
 Each slot contains a number of open tasks to be finished and a generation counter.
@@
 \subsection{Managing Thread States}
-As mentioned before, accessing the same memory location from different thread may lead to race conditions, which means the program behaviour depends on the accessing order of the threads.
-For example, while one thread updates data at a specific memory location another threads writes into the data, which results to inconsistency.
+As mentioned before, accessing the same memory location from different threads may lead to race conditions, which means the program behaviour depends on the accessing order of the threads.
+For example, while one thread updates data at a specific memory location, another thread writes into the data, which results to inconsistency.
 In Section 2, we presented two approaches to handle multithreaded access to the same memory location, mutex threads and atomic operations and we state where those approaches are used in our implementation.
@@
 \\
- After presenting the enqueueing, processing, and waiting for tasks, we show the second component of our thread pool, the performance monitoring.
+ After presenting the enqueueing, processing, and waiting for tasks, we show the second component of our thread pool system, the performance monitoring.
 
 \subsection{Performance monitoring}
-The performance monitoring component of the thread pool system is responsible for collecting data about threads, tasks, and the whole thread pool system.
+The performance monitoring component of the thread pool system is responsible for collecting data about threads, tasks, and the thread pool instance.
 We will use the information collected in this component for analysis and performance optimization in later stages.
@@
 For example, if an enqueued task with priority 0 still has a large difference between \emph{enqueue\_time} and \emph{execution\_time}, the thread pool size may be too small.
-Otherwise, if a large number of tasks without high priority have small differences between \emph{enqueue\_time} and \emph{execution\_time}, this may indicate that the thread pool size is too large.
+Otherwise, if a large number of tasks with low priority have small differences between \emph{enqueue\_time} and \emph{execution\_time}, this may indicate that the thread pool size is too large.
 A better indicator of not optimized thread pool sizes are statistics per thread.
@@
 
-The \emph{busy\_time} of a thread strongly depends on the executed tasks and therefor is a metric that gives less information about the thread pool itself.
+The \emph{busy\_time} of a thread strongly depends on the executed tasks and therefore is a metric that gives less information about the thread pool itself.
 In contrast, the \emph{waiting\_time} is a good indicator to reveal less used threads in the thread pool to determine an optimal amount of threads for the current use case.

### 2018-06-23 Presentations/Paper/paper_thread_pool.tex
Add first version of related work to paper

@@
 \subsection{Experimental Configuration}
-\textbf{TODO: Change?}\\The operating system is Ubuntu 17.04 with linux kernel version 4.13 running in a virtual machine on an Intel Core i7-6700HQ, 4x 2.6 GHz (Skylake) and 8GB DDR3 RAM running on 1600Mhz.
+\textbf{TODO: Change!}\\The operating system is Ubuntu 17.04 with linux kernel version 4.13 running in a virtual machine on an Intel Core i7-6700HQ, 4x 2.6 GHz (Skylake) and 8GB DDR3 RAM running on 1600Mhz.
 We reduce the number of cores of the virtual machine to 4 instead of 8 virtual cores of the CPU.
@@
 \section{Related Work}
+Implementing multi threading can be realised using different approaches from spawning new threads to reuse them in thread pools.
+Pyarali et al.
+show that a thread pool model can significantly improve system performance and reduce response time \cite{pyarali2001evaluating}.
+They present different approaches to implement thread pools and show the dependency on use cases of these implementations.
 
+Schmidt et al.
+present different implementations of multi threading, including two types of thread pools.
+They distinguish between the \emph{Worker Thread Pool Architecture} and the \emph{Leader/Follower Thread Pool Architecture} \cite{schmidt1998evaluating}.
+In the first approach, an IO-thread selects a connection, reads the requests and dispatches the tasks to an array of threads.
+Drawbacks of this system are excessive context switching effort along with synchronization required to manage the request queue.
+In the second approach, one thread of the pool becomes the leader and does the reading of the task into an internal buffer.
+If the request is valid, the thread itself processes the task and a follower thread becomes the new leader.
+This approach minimizes the context switching effort compared to the first one.
+
+Along with the design of thread pool systems, the performance strongly depends on the number of threads used in a thread pool.
+Xu et al.
+present a performance study for a set of performance metrics combined with a heuristic algorithm for dynamic resizing of thread pools \cite{xu2004performance}.
+They show the correlation between average task idle time and the throughput of thread pools and present the effectiveness of dynamically adjusting the number of threads.
+
+Compared to heuristically adjust the number of threads, Ling et al.
+formalized the problem of finding the optimal number of threads by establishing mathematical relationship among optimal thread pool size, system load and the associated costs \cite{ling2000analysis}.
+They show the applicability of their model and plan to further examine it based on additional metrics.
 \section{Conclusion and Future Work}

### 2018-06-25 include/benchmark_switches.h
+ better distribution for wait slots with compile time switch

@@ -1,0 +1,4 @@
+
+#define UNIFORM_SLOT_DISTRIBUTION
+// 0.848815, 0.847186, 0.847997
+// 9.275028, 9.279833, 9.276998

### 2018-06-25 include/thread_pool.h

@@ -23,6 +23,8 @@
 //	GLOBAL VARIABLES
 //
 
+#define MAX_NUM_TASKS 4097
+
 //
 //	STRUCTS
 //

### 2018-06-25 src/main.c

@@ -6,7 +6,7 @@
 #include "thread_pool.h"
 #include "thread_pool_monitoring.h"
 
-#define WORK_SIZE 40000
+#define WORK_SIZE 128
 #define NUM_TASKS 2048
 
 void work_large(void* args)
@@ -15,7 +15,7 @@ void work_large(void* args)
 	double exp = *res;
 
 	double v = 0.0;
-	for(double i = 0; i < 1024; ++i)
+	for(double i = 0; i < WORK_SIZE; ++i)
 	{
 		v += pow(i, 1 / exp);
 	}
@@ -63,7 +63,7 @@ void resize_test()
 		sum += i*i / 8;
 		if(results[i] != sum) printf("error at %d: %d != %d\n", i, sum, results[i]);
 	}
-	printf("interleaving test completed.");
+	printf("interleaving test completed.\n");
 
 }
 
@@ -99,6 +99,31 @@ void performance_test(int numThreads, int numTasks)
 	thread_pool_free(pool);
 }
 
+float wait_performance_test(int numThreads, int numTasks)
+{
+	thread_pool* pool = thread_pool_create(numThreads, 1);
+
+	double results[numTasks];
+	thread_task tasks[numTasks];
+
+	for(int i = 0; i < numTasks; ++i)
+	{
+		results[i] = (double)(i+1);
+		tasks[i].args = &results[i];
+		tasks[i].routine = work_large;
+		tasks[i].priority = i;
+		task_handle hndl;
+		thread_pool_enqueue_task(&tasks[i], pool, &hndl);
+		if(i % 11 == 0) thread_pool_wait_for_task(pool, &hndl);
+	}
+	thread_pool_wait_for_all(pool);
+
+	float a = thread_pool_get_time_working(pool);
+//	printf("fraction of time spend working: %f\n",a);
+	thread_pool_free(pool);
+	return a;
+}
+
 
 int main()
 {
@@ -118,6 +143,13 @@ int main()
 	performance_test(2, 4000);
 	resize_test();
 
+	float sum = 0.f;
+	for(int i = 0; i < 1000; ++i)
+	{
+		sum += wait_performance_test(2, 1000);
+	}
+	sum /= 1000;
+	printf("average: %f\n", sum);
 	getchar();
 	return 0;
 }

### 2018-06-25 src/thread_pool.c

@@ -2,6 +2,7 @@
 #include <stdlib.h>
 #include <string.h>
 #include <stdio.h>
+#include "benchmark_switches.h"
 
 static inline void __execute_task();
 
@@ -19,7 +20,7 @@ thread_pool* thread_pool_create(size_t num_threads, int enable_monitoring) {
 
   pool->thread_tasks = calloc(num_threads, sizeof(thread_task*));
   pool->thread_infos = calloc(sizeof(__thread_information*) * pool->capacity, 1);
-  pool->task_state_capacity = 4096;
+  pool->task_state_capacity = MAX_NUM_TASKS;
   pool->task_group_states = calloc(pool->task_state_capacity, sizeof(__task_state));
   pool->enable_monitoring = enable_monitoring;
 
@@ -138,9 +139,12 @@ status_e thread_pool_enqueue_tasks(thread_task* tasks, thread_pool* pool, size_t
   
   // find unused slot
   size_t ind = 0;
+#ifdef UNIFORM_SLOT_DISTRIBUTION
+  for(; pool->task_group_states[ind].task_count;ind = (ind + 8) % MAX_NUM_TASKS);
+#else
   for(; ind < pool->task_state_capacity && pool->task_group_states[ind].task_count; ++ind);
   if(ind == pool->task_state_capacity) return status_failed;
-
+#endif
   // increment generation first to always be identifiable as finished
   ++pool->task_group_states[ind].generation;
   pool->task_group_states[ind].task_count = num_tasks;

### 2018-06-26 Presentations/Paper/paper_thread_pool.tex
Enhance Setup section paper

@@
 \section{Experimental Environment}	
-In this section, we present our experimental environment focussing on the design of our benchmarks and the system configuration we used to evaluate our thread pool implementation.
+In this section, we present our experimental environment focusing on the design of our benchmarks and the system configuration we used to evaluate our thread pool implementation.
 
@@
 \subsection{Experimental Configuration}
-\textbf{TODO: Change!}\\The operating system is Ubuntu 17.04 with linux kernel version 4.13 running in a virtual machine on an Intel Core i7-6700HQ, 4x 2.6 GHz (Skylake) and 8GB DDR3 RAM running on 1600Mhz.
-We reduce the number of cores of the virtual machine to 4 instead of 8 virtual cores of the CPU.
-The CPU has 256KB L1 cache, 1MB L2 cache, and 6MB L3 cache.
-All simulations are repeated three times and a median value of all repetitions is used for evaluation.
+%\textbf{TODO: Change!}\\The operating system is Ubuntu 17.04 with linux kernel
+%version 4.13 running in a virtual machine on an Intel Core i7-6700HQ, 4x 2.6 GHz
+%(Skylake) and 8GB DDR3 RAM running on 1600Mhz. We reduce the number of cores of
+%the virtual machine to 4 instead of 8 virtual cores of the CPU. The CPU has
+%256KB L1 cache, 1MB L2 cache, and 6MB L3 cache. All benchmarks have been
+%repeated three times and a median value of all repetitions is being used for evaluation.
+
+For testing purposes we used primarly Linux setups, for one Ubuntu 17.04
+(linux 4.13) in an virtual machine environment, second Archlinux (linux
+4.17.2-1) running native, both updated on stable branch. Tests have also been
+conducted on Windows 10 and macOs HighSierra machines.
+
+In our setups mostly Sky- and Kabylake processors have been used with memory
+ranging from 8 to 24 GB of both DDR3-SODIMM and DDR4-SODIMM.
+
+All benchmarks have been repeated three times and a median value of all
+repetitions is being used for evaluation.
 

### 2018-06-26 include/thread_pool.h
Refactor to comply with conventions, Fix resize test
Also updated some README specifics

@@ -31,92 +31,92 @@
 
 typedef void (*task_routine)(void* routine);
 
-typedef struct thread_task {
+typedef struct thread_task_t {
 	void* args;
 	pthread_attr_t *attr;
 	task_routine routine;
 	size_t group_id;
 	size_t priority;
 	task_stats statistics; 
-} thread_task;
+} thread_task_t;
 
-struct __thread_information;
+struct __thread_information_t;
 
 
-typedef struct __task_state{
+typedef struct __task_state_t{
 	atomic_int task_count; //< remaining tasks in this group
 	unsigned generation;
-} __task_state;
+} __task_state_t;
 
-typedef struct task_handle{
+typedef struct task_handle_t{
 	size_t index;
 	unsigned generation;
-} task_handle;
+} task_handle_t;
 
-typedef struct thread_pool {
+typedef struct thread_pool_t {
 	char* name;
 	pthread_t* pool;
 	priority_queue_t* waiting_tasks;
-	__task_state* task_group_states;
+	__task_state_t* task_group_states;
 	size_t task_state_capacity; // number of tasks that can be tracked
 	size_t size;
 	size_t capacity;
-	struct __thread_information** thread_infos;
-	thread_task** thread_tasks;
+	struct __thread_information_t** thread_infos;
+	thread_task_t** thread_tasks;
 	thread_pool_stats* statistics;
 	int enable_monitoring;
-} thread_pool;
+} thread_pool_t;
 
-typedef struct __thread_information {
+typedef struct __thread_information_t {
 	char name[12];
-	thread_pool* pool;
+	thread_pool_t* pool;
 	size_t id;
 	atomic_int status;
 	struct timespec creation; // cannot be removed, leads to segfault
 	thread_stats* statistics;
-} __thread_information;
+} __thread_information_t;
 
 //
 //	EXTERNAL METHODS
 //
 
-thread_pool* thread_pool_create(size_t num_threads, int enable_monitoring);
-thread_pool* thread_pool_create_named(size_t num_threads, const char* name, int enable_monitoring);
+thread_pool_t* thread_pool_create(size_t num_threads, int enable_monitoring);
+thread_pool_t* thread_pool_create_named(size_t num_threads, const char* name, int enable_monitoring);
 // Releases all resources hold by the threadpool. 
 // Currently working threads may finish but tasks left in the queue will be discarded.
-void thread_pool_free(thread_pool* pool);
+void thread_pool_free(thread_pool_t* pool);
 
[…]

### 2018-06-26 include/thread_pool_monitoring.h

@@ -8,13 +8,13 @@
 //
 
 // Returns the average fraction of time the active threads have been working.
-double thread_pool_get_time_working(thread_pool* pool);
+double thread_pool_get_time_working(thread_pool_t* pool);
 
 // Fill all stats of the passed thread pool instance
-thread_pool_stats thread_pool_get_stats(thread_pool* pool);
+thread_pool_stats thread_pool_get_stats(thread_pool_t* pool);
 
 // Fill all stats of the thread matching the given id in the thread pool
-thread_stats thread_pool_get_thread_stats(thread_pool* pool, size_t id);
+thread_stats thread_pool_get_thread_stats(thread_pool_t* pool, size_t id);
 
 
 

### 2018-06-26 src/main.c

@@ -35,10 +35,10 @@ void work(void* args)
 
 void resize_test()
 {
-	thread_pool* pool = thread_pool_create(2, 0);
-	task_handle hndl;
+	thread_pool_t* pool = thread_pool_create(2, 0);
+	task_handle_t hndl;
 	const int numThreads = 1 << 11;
-	thread_task tasks[numThreads];
+	thread_task_t tasks[numThreads];
 	int results[numThreads];
 	for(int i = numThreads-1; i >= 0; --i){
 		results[i] = i;
@@ -78,10 +78,10 @@ void performance_test(int numThreads, int numTasks)
 	time /= 2048;
 	printf("time for one task: time: %f | %f\n", time, exp);
 
-	thread_pool* pool = thread_pool_create(numThreads, 1);
+	thread_pool_t* pool = thread_pool_create(numThreads, 1);
 
 	double results[numTasks];
-	thread_task tasks[numTasks];
+	thread_task_t tasks[numTasks];
 
 	for(int i = 0; i < numTasks; ++i)
 	{
@@ -101,10 +101,10 @@ void performance_test(int numThreads, int numTasks)
 
 float wait_performance_test(int numThreads, int numTasks)
 {
-	thread_pool* pool = thread_pool_create(numThreads, 1);
+	thread_pool_t* pool = thread_pool_create(numThreads, 1);
 
 	double results[numTasks];
-	thread_task tasks[numTasks];
+	thread_task_t tasks[numTasks];
 
 	for(int i = 0; i < numTasks; ++i)
 	{
@@ -112,7 +112,7 @@ float wait_performance_test(int numThreads, int numTasks)
 		tasks[i].args = &results[i];
 		tasks[i].routine = work_large;
 		tasks[i].priority = i;
-		task_handle hndl;
+		task_handle_t hndl;
 		thread_pool_enqueue_task(&tasks[i], pool, &hndl);
 		if(i % 11 == 0) thread_pool_wait_for_task(pool, &hndl);
 	}

### 2018-06-26 src/thread_pool.c

@@ -10,18 +10,18 @@ static inline void __execute_task();
 //  EXTERNAL METHODS
 //
 
-thread_pool* thread_pool_create(size_t num_threads, int enable_monitoring) {
-  thread_pool* pool = malloc(sizeof(thread_pool));
+thread_pool_t* thread_pool_create(size_t num_threads, int enable_monitoring) {
+  thread_pool_t* pool = malloc(sizeof(thread_pool_t));
  
   pool->name = NULL;
   pool->size = num_threads;
   pool->capacity = num_threads * 2;
   pool->waiting_tasks = calloc(1, sizeof(priority_queue_t));
 
-  pool->thread_tasks = calloc(num_threads, sizeof(thread_task*));
-  pool->thread_infos = calloc(sizeof(__thread_information*) * pool->capacity, 1);
+  pool->thread_tasks = calloc(num_threads, sizeof(thread_task_t*));
+  pool->thread_infos = calloc(sizeof(__thread_information_t*) * pool->capacity, 1);
   pool->task_state_capacity = MAX_NUM_TASKS;
-  pool->task_group_states = calloc(pool->task_state_capacity, sizeof(__task_state));
+  pool->task_group_states = calloc(pool->task_state_capacity, sizeof(__task_state_t));
   pool->enable_monitoring = enable_monitoring;
 
   pthread_t* threads = malloc(sizeof(pthread_t) * pool->capacity);
@@ -34,7 +34,7 @@ thread_pool* thread_pool_create(size_t num_threads, int enable_monitoring) {
 
   for(size_t i = 0; i < pool->capacity; i++) {
     // one block per thread to reduce risk of two threads sharing the same cache line
-    __thread_information* thread_info = malloc(sizeof(__thread_information));
+    __thread_information_t* thread_info = malloc(sizeof(__thread_information_t));
     pool->thread_infos[i] = thread_info;
     thread_info->pool = pool;
     thread_info->id = i;
@@ -52,8 +52,8 @@ thread_pool* thread_pool_create(size_t num_threads, int enable_monitoring) {
    return pool;
 }
 
-thread_pool* thread_pool_create_named(size_t num_threads, const char* name, int enable_monitoring) {
-  thread_pool* pool = thread_pool_create(num_threads, enable_monitoring);
+thread_pool_t* thread_pool_create_named(size_t num_threads, const char* name, int enable_monitoring) {
[…]

### 2018-06-26 src/thread_pool_monitoring.c

@@ -5,7 +5,7 @@
 // EXTERNAL METHODS
 // 
 
-thread_pool_stats thread_pool_get_stats(thread_pool* pool) {
+thread_pool_stats thread_pool_get_stats(thread_pool_t* pool) {
 
   // In case no tasks care completed, no averages can be calculated
   if(pool->statistics->task_complete_count) {
@@ -15,7 +15,7 @@ thread_pool_stats thread_pool_get_stats(thread_pool* pool) {
   return *pool->statistics;
 }
 
-thread_stats thread_pool_get_thread_stats(thread_pool* pool, size_t id) {
+thread_stats thread_pool_get_thread_stats(thread_pool_t* pool, size_t id) {
   thread_stats* thread_stats = pool->thread_infos[id]->statistics;
   
   // busy_time = running_time - idle_time
@@ -29,7 +29,7 @@ thread_stats thread_pool_get_thread_stats(thread_pool* pool, size_t id) {
 // HELP METHODS
 //
 
-double thread_pool_get_time_working(thread_pool* pool){
+double thread_pool_get_time_working(thread_pool_t* pool){
   struct timespec end;
   
   clock_gettime(CLOCK_MONOTONIC, &end);

### 2018-06-26 testing/tests/threadpool/test_pool.hpp

@@ -8,14 +8,14 @@ using atomic_int = std::atomic<int>;
 
 extern "C" 
 {    
-    #include "../../../include/thread_pool.h"
+    #include "thread_pool.h"
     #include "../../../include/thread_pool_monitoring.h"
 }
 
 // Test the creatrion of thread pools
 TEST(ThreadPool, Create) {
 
-    thread_pool* pool = thread_pool_create(2, 0);
+    thread_pool_t* pool = thread_pool_create(2, 0);
     EXPECT_TRUE(pool);
 
     thread_pool_free(pool);
@@ -24,7 +24,7 @@ TEST(ThreadPool, Create) {
 
 // Test the creatrion of group Ids of thread pools
 TEST(ThreadPool, Name) {
-    thread_pool* pool = thread_pool_create_named(2, "ThreadPool", 0);
+    thread_pool_t* pool = thread_pool_create_named(2, "ThreadPool", 0);
     EXPECT_TRUE(pool);
     ASSERT_STREQ(pool->name, "ThreadPool");
 
@@ -58,7 +58,7 @@ void work(void* args)
 
 TEST(ThreadPool, Resize) {
 
-    thread_pool* pool = thread_pool_create(4, 0);
+    thread_pool_t* pool = thread_pool_create(4, 0);
     EXPECT_TRUE(pool);
 
     status_e status = thread_pool_resize(pool, 6);
@@ -74,8 +74,8 @@ TEST(ThreadPool, Resize) {
 TEST(ThreadPool, WAIT){
 
     int test[] = {1000000, 1000000, 1000000, 1000000, 1000000, 1000000 };
-    thread_pool* pool = thread_pool_create(2, 0);
-    thread_task tasks[6];
+    thread_pool_t* pool = thread_pool_create(2, 0);
+    thread_task_t tasks[6];
 
     for(int i = 0; i < 6; i++){
         tasks[i].args = (void*)&test[i];
@@ -97,8 +97,8 @@ TEST(ThreadPool, WAIT){
 TEST(ThreadPool, WAIT_FOR_ALL){
 
     int test[] = {1000000, 1000000, 1000000, 1000000, 1000000, 1000000 };
-    thread_pool* pool = thread_pool_create(2, 0);
-    thread_task tasks[6];
+    thread_pool_t* pool = thread_pool_create(2, 0);
[…]

### 2018-06-26 testing_c/tests/thread_pool/test_thread_pool.h

@@ -1,7 +1,7 @@
 #include <check.h>
 #include <stdio.h>
 #include <math.h>
-#include "../../../include/thread_pool.h"
+#include "thread_pool.h"
 
 // Test the creation of thread pools
 START_TEST(thread_pool_create_) 

### 2018-06-26 src/main.c
Change README
Forgot to make some changes

@@ -38,7 +38,7 @@ void resize_test()
 	thread_pool_t* pool = thread_pool_create(2, 0);
 	task_handle_t hndl;
 	const int numThreads = 1 << 11;
-	thread_task_t tasks[numThreads];
+	thread_task_t    tasks[numThreads];
 	int results[numThreads];
 	for(int i = numThreads-1; i >= 0; --i){
 		results[i] = i;

### 2018-06-26 include/thread_pool.h
Add basic graphs and csv for pool statistics(broken)

@@ -44,7 +44,7 @@ struct __thread_information_t;
 
 
 typedef struct __task_state_t{
-	atomic_int task_count; //< remaining tasks in this group
+	atomic_int task_count; // remaining tasks in this group
 	unsigned generation;
 } __task_state_t;
 
@@ -72,7 +72,6 @@ typedef struct __thread_information_t {
 	thread_pool_t* pool;
 	size_t id;
 	atomic_int status;
-	struct timespec creation; // cannot be removed, leads to segfault
 	thread_stats* statistics;
 } __thread_information_t;
 

### 2018-06-26 testing/tests/threadpool/test_pool.hpp

@@ -1,6 +1,7 @@
 #include <gtest/gtest.h>
 #include <atomic>
 #include <cmath>
+#include <fstream>
 
 // https://gcc.gnu.org/bugzilla/show_bug.cgi?id=60932
 using atomic_int = std::atomic<int>;
@@ -194,30 +195,40 @@ TEST(ThreadPool, TASK_STATISTICS){
     thread_pool_free(pool);
 }
 
-TEST(ThreadPool, POOL_STATISTICS){
+TEST(ThreadPool, POOL_STATISTICS) {
 
-    int test[] = {1000000, 1000000, 1000000, 1000000, 1000000, 1000000 };
-    thread_pool_t* pool = thread_pool_create(2, 1);
-    thread_task_t tasks[6];
-    thread_pool_stats pool_stats;
+    std::ofstream stats;
+    stats.open("Statistics/pool_avg.csv");
+    stats << "Iteration Avg.CompletionTime Avg.WaitingTime" << "\n";
+    for (size_t counter = 0; counter < 10; ++counter) {
 
-    for(int i = 0; i < 6; i++){
-        tasks[i].args = (void*)&test[i];
-        tasks[i].routine = work;
-        thread_pool_enqueue_task(&tasks[i], pool, NULL);
+        int test[] = {1000000, 1000000, 1000000, 1000000, 1000000, 1000000};
+        thread_pool_t *pool = thread_pool_create(counter, 1);
+        thread_task_t tasks[6];
+        thread_pool_stats pool_stats;
+
+        for (int i = 0; i < 6; i++) {
+            tasks[i].args = (void *) &test[i];
+            tasks[i].routine = work;
+            thread_pool_enqueue_task(&tasks[i], pool, NULL);
+        }
+
+        pool_stats = thread_pool_get_stats(pool);
+        ASSERT_EQ(pool_stats.task_enqueued_count, 6);
+        ASSERT_NE(pool_stats.task_complete_count, 6);
+
+        thread_pool_wait_for_all(pool);
+
+        pool_stats = thread_pool_get_stats(pool);
+        ASSERT_EQ(pool_stats.task_complete_count, pool_stats.task_enqueued_count);
+
+        //save to csv
+        stats << counter << pool_stats.avg_complete_time << " " << pool_stats.avg_wait_time << "\n";
+
+
+        thread_pool_free(pool);
     }
-
-    pool_stats = thread_pool_get_stats(pool);
[…]

### 2018-06-28 include/thread_pool.h
Minor changes

@@ -117,5 +117,6 @@ thread_task_t* __get_next_task(thread_pool_t *pool);
 
 status_e __create_thread(__thread_information_t* thread_info, pthread_t* pp);
 thread_status_e __update_thread_status(thread_pool_t* pool, size_t thread_id, thread_status_e thread_status);
+void __sig_seg(int sig);
 
 #endif //THREAD_POOL_H

### 2018-06-28 src/main.c

@@ -49,7 +49,7 @@ void resize_test()
 		thread_pool_enqueue_task(&tasks[i], pool, &hndl);
 		if( i == numThreads * 1 / 3) thread_pool_resize(pool, 4);
 		if(i == numThreads * 2 / 3) thread_pool_resize(pool, 3);
-		if(i == numThreads * 3 / 4) thread_pool_resize(pool,2);
+		if(i == numThreads * 3 / 4) thread_pool_resize(pool, 2);
 	}
 
 //	thread_pool_wait_for_task(pool, &hndl);
@@ -99,7 +99,7 @@ void performance_test(int numThreads, int numTasks)
 	thread_pool_free(pool);
 }
 
-float wait_performance_test(int numThreads, int numTasks)
+float wait_performance_test(int numThreads, int numTasks, FILE *ps)
 {
 	thread_pool_t* pool = thread_pool_create(numThreads, 1);
 
@@ -120,6 +120,8 @@ float wait_performance_test(int numThreads, int numTasks)
 
 	float a = thread_pool_get_time_working(pool);
 //	printf("fraction of time spend working: %f\n",a);
+    thread_pool_stats stats = thread_pool_get_stats(pool);
+    fprintf(ps, "%i %lli %lli \n", numThreads, stats.avg_complete_time, stats.avg_wait_time);
 	thread_pool_free(pool);
 	return a;
 }
@@ -142,14 +144,18 @@ int main()
 	}*/
 	performance_test(2, 4000);
 	resize_test();
+	FILE *ps;
+
+	ps = fopen("Statistics/pool_avg.csv","w+");
+    fprintf(ps, "Threads BusyTime IdleTime \n");
 
 	float sum = 0.f;
-	for(int i = 0; i < 1000; ++i)
+	for(int i = 2; i < 16; ++i)
 	{
-		sum += wait_performance_test(2, 1000);
+		sum += wait_performance_test(i, 1000,ps);
 	}
 	sum /= 1000;
 	printf("average: %f\n", sum);
-	getchar();
+//	getchar();
 	return 0;
 }

### 2018-06-28 src/thread_pool.c

@@ -2,6 +2,7 @@
 #include <stdlib.h>
 #include <string.h>
 #include <stdio.h>
+#include <signal.h>
 #include "benchmark_switches.h"
 
 static inline void __execute_task();
@@ -309,8 +310,24 @@ status_e __create_thread(__thread_information_t* thread_info, pthread_t* pp){
   return status_ok;
 }
 
+void* faulty;
+size_t num;
+
 void __execute_task(thread_pool_t* pool, thread_task_t* task)
 {
-  (*task->routine)(task->args);
-  --pool->task_group_states[task->group_id].task_count;
+    faulty = task->routine;
+    num++;
+    (*task->routine)(task->args);
+    --pool->task_group_states[task->group_id].task_count;
+}
+
+void __sig_seg(int sig) {
+    if (sig != SIGSEGV) {
+        return;
+    }
+    else {
+        printf("%p \n", faulty);
+        printf("occured after: %li", num);
+        exit(1);
+    }
 }

### 2018-06-28 testing/tests/threadpool/test_pool.hpp

@@ -2,6 +2,7 @@
 #include <atomic>
 #include <cmath>
 #include <fstream>
+#include <signal.h>
 
 // https://gcc.gnu.org/bugzilla/show_bug.cgi?id=60932
 using atomic_int = std::atomic<int>;
@@ -13,8 +14,8 @@ extern "C"
     #include "../../../include/thread_pool_monitoring.h"
 }
 
-// Test the creatrion of thread pools
-TEST(ThreadPool, Create) {
+// Test the creation of thread pools
+TEST(ThreadPool, CREATE) {
 
     thread_pool_t* pool = thread_pool_create(2, 0);
     EXPECT_TRUE(pool);
@@ -23,18 +24,13 @@ TEST(ThreadPool, Create) {
 
 }
 
-// Test the creatrion of group Ids of thread pools
-TEST(ThreadPool, Name) {
+// Test the creation of group Ids of thread pools
+TEST(ThreadPool, NAME) {
     thread_pool_t* pool = thread_pool_create_named(2, "ThreadPool", 0);
     EXPECT_TRUE(pool);
     ASSERT_STREQ(pool->name, "ThreadPool");
 
     thread_pool_free(pool);
-}
-
-void basicTask (void* args) {
-    int* number = (int*) args;
-    std::cout << "number has value: " <<  *number << std::endl;
 }
 
 void work(void* args)
@@ -51,13 +47,13 @@ void work(void* args)
 
 	for(; *res > 0; (*res)--)
     {
-        
+
     }
 
     //std::cout << "Work finished" << std::endl;
 }
 
-TEST(ThreadPool, Resize) {
+TEST(ThreadPool, RESIZE) {
 
     thread_pool_t* pool = thread_pool_create(4, 0);
     EXPECT_TRUE(pool);
@@ -97,7 +93,7 @@ TEST(ThreadPool, WAIT){
 
 TEST(ThreadPool, WAIT_FOR_ALL){
 
-    int test[] = {1000000, 1000000, 1000000, 1000000, 1000000, 1000000 };
+    int test[] = { 1000000, 1000000, 1000000, 1000000, 1000000, 1000000 };
     thread_pool_t* pool = thread_pool_create(2, 0);
     thread_task_t tasks[6];
 
@@ -197,25 +193,29 @@ TEST(ThreadPool, TASK_STATISTICS){
 
 TEST(ThreadPool, POOL_STATISTICS) {
 
-    std::ofstream stats;
[…]

### 2018-06-28 src/main.c
Add utilization chart

@@ -99,7 +99,7 @@ void performance_test(int numThreads, int numTasks)
 	thread_pool_free(pool);
 }
 
-float wait_performance_test(int numThreads, int numTasks, FILE *ps)
+float wait_performance_test(int numThreads, int numTasks, FILE *ps, FILE *pu)
 {
 	thread_pool_t* pool = thread_pool_create(numThreads, 1);
 
@@ -121,6 +121,7 @@ float wait_performance_test(int numThreads, int numTasks, FILE *ps)
 	float a = thread_pool_get_time_working(pool);
 //	printf("fraction of time spend working: %f\n",a);
     thread_pool_stats stats = thread_pool_get_stats(pool);
+    fprintf(pu, "%i %f \n",numThreads, a);
     fprintf(ps, "%i %lli %lli \n", numThreads, stats.avg_complete_time, stats.avg_wait_time);
 	thread_pool_free(pool);
 	return a;
@@ -145,17 +146,19 @@ int main()
 	performance_test(2, 4000);
 	resize_test();
 	FILE *ps;
+	FILE *pu;
 
 	ps = fopen("Statistics/pool_avg.csv","w+");
     fprintf(ps, "Threads BusyTime IdleTime \n");
-
-	float sum = 0.f;
-	for(int i = 2; i < 16; ++i)
+    pu = fopen("Statistics/pool_util.csv","w+");
+    fprintf(pu, "Threads Util \n");
+//	float sum = 0.f;
+	for(int i = 2; i < 33; ++i)
 	{
-		sum += wait_performance_test(i, 1000,ps);
+		wait_performance_test(i, 6000,ps, pu);
 	}
-	sum /= 1000;
-	printf("average: %f\n", sum);
+//	sum /= 1000;
+//	printf("average: %f\n", sum);
 //	getchar();
 	return 0;
 }

### 2018-07-01 src/main.c
Some Statistic stuff

@@ -5,9 +5,19 @@
 #include "priority_queue.h"
 #include "thread_pool.h"
 #include "thread_pool_monitoring.h"
+#include <time.h>
+#include <pthread.h>
 
-#define WORK_SIZE 128
+#define WORK_SIZE 4096
 #define NUM_TASKS 2048
+#define MEASURE_COUNT 10 
+
+static double test_results_pu[64];
+static long long test_results_ps1[64];
+static long long test_results_ps2[64];
+
+static long long test_results_baseline[64];
+static long long test_results_pool[64];
 
 void work_large(void* args)
 {
@@ -99,7 +109,7 @@ void performance_test(int numThreads, int numTasks)
 	thread_pool_free(pool);
 }
 
-float wait_performance_test(int numThreads, int numTasks, FILE *ps, FILE *pu)
+float wait_performance_test(int numThreads, int numTasks)
 {
 	thread_pool_t* pool = thread_pool_create(numThreads, 1);
 
@@ -114,19 +124,72 @@ float wait_performance_test(int numThreads, int numTasks, FILE *ps, FILE *pu)
 		tasks[i].priority = i;
 		task_handle_t hndl;
 		thread_pool_enqueue_task(&tasks[i], pool, &hndl);
-		if(i % 11 == 0) thread_pool_wait_for_task(pool, &hndl);
+		//if(i % 11 == 0) thread_pool_wait_for_task(pool, &hndl);
 	}
 	thread_pool_wait_for_all(pool);
 
 	float a = thread_pool_get_time_working(pool);
 //	printf("fraction of time spend working: %f\n",a);
     thread_pool_stats stats = thread_pool_get_stats(pool);
-    fprintf(pu, "%i %f \n",numThreads, a);
-    fprintf(ps, "%i %lli %lli \n", numThreads, stats.avg_complete_time, stats.avg_wait_time);
+    
+	test_results_pu[numThreads] += a;
+	test_results_ps1[numThreads] += stats.avg_complete_time;
+	test_results_ps2[numThreads] += stats.avg_wait_time;
+	//fprintf(pu, "%i %f \n",numThreads, a);
+    	//fprintf(ps, "%i %lli %lli \n", numThreads, stats.avg_complete_time, stats.avg_wait_time);
 	thread_pool_free(pool);
 	return a;
 }
 
[…]

### 2018-07-01 Presentations/Paper/paper_thread_pool.tex
Start evaluation in paper

@@
 \usepackage{textcomp}
+\usepackage{listings}
+\usepackage{xcolor}
+\usepackage{listings}
+
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 \begin{document}
+
+\definecolor{mGreen}{rgb}{0,0.6,0}
+\definecolor{mGray}{rgb}{0.5,0.5,0.5}
+\definecolor{mPurple}{rgb}{0.58,0,0.82}
+\definecolor{backgroundColour}{rgb}{0.95,0.95,0.92}
+
+\lstdefinestyle{CStyle}{
+	backgroundcolor=\color{backgroundColour},   
+	commentstyle=\color{mGreen},
+	keywordstyle=\color{blue},
+	numberstyle=\tiny\color{mGray},
+	stringstyle=\color{mPurple},
+	basicstyle=\footnotesize,
+	breakatwhitespace=false,         
+	breaklines=true,                 
+	captionpos=b,                    
+	keepspaces=true,                 
+	numbers=left,                    
+	numbersep=5pt,                  
+	showspaces=false,                
+	showstringspaces=false,
+	showtabs=false,                  
+	tabsize=2,
+	language=C,
+	morekeywords={*,thread\_pool\_t, thread\_task\_t}
+}
 
@@
 
-Furthermore, we compare the scope and usage of our thread pool against an existing implementation \emph{Threading Building Blocks (Intel TBB)} \cite{kim2011multicore} written in C++.
+%Furthermore, we compare the scope and usage of our thread pool against an existing implementation \emph{Threading Building Blocks (Intel TBB)} \cite{kim2011multicore} written in C++.
 Since Intel TBB uses completely different approaches to parallel the execution of tasks, a direct performance comparison may not lead to reliable results.
@@
 \section{Evaluation}
+In this Section, we present the results of our evaluation according to the evaluation setup we stated in the previous section.
+We start with an introduction of the benchmark methods to compare our thread pool against the baseline implementation and present the results.
+Furthermore, we show the relation of waiting time and completing regarding different thread pool configurations.
+
+\subsection{Comparison against Baseline Implementation}
+As most important aspect of our evaluation, we compare our thread pool against an approximated version of the baseline implementation.
+Since BOLSTER is deeply integrated into geckoDB and the workload and usage of multithreading depends on the graph data in the system, we can not directly use the BOLSTER implementation for our evaluation.
+Thus, we extracted the respecting code parts into an own implementation and use this for a direct comparison against the thread pool.
+In Figure \ref{fig4}, we show the approximated version of the BOLSTER multithreading code that we use as baseline implementation.
+In our benchmark.
+the method \emph{cmp\_baseline} is called multiple times with a different \emph{numThreads} parameter.
+Within each call, \emph{numThreads} threads are created and all of these execute the benchmark method v, that calculates a huge amount of exponential values in a loop.
+After the calling thread takes part in execution, it waits until all created threads are joined.
+The elapsed processing time is measured from the begin of the method until joining all threads finished.
+
+\begin{figure}
+	\begin{lstlisting}[style=CStyle]
+	void cmp_baseline(int numThreads)
+	{
+	struct timespec begin, end;
+	clock_gettime(CLOCK_MONOTONIC, &begin);
+
+	double results[numThreads];
+	thread_task_t tasks[numThreads];
+
+	pthread_t threads[numThreads];
+	for (int tid = 1; tid < numThreads; tid++)
+	{
+	results[tid] = (double)(tid+1);
+	pthread_create(&threads[tid] , NULL, 
+	&work_large, &results[tid]);
+	}
+
+	// Let the calling thread process one 
+	// task before waiting
+	work_large(&results[0]);
+
+	for (int tid = 1; tid < numThreads; tid++) {
+	pthread_join(threads[tid], NULL);
+	}
+
+	clock_gettime(CLOCK_MONOTONIC, &end);
+	test_results_baseline[numThreads] +=
+	__get_time_diff(&begin, &end);
+	}
+	\end{lstlisting}
+	\caption{Baseline multithreading implementation extracted from BOLSTER}
+	\label{fig4}
+\end{figure}
+
+To compare against the presented baseline implementation, we developed a method that does the same amount of calls of \emph{work\_large} as the baseline, but uses a thread pool instance instead of creating new threads.
+In Figure \ref{fig5}, we show this implementation called \emph{cmp\_pool} that creates and fills tasks and enqueues these into a thread pool instance.
+Since we assume, that the thread pools are reused for multiple requests, we exclude the thread pool creation and deallocation from the time measure.
+Consequently, only creating and processing tasks is included in the considered elapsed time.
+
+\begin{figure}
+	\begin{lstlisting}[style=CStyle]
+	void cmp_pool(int numThreads)
+	{
+	thread_pool_t* pool = 
+	thread_pool_create(numThreads, 1);
+
+	struct timespec begin, end;
+	clock_gettime(CLOCK_MONOTONIC, &begin);
+	double results[numThreads];
+	thread_task_t tasks[numThreads];
+
+	for(int i = 0; i < numThreads; ++i)
+	{
+	results[i] = (double)(i+1);
+	tasks[i].args = &results[i];
+	tasks[i].routine = work_large;
+	thread_pool_enqueue_task(&tasks[i], pool, NULL);
+	}
+
+	thread_pool_wait_for_all(pool);
+
+	clock_gettime(CLOCK_MONOTONIC, &end);
+	test_results_pool[numThreads] += 
+	__get_time_diff(&begin, &end);
+
+	thread_pool_free(pool);
+	}
+	\end{lstlisting}
+	\caption{Thread pool benchmark to compare against baseline implementation}
+	\label{fig5}
+\end{figure}
+
+In our performance comparison, both implementations are executed three times with $4, 8, 16, 32$ as \emph{numThreads} parameter and the summed results are divided by three.
+We consider a number between $4$ and $32$ as a useful amount of threads to reach a sufficient performance increase due to parallelization, but also to have the possibility to debug and profile them.
+
+In Figure \ref{fig6}, we show the comparison results of both mentioned implementations and the different \emph{numThreads} parameters.
+
+\begin{figure}
+	\includegraphics[width=0.5\textwidth]{img/pool_baseline.png}
+	\caption{Time measures in nanoseconds of thread pool (blue) and baseline (orange) implementation with different numbers of threads.  }
+	\label{fig6}
+\end{figure}
 

### 2018-07-01 Presentations/Paper/paper_thread_pool.tex

--- 2018-06-26 Presentations/Paper/paper_thread_pool.tex
@@
 
-\title{Thread Pool in GeckoDB/BOLSTER - Implementing a Thread Pool for a high performance Database System \\
+\title{Thread Pool in GeckoDB/BOLSTER - Implementing a Thread Pool for a High Performance Database System \\
 }
@@
 \begin{abstract}
-Processing tasks in parallel is used in nearly all applications to keep up with the requirements of modern software systems.
-However, the implementation of parallel processing can reach from spawning dozens of threads to reuse threads in thread pools.
-In this paper, we show an implementation of a thread pool to process independent tasks in parallel including the waiting for a set of processed tasks.
-We also compare the thread pool implementation against a baseline.
-The evaluation of the thread pool shows, that designing and implementing a thread pool is a balancing act between generalization of tasks and the processing performance.
-Additionally, we show that the task and thread pool configuration depending on the use case has a high impact on the thread pool performance.
+        Processing tasks in parallel is used in nearly all applications to keep
+        up with the requirements of modern software systems. However, the
+        current implementation of parallel processing in GeckoDB requires
+        spawning many short-lived threads that execute a single task and then
+        terminate. This creates a lot of overhead since the threads are not
+        being reused. To counter this effect we implemented a thread pool to be
+        used in combination with GeckoDB to reduce the additional time required
+        to create and close thousands of threads. In this paper, we show our
+        implementation of a thread pool to process independent tasks in parallel
+        including the waiting for a set of processed tasks. We also compare the
+        thread pool implementation against the current one. The evaluation of
+        the thread pool shows, that designing and implementing a thread pool is
+        a balancing act between generalization of tasks and the processing
+        performance. Additionally, we show that the task and thread pool configuration
+        depending on the use case has a high impact on the thread
+        pool performance. At the end of our evalutation we rate the implementation a
+        success because we fulfilled all given constraints and generally reduce
+        overhead compared to creating and spawning threads individually.
 \end{abstract}
@@
 \subsection{Task Configuration}
-We define a Task as a structure containing data that has to be processed and an operation that has to be executed on the data.
+We define a Task as a structure referencing data that has to be processed and an operation that has to be executed on the data.
 In this work, we define tasks as \emph{independent}, which means tasks do not have dependencies on other tasks and can be processed independently.
@@
 	\begin{center}
-		\begin{tabular}{ c c }
+		\begin{tabular}{ l l }
 			\hline
@@
 
-To solve this problem, we implement the waiting for a set of tasks using a slotmap.
-In Figure 3, we show the concept of our slotmap.
-During the enqueueing process of tasks, a set of tasks is mapped to a slot in the slotmap.
+To solve this problem, we implement the waiting for a set of tasks using a slot-map.
+In Figure 3, we show the concept of our slot-map.
+During the enqueueing process of tasks, a set of tasks is mapped to a slot in the slot-map.
 Each slot contains a number of open tasks to be finished and a generation counter.
@@
 	\includegraphics[width=0.4\textwidth]{img/waitingconcept.png}
-	\caption{Concept of waiting for tasks using a slotmap}
+	\caption{Concept of waiting for tasks using a slot-map}
 	\label{fig2}

### 2018-07-01 Presentations/Paper/paper_thread_pool.tex

--- 2018-07-01 Presentations/Paper/paper_thread_pool.tex
@@
 
-\title{Thread Pool in GeckoDB/BOLSTER - Implementing a Thread Pool for a high performance Database System \\
-}
+\title{Thread Pool in GeckoDB/BOLSTER - Implementing a Thread Pool for a High Performance Database System \\}
 
@@
 
+\subsection{TODO: Validity of Results}
+
 \subsection{Comparison against Baseline Implementation}
 As most important aspect of our evaluation, we compare our thread pool against an approximated version of the baseline implementation.
+To justify the performance comparison results, we present the different benchmark implementations in detail before we show the evaluation results.
+
+\subsubsection{Benchmark Implementations}
 Since BOLSTER is deeply integrated into geckoDB and the workload and usage of multithreading depends on the graph data in the system, we can not directly use the BOLSTER implementation for our evaluation.
 Thus, we extracted the respecting code parts into an own implementation and use this for a direct comparison against the thread pool.
+
 In Figure \ref{fig4}, we show the approximated version of the BOLSTER multithreading code that we use as baseline implementation.
@@
 
+\subsubsection{Comparison Results}
 In our performance comparison, both implementations are executed three times with $4, 8, 16, 32$ as \emph{numThreads} parameter and the summed results are divided by three.
@@
 In Figure \ref{fig6}, we show the comparison results of both mentioned implementations and the different \emph{numThreads} parameters.
+Our thread pool implementation is faster than the baseline for all \emph{numThreads} parameters.
+Based on Figure \ref{fig6}, we make the following conclusions.
+
+Firstly, compared to strict processing workload, create, and later on join threads, consumes much time.
+Our thread pool implementation reaches up to a multiple of performance increase compared to the baseline.
+Since the creation and join of threads is the main difference between both tests, reusing threads will lead to this noticeable performance increase.
+
+Secondly, the task creation and enqueuing does not have heavy impact on the performance, since the difference in execution time \textbf{TODO: Complete!}
 
@@
 	\label{fig6}
+\end{figure}
+
+\subsubsection{Summary}
+As mentioned before, our implementation reaches better performance than the baseline regarding our specific tests.
+
+\subsection{Waiting Time Evaluation}
+\begin{figure}
+	\includegraphics[width=0.5\textwidth]{img/pool_avg.png}
+	\caption{Time measures in nanoseconds of thread pool (blue) and baseline (orange) implementation with different numbers of threads.  }
+	\label{fig7}
 \end{figure}

### 2018-07-02 Presentations/Paper/paper_thread_pool.tex
Continue Evaluation on Paper

@@
 \section{Evaluation}
-In this Section, we present the results of our evaluation according to the evaluation setup we stated in the previous section.
+In this Section, we present the results of our evaluation according to the setup we stated in the previous section.
 We start with an introduction of the benchmark methods to compare our thread pool against the baseline implementation and present the results.
@@
 
-\subsection{TODO: Validity of Results}
+\subsection{Validity of Results}
+Before presenting our evaluation results, we give some preliminaries about the validity of our results.
+At first, all measurements are executed on all machines we listed in our experimental configuration, but in this work we present only the results measured on Archlinux with kernel 4.17.2-1 and 24 GB DDR4-SODIMM.
+We decided to use this system, since it is running on a high kernel version and not in a virtual environment.
+Additionally, we use gnuplot to generate diagrams from our results, consequently using a windows machine is not optimal.
+
+Secondly, the measured results represent the usage of our thread pool implementation, but may differ strongly in other applications.
+Depending on the number of threads used outside the thread pools and the ability to parallelize work, the performance may increase less than shown in our tests.
+
+Thirdly, measuring multi threading performance highly depends on the current workload of the system and the scheduler.
+We determine this in our evaluation since tests differ strongly in their results.
+Although we make average measurements for each benchmark, these represent the performance increase at the current system workload.
 
@@
 				tasks[i].routine = work_large;
-	thread_pool_enqueue_task(&tasks[i], pool, NULL);
+				thread_pool_enqueue_task(&tasks[i], pool, 
+					NULL);
 			}

### 2018-07-02 src/main.c
Update README && Fix Statistics save to file

@@ -13,11 +13,16 @@
 #define MEASURE_COUNT 10 
 
 static double test_results_pu[64];
+static double test_results_puw[64];
 static long long test_results_ps1[64];
 static long long test_results_ps2[64];
+static long long test_results_psw1[64];
+static long long test_results_psw2[64];
+
 
 static long long test_results_baseline[64];
 static long long test_results_pool[64];
+static long long test_results_batch[64];
 
 void work_large(void* args)
 {
@@ -47,7 +52,7 @@ void resize_test()
 {
 	thread_pool_t* pool = thread_pool_create(2, 0);
 	task_handle_t hndl;
-	const int numThreads = 1 << 11;
+	const size_t numThreads = 1 << 11;
 	thread_task_t    tasks[numThreads];
 	int results[numThreads];
 	for(int i = numThreads-1; i >= 0; --i){
@@ -68,32 +73,23 @@ void resize_test()
 
 	// verify results
 	int sum = 0;
-	for(int i = 0; i < numThreads; ++i)
+	for(size_t i = 0; i < numThreads; ++i)
 	{
 		sum += i*i / 8;
-		if(results[i] != sum) printf("error at %d: %d != %d\n", i, sum, results[i]);
+		if(results[i] != sum) printf("error at %lu: %d != %d\n", i, sum, results[i]);
 	}
 	printf("interleaving test completed.\n");
 
 }
 
-void performance_test(int numThreads, int numTasks, FILE *pu, FILE *ps)
+void performance_test(size_t numThreads, size_t numTasks)
 {
-//	clock_t begin = clock();
-//	double exp = 2.7;
-//	for(int i = 0; i < 2048; ++i)
-//		work_large(&exp);
-//	clock_t end = clock();
-//	float time = (double)(end - begin) / CLOCKS_PER_SEC;
-//	time /= 2048;
-//	printf("time for one task: time: %f | %f\n", time, exp);
-
 	thread_pool_t* pool = thread_pool_create(numThreads, 1);
 
 	double results[numTasks];
 	thread_task_t tasks[numTasks];
 
-	for(int i = 0; i < numTasks; ++i)
+	for(size_t i = 0; i < numTasks; ++i)
 	{
 		results[i] = (double)(i+1);
 		tasks[i].args = &results[i];
[…]

### 2018-07-02 Presentations/Paper/paper_thread_pool.tex
Add Baseline compare to paper

@@
 			double results[numThreads];
-			thread_task_t tasks[numThreads];
 
@@
 
-Secondly, the task creation and enqueuing does not have heavy impact on the performance, since the difference in execution time \textbf{TODO: Complete!}
+Secondly, the task creation and enqueuing have only slight on the performance, since the baseline does not include these operations, but is still clearly faster than the thread pool implementation.
 

### 2018-07-02 Presentations/Paper/paper_thread_pool.tex
Add first complete version of evaluation

@@
 
-\subsubsection{Benchmark Implementations}
+\subsubsection{Benchmark Implementation}
 Since BOLSTER is deeply integrated into geckoDB and the workload and usage of multithreading depends on the graph data in the system, we can not directly use the BOLSTER implementation for our evaluation.
@@
 
-\subsubsection{Comparison Results}
+\subsubsection{Evaluation Results}
 In our performance comparison, both implementations are executed three times with $4, 8, 16, 32$ as \emph{numThreads} parameter and the summed results are divided by three.
@@
 
-\subsubsection{Summary}
-As mentioned before, our implementation reaches better performance than the baseline regarding our specific tests.
+\subsection{Waiting Time Evaluation}
+As mentioned before, instead of implementing explicit scheduling to assign tasks to threads, we use the \emph{first come first serve} approach.
+Each thread that currently has no task, checks the priority queue to pop a new task in a loop.
+Since our priority queue manages the memory access from different threads using a mutex thread, the queue is locked while a thread asks for a task.
+Consequently, in large thread pools, the accessing time of the priority queue may increase along with the waiting time of threads.
+To examine this behaviour with increasing thread pool size, we measure the average waiting time of thread pools with different sizes.
 
-\subsection{Waiting Time Evaluation}
+\subsubsection{Benchmark Implementation}
+To examine the waiting time behaviour of larger thread pools, we use a similar implementation than the \emph{cmp\_pool} method, which creates 10000 tasks at different thread pool sizes.
+Again, we consider sizes from $4$ to $32$ threads as useful test case and take the average value of three measurements.
+To measure waiting and completion time, we use the monitoring component of our system and evaluate the thread pool statistics of each instance.
+
+\subsubsection{Evaluation Results}
+In Figure \ref{fig7}, we show the relation of waiting and completion time of different thread pool sizes.
+The waiting time is, compared to the completion time, very small and nearly constant for all considered thread pool sizes.
+From this observation, we draw the following conclusions for our \emph{first come first serve} scheduling approach.
+Firstly, constant average waiting time indicates only small performance decreases due to priority queue locking.
+Secondly, even for $32$ threads in one pool instance, the waiting time occupies only a small fraction of the completion time, which indicates good thread usage.
+
+As mentioned before in this section, the performance results depend on the current workload of the system, as seen by the waiting time shown in Figure \ref{fig7}.
+Although the priority queue locking by the mutex thread should increase the waiting time slightly with bigger thread pool sizes, the last measuring point is lower than the previous points.
+
 \begin{figure}
 	\includegraphics[width=0.5\textwidth]{img/pool_avg.png}
-	\caption{Time measures in nanoseconds of thread pool (blue) and baseline (orange) implementation with different numbers of threads.  }
+	\caption{Relation between waiting time (orange) and completion time (blue) of thread pools with different sizes processing 10000 tasks}
 	\label{fig7}
 \end{figure}
+
+\subsection{Summary}
+As last step of the evaluation, we summarize our results and compare them to the requirements of geckoDB.
+As mentioned before, our implementation reaches better performance than the baseline regarding our specific tests.
+Consequently, we solved the task for implementing and integrating the thread pool into BOLSTER to reach at least the baseline performance.
+Additionally, using our thread pool implementation solves the problem, that debugging and profiling with a huge number of threads is not possible.
+
+Although a small and constant waiting time was not a requirement for integration into geckoDB, we ensure that assigning many tasks work fast for different thread pool sizes.
+Regarding the performance and also the design requirements of geckoDB and BOLSTER, we fulfil these requirements and successfully implemented and evaluated our thread pool implementation.
 
@@
 They show the applicability of their model and plan to further examine it based on additional metrics.
+
 \section{Conclusion and Future Work}

### 2018-07-02 Presentations/Paper/paper_thread_pool.tex
Add first version of conclusion

@@
         including the waiting for a set of processed tasks. We also compare the
-        thread pool implementation against the current one. The evaluation of
-        the thread pool shows, that designing and implementing a thread pool is
-        a balancing act between generalization of tasks and the processing
-        performance. Additionally, we show that the task and thread pool configuration
+        thread pool implementation against the current one. Additionally, we show that the task and thread pool configuration
         depending on the use case has a high impact on the thread
-        pool performance. At the end of our evalutation we rate the implementation a
-        success because we fulfilled all given constraints and generally reduce
+        pool performance. At the end of our evaluation, we show that the implementation 
+        fulfils all given constraints and generally reduce
         overhead compared to creating and spawning threads individually.
@@
 	\item We compare our thread pool against the existing implementation in BOLSTER
-	\item We measure and evaluate statistics for threads, tasks and the thread pool
+	\item We measure and evaluate additional statistics for different thread pool configurations
 \end{itemize}
@@
 Processing tasks in a thread pool is a balancing act between generalization the of tasks and the performance of their execution.
+With higher generalization, the implementation can be used for more use cases, but the execution performance may decrease.
 Therefore, we provide a set of functions to process task in the thread pool to match different requirements of BOLSTER.
@@
 Secondly, even for $32$ threads in one pool instance, the waiting time occupies only a small fraction of the completion time, which indicates good thread usage.
+Thirdly, the execution performance increases along with the thread pool size, which shows that even for $32$ threads, parallelization leads to faster task execution.
 
@@
 \section{Conclusion and Future Work}
+In this work, we describe our implementation of a thread pool system, present in detail the implementation and show our evaluation results containing a comparison against the baseline implementation and the waiting time behaviour over different thread pool sizes.
+Within our implementation, we state the differences of parallel memory access as well as the problem of implementing synchronous waiting for a set of tasks.
+
+As presented in the evaluation, our thread pool implementation reaches better execution performance than the baseline because of avoiding the creation of short living threads for each request.
+We measured the average time threads spent waiting in different thread pool sizes and present, that even for a size of $32$ threads, the waiting time stays constantly small.
+Consequently, we conclude that using \emph{first come first serve} as scheduling approach leads to small waiting times and high execution performance.
+
+To sum up, our thread pool implementation fulfils all performance and design requirements of geckoDB and BOLSTER and reaches even faster execution time than the existing implementation.
+In the future, we plan to make the following adaptations to our thread pool system.
+Firstly, we want to reduce the number of dynamic allocation, since heap allocations mostly are slower than stack allocations.
+Secondly, we want to examine the usage of existing priority queue implementations for our thread pool system.
+Thirdly, we plan to integrate a dynamic resizing function based on the collected statistics of the performance monitoring component.
 

### 2018-07-03 Presentations/Paper/paper_thread_pool.tex
Small paper fixes

@@
         being reused. To counter this effect we implemented a thread pool to be
-        used in combination with GeckoDB to reduce the additional time required
+        used in GeckoDB to reduce the additional time required
         to create and close thousands of threads. In this paper, we show our
-        implementation of a thread pool to process independent tasks in parallel
+        implementation of a thread pool to process independent tasks in parallel,
         including the waiting for a set of processed tasks. We also compare the
-        thread pool implementation against the current one. Additionally, we show that the task and thread pool configuration
-        depending on the use case has a high impact on the thread
+        thread pool implementation against the current one. Additionally, we show that the task and thread pool configuration,
+        depending on the use case, has a high impact on the thread
         pool performance. At the end of our evaluation, we show that the implementation 
-        fulfils all given constraints and generally reduce
+        fulfils all given requirements and generally reduce
         overhead compared to creating and spawning threads individually.
@@
 \section{Implementation}
-In this section, we describe in detail the design and implementation of our thread pool.
+In this section, we describe in detail the design and implementation of our thread pool system.
 We show how tasks are enqueued regarding their priority and how the waiting for tasks is implemented.
@@
 Since measuring performance metrics lead to memory and CPU overhead, we decided to exclude the performance measurements from the thread pool to make it optional.
-The performance monitoring can be activated as a boolean parameter to the thread pool create functions.
+The performance monitoring can be activated by a boolean parameter of the thread pool create functions.
 Consequently, in the target database system, the designers can decide for each thread pool instance, if performance monitoring should be applied.
@@
 One of the requirements of BOLSTER is waiting until a set of tasks is finished.
-To achieve the same performance as the baseline implementation reaches, we let the main thread first participate in the execution of the tasks and then wait until all tasks with the same handle are finished.
+To achieve at least the same performance as the baseline implementation reaches, we let the main thread first participate in the execution of the tasks and then wait until all tasks with the same handle are finished.
 
@@
 Before enqueueing, the function \emph{thread\_pool\_enqueue\_tasks\_wait} creates a handle and links all tasks to it.
-After enqueueing the tasks to the priority queue, the waiting threads in the thread pool pop the enqueued tasks from the priority queue and execute the routine of the task.
+After enqueueing the tasks to the priority queue, the waiting threads in the thread pool pop the enqueued tasks from the priority queue and execute the task routine.
+We decided to let threads look for new tasks instead of explicitly assigning to reduce the scheduling overhead.
+We call this approach \emph{first come first serve}.
 This happens asynchronously to the calling function and therefore is not considered as an own step of the process.
@@
 We implemented every thread state access with this function to ensure atomic thread state updates.
-\\
+
 This approach enables efficient dynamic and lazy resizing of the thread pool.
@@
 If later on the number is increased again, still running but marked threads can be enabled again ensuring that few threads have to be created and that the number of running threads will not temporarily exceed the requested size.
-\\
  After presenting the enqueueing, processing, and waiting for tasks, we show the second component of our thread pool system, the performance monitoring.
@@
 As last aspect of our implementation, we show how we test the thread pool and how we ensure to avoid memory leaks.
-Since geckoDB already uses GTest \cite{sen2010quick} for unit testing, we also wrote our test using this framework.
+Since GeckoDB already uses GTest \cite{sen2010quick} for unit testing, we also wrote our test using this framework.
 GTest provides a simple, powerful library to perform unit tests in C++.
@@
 \subsection{Evaluation Setup}
-In our evaluation setup, we focus primarily on the comparison of using our thread pool against the baseline implementation of creating new threads at runtime of BOLSTER.
+In our evaluation setup, we focus primarily on the comparison of our thread pool against the baseline implementation in BOLSTER, that creates new threads at runtime for each request.
 We simulate calls of BOLSTER primitives with different thread pool and task configurations and compare the results against the baseline.
@@
 Additionally, we evaluate the metrics for threads, tasks and the thread pool mentioned in Section 3.F regarding the thread pool and task configuration used in the BOLSTER primitives.
-To evaluate optimal configurations for the thread pool, we consider the relation between idle and busy time of the thread pool and the average waiting time of tasks.
+To evaluate optimal configurations for the thread pool, we consider the relation between waiting and completion time of  thread pool instances with different sizes.
 
@@
 ranging from 8 to 24 GB of both DDR3-SODIMM and DDR4-SODIMM. 
-
 All benchmarks have been repeated three times and a median value of all
@@
 \subsubsection{Benchmark Implementation}
-Since BOLSTER is deeply integrated into geckoDB and the workload and usage of multithreading depends on the graph data in the system, we can not directly use the BOLSTER implementation for our evaluation.
-Thus, we extracted the respecting code parts into an own implementation and use this for a direct comparison against the thread pool.
+Since BOLSTER is deeply integrated into GeckoDB and the workload and usage of multi threading depends on the graph data in the system, we can not directly use the BOLSTER implementation for our evaluation.
+Thus, we extracted the respecting code parts into an own implementation and use this for a direct comparison against one that uses our thread pool.
 
-In Figure \ref{fig4}, we show the approximated version of the BOLSTER multithreading code that we use as baseline implementation.
+In Figure \ref{fig4}, we show the approximated version of the BOLSTER multi threading code that we use as baseline implementation.
 In our benchmark.
 the method \emph{cmp\_baseline} is called multiple times with a different \emph{numThreads} parameter.
-Within each call, \emph{numThreads} threads are created and all of these execute the benchmark method v, that calculates a huge amount of exponential values in a loop.
+Within each call, \emph{numThreads} threads are created and all of these execute the benchmark method \emph{work\_large}, that calculates a huge amount of exponential values in a loop.
 After the calling thread takes part in execution, it waits until all created threads are joined.
-The elapsed processing time is measured from the begin of the method until joining all threads finished.
+The elapsed processing time is measured from the begin of the method until the join of all threads finished.
 
@@
 	\end{lstlisting}
-	\caption{Baseline multithreading implementation extracted from BOLSTER}
+	\caption{Baseline multi threading implementation extracted from BOLSTER}
 	\label{fig4}
@@
 Based on Figure \ref{fig6}, we make the following conclusions.
-
 Firstly, compared to strict processing workload, create, and later on join threads, consumes much time.
@@
 
-Secondly, the task creation and enqueuing have only slight on the performance, since the baseline does not include these operations, but is still clearly faster than the thread pool implementation.
+Secondly, the task creation and enqueuing have only slight on the performance, since the baseline does not include these operations, but is still clearly slower than the thread pool implementation.
 
@@
 \subsection{Summary}
-As last step of the evaluation, we summarize our results and compare them to the requirements of geckoDB.
+As last step of the evaluation, we summarize our results and compare them to the requirements of GeckoDB.
 As mentioned before, our implementation reaches better performance than the baseline regarding our specific tests.
@@
 
-Although a small and constant waiting time was not a requirement for integration into geckoDB, we ensure that assigning many tasks work fast for different thread pool sizes.
-Regarding the performance and also the design requirements of geckoDB and BOLSTER, we fulfil these requirements and successfully implemented and evaluated our thread pool implementation.
+Although a small and constant waiting time was not a requirement for integration into GeckoDB, we ensure that assigning many tasks works fast for different thread pool sizes.
+Regarding the performance and also the design requirements of GeckoDB and BOLSTER, we fulfil these requirements and successfully implemented and evaluated our thread pool implementation.
 
@@
 \section{Conclusion and Future Work}
-In this work, we describe our implementation of a thread pool system, present in detail the implementation and show our evaluation results containing a comparison against the baseline implementation and the waiting time behaviour over different thread pool sizes.
+In this work, we describe our implementation of a thread pool system.
+We present in detail the implementation and show our evaluation results containing a comparison against the baseline implementation and the waiting time behaviour over different thread pool sizes.
 Within our implementation, we state the differences of parallel memory access as well as the problem of implementing synchronous waiting for a set of tasks.
@@
 
-To sum up, our thread pool implementation fulfils all performance and design requirements of geckoDB and BOLSTER and reaches even faster execution time than the existing implementation.
+To sum up, our thread pool implementation fulfils all performance and design requirements of GeckoDB and BOLSTER and reaches even faster execution time than the existing implementation.
 In the future, we plan to make the following adaptations to our thread pool system.

### 2018-07-05 include/thread_pool.h
priority queue lifetime is correctly handled by the thread_pool

@@ -56,7 +56,7 @@ typedef struct task_handle_t{
 typedef struct thread_pool_t {
 	char* name;
 	pthread_t* pool;
-	priority_queue_t* waiting_tasks;
+	priority_queue_t waiting_tasks;
 	__task_state_t* task_group_states;
 	size_t task_state_capacity; // number of tasks that can be tracked
 	size_t size;

### 2018-07-05 src/thread_pool.c

@@ -17,7 +17,8 @@ thread_pool_t* thread_pool_create(size_t num_threads, int enable_monitoring) {
   pool->name = NULL;
   pool->size = num_threads;
   pool->capacity = num_threads * 2;
-  pool->waiting_tasks = calloc(1, sizeof(priority_queue_t));
+  priority_queue_init(&pool->waiting_tasks);
+
 
   pool->thread_tasks = calloc(num_threads, sizeof(thread_task_t*));
   pool->thread_infos = calloc(sizeof(__thread_information_t*) * pool->capacity, 1);
@@ -84,9 +85,8 @@ void thread_pool_free(thread_pool_t* pool) {
     }
       
 
-    priority_queue_free(pool->waiting_tasks);
+    priority_queue_free(&pool->waiting_tasks);
     free(pool->pool);
-    free(pool->waiting_tasks);
     free(pool->thread_tasks);
     free(pool->thread_infos);
     free(pool->task_group_states);
@@ -159,7 +159,7 @@ status_e thread_pool_enqueue_tasks(thread_task_t* tasks, thread_pool_t* pool, si
     }
     
     tasks[i].group_id = ind;
-    priority_queue_push(pool->waiting_tasks, &tasks[i], tasks[i].priority); 
+    priority_queue_push(&pool->waiting_tasks, &tasks[i], tasks[i].priority); 
   }
 
   if(hndl){
@@ -299,7 +299,7 @@ void *__thread_main(void* args) {
 }
 
 thread_task_t* __get_next_task(thread_pool_t *pool) {
-  thread_task_t* next_task = priority_queue_pop(pool->waiting_tasks);
+  thread_task_t* next_task = priority_queue_pop(&pool->waiting_tasks);
   return next_task;
 }
 

### 2018-07-05 Presentations/Paper/paper_thread_pool.tex
+ some optimizations + evaluation

@@
 Each assigned task contains a function pointer to a routine and a pointer to data.
-We implemented the task queue as a generic priority queue that uses a mutex thread to enqueue tasks threadsafe.
+We implemented the task queue as a generic priority queue utilizing an implicit heap.
+To ensure thread safety a mutex is used while operating on the queue.
 Every time a new task is enqueued to the task queue, two scenarios can occur.
@@
 
+With an increasing number of threads interactions with the queue could bottleneck the task throughput.
+An alternative lock free priority queue \cite{it:2018-003} based on a skip list is considered to improve on this.
 \subsection{Processing Tasks}
@@
 If the slot is reused, $genCount$ increases and the second part of the condition returns false.
+
+Since $openTaskCount$ has to be altered from different threads the required decrements are ensured to be atomic.
+This is in opposition to the usual goal of slot-maps to improve cache locality.
+Atomics on x86 usually prevent their whole cache line from being used while they are changed.
+Thus a sparse distribution of slots should reduces collisions.
+To quantify this difference the baseline implementation searches for free slots linearly from the beginning, while an alternative approach tries to distribute the slots over multiple cache lines:
+\begin{lstlisting}[style=CStyle]
+size_t ind = 0;
+// linear search, compact slot usage
+for(; ind < pool->task_state_capacity && pool->task_group_states[ind].task_count; ++ind);
+// distributed over different cache lines
+for(; pool->task_group_states[ind].task_count; ind = (ind + 8) % MAX_NUM_TASKS);
+\end{lstlisting}
 
@@
 
-In our setups mostly Sky- and Kabylake processors have been used with memory
+In our setups Sky- and Kabylake aswell as Ryzen processors have been used with memory
 ranging from 8 to 24 GB of both DDR3-SODIMM and DDR4-SODIMM. 
@@
 
-Secondly, the task creation and enqueuing have only slight on the performance, since the baseline does not include these operations, but is still clearly slower than the thread pool implementation.
+Secondly, the task creation and enqueuing have only slight impact on the performance, since the baseline does not include these operations, but is still clearly slower than the thread pool implementation.
 
@@
 To examine this behaviour with increasing thread pool size, we measure the average waiting time of thread pools with different sizes.
+
+To further test the performance of the task-completion system, in particular the previously mentioned wait-slot acquisition, the main thread periodicity has to wait for a specific task to finish and the total utilization is measured.
 
@@
 
+The selection strategy of wait-slots has only a small impact on occupation as seen in \ref{fig8}.
+The uniform distributions advantage amounts to $0.1\%$ across different numbers of threads and is far overshadowed by differences in the threads to tasks ratio.
+
+Figure \ref{fig9} shows the lock-free priority queue performing better.
+With a difference in utilization of $2\%$ when using only two threads the improvement over our mutex based implementation is not insignificant.
+A total occupation of over $99.5\%$ also demonstrates that most overhead in our implementation comes from the priority queue when supplying a balanced stream of tasks.
+With an increase in available threads the performance of the mutex solution further decreases, while the lock-free queue-variants utilization remains nearly constant.
+
 \begin{figure}
@@
 	\label{fig7}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.5\textwidth]{img/slot_distr.png}
+	\caption{Utilization \% with different slot distribution strategies}
+	\label{fig8}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.5\textwidth]{img/lock_free.png}
+	\caption{Comparison of Utilization \% of the lock-free queue versus mutex based heap}
+	\label{fig9}
 \end{figure}
@@
 Regarding the performance and also the design requirements of GeckoDB and BOLSTER, we fulfil these requirements and successfully implemented and evaluated our thread pool implementation.
+
+The choice of a different priority queue data structure presents an opportunity for further optimizations, in particular to scale well with large amounts of threads.
 

### 2018-07-06 Presentations/Paper/paper_thread_pool.tex
Add figure tag to slot search code

@@
 Thus a sparse distribution of slots should reduces collisions.
-To quantify this difference the baseline implementation searches for free slots linearly from the beginning, while an alternative approach tries to distribute the slots over multiple cache lines:
-\begin{lstlisting}[style=CStyle]
-size_t ind = 0;
-// linear search, compact slot usage
-for(; ind < pool->task_state_capacity && pool->task_group_states[ind].task_count; ++ind);
-// distributed over different cache lines
-for(; pool->task_group_states[ind].task_count; ind = (ind + 8) % MAX_NUM_TASKS);
-\end{lstlisting}
+To quantify this difference the baseline implementation searches for free slots linearly from the beginning, while an alternative approach tries to distribute the slots over multiple cache lines.
+In Figure \ref{fig10}, we show implementations for both search approaches.
 
@@
 We implemented every thread state access with this function to ensure atomic thread state updates.
+
+\begin{figure}
+	\begin{lstlisting}[style=CStyle]
+	size_t ind = 0;
+	// linear search, compact slot usage
+	for(; ind < pool->task_state_capacity && pool->task_group_states[ind].task_count; ++ind);
+	// distributed over different cache lines
+	for(; pool->task_group_states[ind].task_count; ind = (ind + 8) % MAX_NUM_TASKS);
+	\end{lstlisting}
+	\caption{Different implementations for slot searching}
+	\label{fig10}
+\end{figure}
 
@@
 
-To compare against the presented baseline implementation, we developed a method that does the same amount of calls of \emph{work\_large} as the baseline, but uses a thread pool instance instead of creating new threads.
-In Figure \ref{fig5}, we show this implementation called \emph{cmp\_pool} that creates and fills tasks and enqueues these into a thread pool instance.
-Since we assume, that the thread pools are reused for multiple requests, we exclude the thread pool creation and deallocation from the time measure.
-Consequently, only creating and processing tasks is included in the considered elapsed time.
-
 \begin{figure}
@@
 \end{figure}
+
+To compare against the presented baseline implementation, we developed a method that does the same amount of calls of \emph{work\_large} as the baseline, but uses a thread pool instance instead of creating new threads.
+In Figure \ref{fig5}, we show this implementation called \emph{cmp\_pool} that creates and fills tasks and enqueues these into a thread pool instance.
+Since we assume, that the thread pools are reused for multiple requests, we exclude the thread pool creation and deallocation from the time measure.
+Consequently, only creating and processing tasks is included in the considered elapsed time.
 
@@
 
-As mentioned before in this section, the performance results depend on the current workload of the system, as seen by the waiting time shown in Figure \ref{fig7}.
-Although the priority queue locking by the mutex thread should increase the waiting time slightly with bigger thread pool sizes, the last measuring point is lower than the previous points.
-
-The selection strategy of wait-slots has only a small impact on occupation as seen in \ref{fig8}.
-The uniform distributions advantage amounts to $0.1\%$ across different numbers of threads and is far overshadowed by differences in the threads to tasks ratio.
-
-Figure \ref{fig9} shows the lock-free priority queue performing better.
-With a difference in utilization of $2\%$ when using only two threads the improvement over our mutex based implementation is not insignificant.
-A total occupation of over $99.5\%$ also demonstrates that most overhead in our implementation comes from the priority queue when supplying a balanced stream of tasks.
-With an increase in available threads the performance of the mutex solution further decreases, while the lock-free queue-variants utilization remains nearly constant.
-
 \begin{figure}
@@
 \end{figure}
+
+As mentioned before in this section, the performance results depend on the current workload of the system, as seen by the waiting time shown in Figure \ref{fig7}.
+Although the priority queue locking by the mutex thread should increase the waiting time slightly with bigger thread pool sizes, the last measuring point is lower than the previous points.
+
+The selection strategy of wait-slots has only a small impact on occupation, as seen in Figure \ref{fig8}.
+The uniform distributions advantage amounts to $0.1\%$ across different numbers of threads and is far overshadowed by differences in the threads to tasks ratio.
+
+In Figure \ref{fig9}, we show that the lock-free priority queue performs better.
+With a difference in utilization of $2\%$ when using only two threads the improvement over our mutex based implementation is not insignificant.
+A total occupation of over $99.5\%$ also demonstrates that most overhead in our implementation comes from the priority queue when supplying a balanced stream of tasks.
+With an increase in available threads the performance of the mutex solution further decreases, while the lock-free queue-variants utilization remains nearly constant.
 

### 2018-07-06 Presentations/Paper/paper_thread_pool.tex
Fix typos and or grammar

@@
 (linux 4.13) in an virtual machine environment, second Archlinux (linux
-4.17.2-1) running native, both updated on stable branch. Tests have also been
+4.17.3-1) running native, both updated on stable branch. Tests have also been
 conducted on Windows 10 and macOs HighSierra machines.
@@
 \subsection{Validity of Results} 
-Before presenting our evaluation results, we give some preliminaries about the validity of our results.
-At first, all measurements are executed on all machines we listed in our experimental configuration, but in this work we present only the results measured on Archlinux with kernel 4.17.2-1 and 24 GB DDR4-SODIMM.
-We decided to use this system, since it is running on a high kernel version and not in a virtual environment.
-Additionally, we use gnuplot to generate diagrams from our results, consequently using a windows machine is not optimal.
+Before presenting our evaluation results, we
+give some preliminaries about the validity of our results. At first, all
+measurements are executed on all machines we listed in our experimental
+configuration, but in this work we present only the results measured on
+Archlinux with kernel 4.17.3-1 and 24 GB DDR4-SODIMM. We decided to use this
+system, since it is running a current kernel version and not in a virtual
+environment. Additionally, we use gnuplot to generate diagrams from our results,
+consequently using a windows machine is not optimal.
 

### 2018-07-12 Presentations/Milestone4/presentation.tex
Add presentation 4

--- 2018-06-13 Presentations/Milestone3/presentation.tex
@@
 \title{Thread Pool in GeckoDB/BOLSTER}
-\subtitle{Milestone III: Implementation \& evaluation setup}
+\subtitle{Milestone IV: Final presentation (wrap-up + evaluation results) }
 \author{Johann Wagner, Johannes Wünsche, Marten Wallewein-Eising, Robert Jenderise}
@@
 	\begin{center}
-		\includegraphics[width=0.9\textwidth]{img/pool_queue.png}
+		\includegraphics[width=0.9\textwidth]{img/pool_structure.png}
+		Thread Pool System
 	\end{center}
@@
 	\frametitle{Previously on Thread Pool in GeckoDB/BOLSTER}
-	\begin{columns}
-		\begin{column}{0.48\textwidth}
-			\begin{itemize}
-				\item Problem: waiting for a task to complete could lead to a
-				race condition
-				\item This was caused by the uncertainty of a thread\_task
-				during transition between the priority queue and the
-				thread pool
-				\item Solution: slotmap with an atomic counter
-			\end{itemize}
-		\end{column}
-		\begin{column}{0.48\textwidth}
-			\includegraphics[width=1.0\textwidth]{waitingconcept.png}
-		\end{column}
-	\end{columns}
+	\begin{center}
+		\includegraphics[width=0.9\textwidth]{img/pool_queue.png}
+		Enqueuing and Waiting for Tasks
+	\end{center}
+\end{frame}
+
+\begin{frame}
+	\frametitle{Previously on Thread Pool in GeckoDB/BOLSTER}
+	\begin{center}
+		\includegraphics[width=0.8\textwidth]{img/thread_pool_use.png}
+		Thread Pool Usage
+	\end{center}
 \end{frame}

### 2018-07-12 Presentations/Milestone4/presentation.tex

@@
 	\begin{center}
-		\includegraphics[width=0.9\textwidth]{img/pool_queue.png}
+		\includegraphics[width=0.8\textwidth]{img/pool_queue.png}
 		Enqueuing and Waiting for Tasks
@@
 	\begin{center}
-		\includegraphics[width=0.8\textwidth]{img/thread_pool_use.png}
+		\includegraphics[width=0.9\textwidth]{img/thread_pool_use.png}
 		Thread Pool Usage
@@
 
-\section{Implementation}
+\section{Evaluation}
 \begin{frame}
 	\begin{center}
-		\huge Implementation
-	\end{center}
-\end{frame}
-
-\begin{frame}
-	\frametitle{Design of Thread Pool System}
-	\begin{center}
-		\includegraphics[width=0.9\textwidth]{img/pool_structure.png}
-	\end{center}
-	\begin{itemize}
-		\item Thread pool system contains priority queue and thread pool itself
-		\item Performance monitoring as optional, external component
-	\end{itemize}
-\end{frame}
-
-\begin{frame}
-	\frametitle{Implementation}
-	We...
-	\begin{itemize}
-		\item Implemented task enqueueing and execution
-		\item Implemented performance monitoring as external component
-		\item Added gtest framework to project and wrote tests for pool and queue
-		\item Started writing performance tests and measures for comparison against baseline
-	\end{itemize}
-\end{frame}
-
-\begin{frame}
-	\frametitle{Usage Examples}
-	\begin{center}
-		\includegraphics[width=0.95\textwidth]{img/thread_pool_use.png}
-	\end{center}
-\end{frame}
-
-\begin{frame}
-\frametitle{Usage Examples}
-\begin{center}
-	\includegraphics[width=0.95\textwidth]{img/thread_pool_use2.png}
-\end{center}
-\end{frame}
-
-\begin{frame}
-	\frametitle{Open issues}
-	\begin{itemize}
-		\item GTest still has some problems with c atomics due to itself being implemented in C++
-		\item Some race condition may still occur
-		\item Check for memory leaks using valgrind
-	\end{itemize}
-\end{frame}
-
-\section{Evaluation Setup}
-\begin{frame}
-	\begin{center}
-		\huge Evaluation Setup
+		\huge Evaluation
 	\end{center}
@@
 	\begin{itemize}
-		\item Use of geckodb to stress test the thread pool by with BOLSTER primitives
-		\item Measure different statistics for tasks, thread and thread pools
-		\item Use statistics to determine optimal configurations for different thread pool instances
-		\item Time measures are taken with wall clock time
+		\item Tests run on different environments, including Windows 10, MacOS Sierra and Archlinux
+		\item Presentation results measured on Archlinux with kernel 4.17.2-1 and 24 GB DDR4-SODIMM
+		\item Each measure was repeated 3 times and an average value is used
 	\end{itemize}
@@
 \begin{frame}
-	\frametitle{Task Statistics}
-	\begin{center}
-		\begin{tabular}{ c c }
-			\hline
-			\textbf{Statistic name}&\textbf{Description}\\
-			\hline
-			enqueue\_time & Time when the task was enqueued \\
-			execution\_time & Time when the task execution begins \\
-			complete\_time & Time when the task execution is finished \\
-			\hline
-		\end{tabular}
-		\label{tab2}
-	\end{center}
+	\frametitle{Evaluation Benchmarks}
 	\begin{itemize}
-		\item Minimize waiting time $(execution\_time - enqueue\_time)$
-		\item Evaluate the task setup considering the whole processing time 
+		\item Compare against approximated implementation of BOLSTER (baseline)
+		\item Evaluation of waiting time for different thread pool sizes
+		\item Utilization of different slot distribution strategies
+		\item Utilization of lock-free vs mutex based heap
 	\end{itemize}
@@
 \begin{frame}
-	\frametitle{Thread Statistics}
+	\frametitle{Comparison against Baseline Implementation}
 	\begin{center}
-		\begin{tabular}{ c c }
-			\hline
-			\textbf{Statistic name}&\textbf{Description}\\
-			\hline
-			waiting\_time & Time the thread spend waiting (ns) \\
-			busy\_time & Time the thread spend executing tasks (ns)\\
-			task\_count & Number of tasks the thread has executed \\
-			\hline
-		\end{tabular}
-		\label{tab3}
+		\includegraphics[width=0.8\textwidth]{img/pool_baseline.png}
 	\end{center}
-	\begin{itemize}
-		\item $waiting\_time$ of threads indicate a good thread pool size
-		\item $task\_count$ interesting to number of all completed tasks
-	\end{itemize}
+
 \end{frame}
@@
 \begin{frame}
-	\frametitle{Thread Pool Statistics}
+	\frametitle{Waiting Time Evaluation}
 	\begin{center}
-		\begin{tabular}{ c c }
-			\hline
-			\textbf{Statistic name}&\textbf{Description}\\
-			\hline
-			working\_time & Time the thread pool runs (ns) \\
-			task\_complete\_count & Number of completed tasks \\
-			task\_enqueued\_count & Number of enqueued tasks \\
-			avg\_complete\_time & Average time span from enqueueing until \\
-			& execution of tasks \\
-			avg\_wait\_time & Average time of tasks spend in the queue \\
-			\hline
-		\end{tabular}
-		\label{tab4}
+		\includegraphics[width=0.8\textwidth]{img/pool_avg.png}
 	\end{center}
-	\begin{itemize}
-		\item $avg\_wait\_time$ indicates the quality of the chosen thread pool size
-	\end{itemize}
+
 \end{frame}
@@
 \begin{frame}
-	\frametitle{Possible Optimization Opportunities}
-	\begin{itemize}
-		\item Improve Priority Queue data structure
-		\item Change distribution of Wait-Slots
-		\item Reduce number of dynamic allocations
-		\item Adjust number of threads to increase occupancy
-	\end{itemize}
-\end{frame}
+	\frametitle{Utilization of different slot distribution strategies}
+	\begin{center}
+		\includegraphics[width=0.8\textwidth]{img/slot_distr.png}
+	\end{center}
 
-\section{Paper Progress}
-\begin{frame}
-	\begin{center}
-	\huge Paper Progress \normalsize	
-	\end{center}
 \end{frame}
@@
 \begin{frame}
-	\begin{itemize}
-		\item Introduction and Preliminaries completed
-		\item Implementation described, including our used concepts and basic ideas how to build and operate the pool
-		\item Still required are the evaluation (since we still need to
-			measure the actual performance), the description of our testing environment and our final conclusion
-	\end{itemize}
+	\frametitle{Utilization of lock-free vs mutex based heap}
+	\begin{center}
+		\includegraphics[width=0.8\textwidth]{img/lock_free.png}
+	\end{center}
+
 \end{frame}

### 2018-07-12 Presentations/Milestone4/presentation.tex
added description to pool_avg plot
Add Summary to presentation 4

@@
 		\includegraphics[width=0.9\textwidth]{img/pool_structure.png}
+	\end{center}
+	\begin{center}
 		Thread Pool System
@@
 		\includegraphics[width=0.8\textwidth]{img/pool_queue.png}
+	\end{center}
+	\begin{center}
 		Enqueuing and Waiting for Tasks
@@
 		\includegraphics[width=0.9\textwidth]{img/thread_pool_use.png}
+
+	\end{center}
+	\begin{center}
 		Thread Pool Usage
@@
 
+\section{Summary}
+\begin{frame}
+	\begin{center}
+		\huge Summary
+	\end{center}
+\end{frame}
+
+\begin{frame}
+	\frametitle{Summary}
+	\begin{itemize}
+		\item Better execution performance as baseline implementation
+		\item Constant average waiting time over different pool sizes
+		\item Small performance costs for enqueueing and waiting
+	\end{itemize}
+\end{frame}
+
+\section{Conclusion}
+	\begin{frame}
+	\begin{center}
+		\huge Conclusion
+	\end{center}
+\end{frame}
+
 \begin{frame}

### 2018-07-12 Presentations/Milestone4/presentation.tex
Add some more summary to presentation 4

@@
 	\begin{itemize}
-		\item Tests run on different environments, including Windows 10, MacOS Sierra and Archlinux
+		\item Tests run on different environments, including Windows 10, MacOS High Sierra and Archlinux
 		\item Presentation results measured on Archlinux with kernel 4.17.2-1 and 24 GB DDR4-SODIMM
@@
 	\end{center}
-
+	\begin{tabular}{c|c|c}
+		\#Threads & distributed & compact \\
+		1 & 0.998724 & 0.998787 \\
+		2 & 0.935569 & 0.935331 \\
+		3 & 0.917452 & 0.917384 \\
+		4 & 0.918376 & 0.915736 \\
+		5 & 0.795493 & 0.792239 \\
+		6 & 0.906671 & 0.905388 \\
+		7 & 0.818336 & 0.817147 \\
+		8 & 0.774444 & 0.77419		
+	\end{tabular}
 \end{frame}
@@
 \begin{frame}
-	\frametitle{Utilization of lock-free vs mutex based heap}
+	\frametitle{Utilization of lock-free vs mutex based priority queue}
 	\begin{center}
@@
 	\end{center}
-
 \end{frame}
@@
 	\frametitle{Summary}
+	\textbf{Evaluation results:}
 	\begin{itemize}
@@
 
-\section{Conclusion}
 \begin{frame}
-	\begin{center}
-		\huge Conclusion
-	\end{center}
+\frametitle{Summary}
+	\textbf{Project Evaluation:}
+	\begin{itemize}
+		\item All requirements of the thread pool fulfilled
+		\item Compact but appropriate implementation feedback
+		\item Literature research problematic 
+		\begin{itemize}
+			\item Few paper about thread pool implementations
+			\item Focus on resizing
+		\end{itemize}
+	\end{itemize}
 \end{frame}
+
+%\section{Conclusion}
+%	\begin{frame}
+%	\begin{center}
+%		\huge Conclusion
+%	\end{center}
+%\end{frame}
 

### 2018-07-27 Presentations/Paper/paper_thread_pool.tex
Preparation for technical report

@@
 \usepackage{listings}
+\usepackage{url}
 
@@
 
-\title{Thread Pool in GeckoDB/BOLSTER - Implementing a Thread Pool for a High Performance Database System \\}
+\title{Thread Management for High Performance Database Systems - Design and Implementation \\{\large Technical Report}\\}
 
 \author{
-	\IEEEauthorblockN{Robert Jendersie, Johannes Wuensche, Johann Wagner, Marten Wallewein-Eising}
-	\IEEEauthorblockA{\textit{Otto-von-Guericke University} \\
+	\IEEEauthorblockN{Robert Jendersie, Johannes Wuensche, Johann Wagner, \\Marten Wallewein-Eising, Marcus Pinnecke, Gunter Saake}
+	\IEEEauthorblockA{\textit{Database and Software Engineering Group, Otto-von-Guericke University} \\
 		Magdeburg, Germany \\
-		\emph{firstname}.\emph{lastname}@st.ovgu.de} \and
+		\emph{firstname}.\emph{lastname}@ovgu.de} \and
 }
@@
 Additionally, we measure metrics like \emph{waiting and busy time} of threads to evaluate correct thread pool sizes for the considered use cases.
+
 In this work we make the following contributions:
 \begin{itemize}
-	\item We describe our design and implementation of the thread pool
-	\item We evaluate the possibility to wait for a group of task in the calling thread
-	\item We compare our thread pool against the existing implementation in BOLSTER
-	\item We measure and evaluate additional statistics for different thread pool configurations
+	\item \textbf{Design and Implementation} We describe our design and implementation of the thread pool
+	\item \textbf{Waiting Strategies} We evaluate the possibility to wait for a group of task in the calling thread
+	\item \textbf{Evaluation} We compare our thread pool against the existing implementation in BOLSTER
+	\item \textbf{Statistics per Configuration} We measure and evaluate additional statistics for different thread pool configurations
 \end{itemize}
@@
 
+\section{Acknowledgements}
+This work was partially funded by the DFG (grant no.: SA 465/50-1).
+
 \section{Conclusion and Future Work}

### 2018-08-17 Presentations/Paper/paper_thread_pool.tex
Feedback implemented as given by Gunter

@@
 Parallel data processing can be achieved by different approaches, like instruction and data parallelism or multi threading.
-In this paper, we focus on multi threading by implementing a thread pool for the graph database system GeckoDB.
+In this paper, we focus on multi threading by implementing a thread pool for the graph database system GeckoDB\footnote{GeckoDB source code repository: \url{https://github.com/geckodb}}.
 The thread pool will be integrated into BOLSTER, a high performance library for parallel execution of primitives like for or filter on large data sets.
+BOLSTER has similarities to Intel TBB~\cite{books/daglib/0018624} but is handwritten from scratch to fit perfectly into GeckoDBs storage engine written in C11~\cite{pinneckegridtables}.
 In the current implementation, BOLSTER creates a fix number of threads for each call of a primitive.
@@
 We organized the rest of the paper as follows.
-In Section 2, we give preliminaries about the considered task configuration and about thread safe access of memory.
-In Section 3, we show our design and implementation of the thread pool and examine our experimental environment in Section 4.
-In Section 5, we describe the results of our performance evaluation.
-In Section 6, we name related work and state our conclusion and future work in Section 7.
+In Section \ref{sec:pre}, we give preliminaries about the considered task configuration and about thread safe access of memory.
+In Section III, we show our design and implementation of the thread pool and examine our experimental environment in Section IV.
+In Section \ref{sec:eval}, we describe the results of our performance evaluation.
+In Section \ref{sec:related}, we name related work and state our conclusion and future work in Section \ref{sec:future}.
 
-\section{Preliminaries}
+\section{Preliminaries}\label{sec:pre}
 In this section, we define our configuration of tasks that are processed by the thread pool and state difficulties of synchronizing thread access to memory.
@@
 For example, while one thread updates data at a specific memory location, another thread writes into the data, which results to inconsistency.
-In Section 2, we presented two approaches to handle multithreaded access to the same memory location, mutex threads and atomic operations and we state where those approaches are used in our implementation.
+In Section \ref{sec:pre}, we presented two approaches to handle multithreaded access to the same memory location, mutex threads and atomic operations and we state where those approaches are used in our implementation.
 In this subsection, we focus on atomic operations to manage thread states.
@@
 
-\subsection{Performance monitoring}
+\subsection{Performance monitoring}\label{subsec:monitor}
 The performance monitoring component of the thread pool system is responsible for collecting data about threads, tasks, and the thread pool instance.
@@
 We wrote unit tests for the priority queue, the thread pool itself and the performance monitoring.
-As mentioned before, we use functions from the stdatomic library of c.
-Using this library in GTest results in compiler errors with GCC, since the c atomic implementation differs from the c++ one of the standard library and the GCC developers will not fix this issue.
-Consequently, we wrote a workaround not using the c stdatomic library in the GTest compilation unit.
+As mentioned before, we use functions from the stdatomic library of C.
+Using this library in GTest results in compiler errors with GCC, since the C atomic implementation differs from the C++ one of the standard library and the GCC developers will not fix this issue.
+Consequently, we wrote a workaround not using the C stdatomic library in the GTest compilation unit.
 
-Since c as low level language does not have an own memory management, memory leaks may occur using heap space allocations.
+Since C as low level language does not have an own memory management, memory leaks may occur using heap space allocations.
 To find and remove all memory leaks in our implementation, we use the tool \emph{valgrind}, which offers different functionalities to find race and jump condition and to count unfreed heap blocks.
@@
 
-Additionally, we evaluate the metrics for threads, tasks and the thread pool mentioned in Section 3.F regarding the thread pool and task configuration used in the BOLSTER primitives.
+Additionally, we evaluate the metrics for threads, tasks and the thread pool mentioned in Section \ref{subsec:monitor} regarding the thread pool and task configuration used in the BOLSTER primitives.
 To evaluate optimal configurations for the thread pool, we consider the relation between waiting and completion time of  thread pool instances with different sizes.
@@
 
-\section{Evaluation}
-In this Section, we present the results of our evaluation according to the setup we stated in the previous section.
+\section{Evaluation}\label{sec:eval}
+In this section, we present the results of our evaluation according to the setup we stated in the previous section.
 We start with an introduction of the benchmark methods to compare our thread pool against the baseline implementation and present the results.
@@
 
-\section{Related Work}
+\section{Related Work}\label{sec:related}
 Implementing multi threading can be realised using different approaches from spawning new threads to reuse them in thread pools.
@@
 
-\section{Acknowledgements}
-This work was partially funded by the DFG (grant no.: SA 465/50-1).
-
-\section{Conclusion and Future Work}
+\section{Conclusion and Future Work}\label{sec:future}
 In this work, we describe our implementation of a thread pool system.
@@
 
+\section{Acknowledgements}
+This work was partially funded by the DFG (grant no.: SA 465/50-1).
+
 \bibliographystyle{splncs}

### 2018-08-17 Presentations/Paper/paper_thread_pool.tex
More references and text adaption

@@
 \usepackage{xcolor}
+\usepackage{balance}
 \usepackage{listings}
@@
 \definecolor{mPurple}{rgb}{0.58,0,0.82}
-\definecolor{backgroundColour}{rgb}{0.95,0.95,0.92}
+\definecolor{backgroundColour}{rgb}{1,1,1}
 
@@
 	keywordstyle=\color{blue},
-	numberstyle=\tiny\color{mGray},
+	numberstyle=\tiny\color{mGray}\fontsize{4}{4}\selectfont\ttfamily,
 	stringstyle=\color{mPurple},
@@
 	language=C,
+	basicstyle=\fontsize{7}{7}\selectfont\ttfamily,
 	morekeywords={*,thread\_pool\_t, thread\_task\_t}
@@
         up with the requirements of modern software systems. However, the
-        current implementation of parallel processing in GeckoDB requires
-        spawning many short-lived threads that execute a single task and then
+        current implementation of parallel processing in GeckoDB, a graph database system developed in our group, requires
+        to spawn many short-lived threads that execute a single task and then
         terminate. This creates a lot of overhead since the threads are not
@@
 \begin{IEEEkeywords}
-Thread Pool, GeckoDB, BOLSTER
+Database Management System, Architecture of Parallel Processing Systems, Thread and Task Management
 \end{IEEEkeywords}
@@
 \section{Introduction}
-Since the amount of data that is stored and processed by modern database systems is growing fast, sequential data processing as only possibility is inconceivable.
-Applications have to process data in parallel to reach sufficient throughput to fulfil appropriate requirements.
+Since the amount of data that is stored and processed by modern database systems is growing fast, sequential data processing as only possibility is inconceivable~\cite{Alonso:2013:HKS:2510649.2511162}.
+Hence, applications and systems have to process data in parallel to reach sufficient throughput to fulfil appropriate requirements~\cite{DBLP:series/synthesis/2017Ailamaki, 10.1007/978-3-319-72413-3_8, 10.5120/ijca2016912260} which in turn affects the architectural design and query processing of modern database systems~\cite{agrawal2017many, Makreshanski2017, schuh2016experimental, psaroudakis2016adaptive, psaroudakis2013task}.
+
+The growing number of cores for both traditional CPUs as well as trending co-processors like GPUs require to re-evaluate core design descicion in data-intense systems.
+For instance, Wu et al.
+examined the scalability of multi-version concurrency control (MVCC) for a growning number of threads in  2017\cite{wu2017empirical}.
+They concluded that multi-threading in particular and modern hardware in general both promise notable performance gains for high-performance database systems using MVCC.
+However, they also observed that there is no clear winner combination for several concurrency control protocol, version storage, and garbage collection as well as the index management strategies in MVCC.
+
+For this paper, we focus on managing a far more lower level: general parallel data processing.
+In particular, with this paper we provide insights into our thread managing strategy in GeckoDB/BOOLSTER, its design and its implementation.
 
 Parallel data processing can be achieved by different approaches, like instruction and data parallelism or multi threading.
-In this paper, we focus on multi threading by implementing a thread pool for the graph database system GeckoDB\footnote{GeckoDB source code repository: \url{https://github.com/geckodb}}.
+We focus on multi threading by implementing a thread pool for the graph database system GeckoDB\footnote{GeckoDB source code repository: \url{https://github.com/geckodb}}.
 The thread pool will be integrated into BOLSTER, a high performance library for parallel execution of primitives like for or filter on large data sets.
-BOLSTER has similarities to Intel TBB~\cite{books/daglib/0018624} but is handwritten from scratch to fit perfectly into GeckoDBs storage engine written in C11~\cite{pinneckegridtables}.
+BOLSTER has similarities to Intel TBB~\cite{books/daglib/0018624}, a C++ template library for parallel programming, but BOLSTER is written from scratch to perfectly fit into GeckoDBs specialized storage engine and vectorized query engine, both written in C11~\cite{pinneckegridtables}.
+
 In the current implementation, BOLSTER creates a fix number of threads for each call of a primitive.
@@
 Since many primitives are executed at the same time, a couple of drawbacks arise from this implementation.
-First of all, the creation of threads comes along with overhead like stack initialisation and memory allocation.
+First of all, the creation of threads comes along with overhead like stack initialization and memory allocation.
 Secondly, creating a huge number of threads simultaneously may lead to large context switch overhead of the scheduler.
-Additionally, debugging and profiling applications that create many threads during runtime is 	time-consuming.
+Additionally, debugging and profiling applications that create many threads during runtime is time consuming.
 
-To overcome these drawbacks, we integrate an optimised thread pool in BOLSTER.
+To overcome these drawbacks, we integrate an optimized thread pool in BOLSTER.
 Along with the implementation, we measure the performance of the primitives to determine the thread pool overhead.
@@
 \begin{figure*}[htbp]
-	\includegraphics[width=1.0\textwidth]{img/pool_structure.png}
+	\centering \includegraphics[width=0.7\textwidth]{img/pool_structure.png}
 	\caption{Design of the thread pool system}
@@
 \begin{figure}
-	\includegraphics[width=0.5\textwidth]{img/pool_queue.png}
+	\centering\includegraphics[width=0.4\textwidth]{img/pool_queue.png}
 	\caption{Workflow of task enqueueing and execution in the thread pool}
@@
 \begin{figure}
-	\includegraphics[width=0.4\textwidth]{img/waitingconcept.png}
+\centering	\includegraphics[width=0.25\textwidth]{img/waitingconcept.png}
 	\caption{Concept of waiting for tasks using a slot-map}
@@
 \begin{figure}
-	\begin{lstlisting}[style=CStyle]
+	\begin{lstlisting}[style=CStyle,numbers=none]
 	size_t ind = 0;
@@
 \begin{figure}
-	\begin{lstlisting}[style=CStyle]
+	\begin{lstlisting}[style=CStyle,numbers=none]
 		void cmp_baseline(int numThreads)
@@
 \begin{figure}
-	\begin{lstlisting}[style=CStyle]
+	\begin{lstlisting}[style=CStyle,numbers=none]
 		void cmp_pool(int numThreads)
@@
 \begin{figure}
-	\includegraphics[width=0.5\textwidth]{img/slot_distr.png}
+	\centering\includegraphics[width=0.4\textwidth]{img/slot_distr.png}
 	\caption{Utilization \% with different slot distribution strategies}
@@
 \begin{figure}
-	\includegraphics[width=0.5\textwidth]{img/lock_free.png}
+	\centering\includegraphics[width=0.4\textwidth]{img/lock_free.png}
 	\caption{Comparison of Utilization \% of the lock-free queue versus mutex based heap}
@@
 \section{Acknowledgements}
-This work was partially funded by the DFG (grant no.: SA 465/50-1).
+This work was partially funded by the DFG (grant no.: SA 465/50-1). Also we would like to thank Sebastian Bre\ss, Thomas Leich, David Broneske, and Gabriel Campero Durand.
 
 \bibliographystyle{splncs}
+\balance
 \bibliography{paper_thread_pool}

### 2018-08-17 Presentations/Paper/paper_thread_pool.tex
Grammar and typos eliminated

@@
         terminate. This creates a lot of overhead since the threads are not
-        being reused. To counter this effect we implemented a thread pool to be
+        being reused. To counter this effect, we implemented a thread pool to be
         used in GeckoDB to reduce the additional time required
@@
 
-The growing number of cores for both traditional CPUs as well as trending co-processors like GPUs require to re-evaluate core design descicion in data-intense systems.
+The growing number of cores for both traditional CPUs as well as trending co-processors like GPUs require to re-evaluate core design decision in data-intense systems.
 For instance, Wu et al.
-examined the scalability of multi-version concurrency control (MVCC) for a growning number of threads in  2017\cite{wu2017empirical}.
+examined the scalability of multi-version concurrency control (MVCC) for a growing number of threads in  2017\cite{wu2017empirical}.
 They concluded that multi-threading in particular and modern hardware in general both promise notable performance gains for high-performance database systems using MVCC.
@@
 
-For this paper, we focus on managing a far more lower level: general parallel data processing.
+For this paper, we focus on managing a far lower level: general parallel data processing.
 In particular, with this paper we provide insights into our thread managing strategy in GeckoDB/BOOLSTER, its design and its implementation.
@@
 The higher the priority of a task is, the further it is placed behind in the queue.
-Additionally, we only consider not preemtable tasks.
+Additionally, we only consider non-preemptable tasks.
 Once a task is assigned to a thread, the thread will finish the operation of the task before getting a new one.
@@
 We use both approaches for different parts of the thread pool.
-Since using a mutex thread can decrease access performance, we decided to execute atomic operations on memory containing the thread state informations.
+Since using a mutex thread can decrease access performance, we decided to execute atomic operations on memory containing the thread state information.
 We use a mutex thread to synchronize the enqueueing of tasks into the priority queue.
@@
 Atomics on x86 usually prevent their whole cache line from being used while they are changed.
-Thus a sparse distribution of slots should reduces collisions.
-To quantify this difference the baseline implementation searches for free slots linearly from the beginning, while an alternative approach tries to distribute the slots over multiple cache lines.
+Thus a sparse distribution of slots should reduce collisions.
+To quantify this difference, the baseline implementation searches for free slots linearly from the beginning, while an alternative approach tries to distribute the slots over multiple cache lines.
 In Figure \ref{fig10}, we show implementations for both search approaches.
@@
 \subsection{Managing Thread States}
-As mentioned before, accessing the same memory location from different threads may lead to race conditions, which means the program behaviour depends on the accessing order of the threads.
+As mentioned before, accessing the same memory location from different threads may lead to race conditions, which means the program behavior depends on the accessing order of the threads.
 For example, while one thread updates data at a specific memory location, another thread writes into the data, which results to inconsistency.
@@
 			waiting\_time & Time the thread spend waiting (ms) \\
-			busy\_time & Time the thread spend executing tasks (ms)\\
+			busy\_time & Time the thread spends executing tasks (ms)\\
 			task\_count & Number of tasks the thread has executed \\
@@
 
-For testing purposes we used primarly Linux setups, for one Ubuntu 17.04
-(linux 4.13) in an virtual machine environment, second Archlinux (linux
+For testing purposes we used primarily Linux setups, for one Ubuntu 17.04
+(linux 4.13) in a virtual machine environment, second Arch Linux (linux
 4.17.3-1) running native, both updated on stable branch. Tests have also been
-conducted on Windows 10 and macOs HighSierra machines.
+conducted on Windows 10 and macOS High Sierra machines.
 
-In our setups Sky- and Kabylake aswell as Ryzen processors have been used with memory
+In our setups Sky- and Kabylake as well as Ryzen processors have been used with memory
 ranging from 8 to 24 GB of both DDR3-SODIMM and DDR4-SODIMM. 
 All benchmarks have been repeated three times and a median value of all
-repetitions is being used for evaluation.
+repetitions are being used for evaluation.
 
@@
 configuration, but in this work we present only the results measured on
-Archlinux with kernel 4.17.3-1 and 24 GB DDR4-SODIMM. We decided to use this
+Arch Linux with kernel 4.17.3-1 and 24 GB DDR4-SODIMM. We decided to use this
 system, since it is running a current kernel version and not in a virtual
 environment. Additionally, we use gnuplot to generate diagrams from our results,
-consequently using a windows machine is not optimal.
+consequently, using a windows machine is not optimal.
 
@@
 Consequently, in large thread pools, the accessing time of the priority queue may increase along with the waiting time of threads.
-To examine this behaviour with increasing thread pool size, we measure the average waiting time of thread pools with different sizes.
+To examine this behavior with increasing thread pool size, we measure the average waiting time of thread pools with different sizes.
 
@@
 \subsubsection{Benchmark Implementation}
-To examine the waiting time behaviour of larger thread pools, we use a similar implementation than the \emph{cmp\_pool} method, which creates 10000 tasks at different thread pool sizes.
+To examine the waiting time behavior of larger thread pools, we use a similar implementation than the \emph{cmp\_pool} method, which creates 10000 tasks at different thread pool sizes.
 Again, we consider sizes from $4$ to $32$ threads as useful test case and take the average value of three measurements.
@@
 \section{Related Work}\label{sec:related}
-Implementing multi threading can be realised using different approaches from spawning new threads to reuse them in thread pools.
+Implementing multi threading can be realized using different approaches from spawning new threads to reuse them in thread pools.
 Pyarali et al.
@@
 In this work, we describe our implementation of a thread pool system.
-We present in detail the implementation and show our evaluation results containing a comparison against the baseline implementation and the waiting time behaviour over different thread pool sizes.
+We present in detail the implementation and show our evaluation results containing a comparison against the baseline implementation and the waiting time behavior over different thread pool sizes.
 Within our implementation, we state the differences of parallel memory access as well as the problem of implementing synchronous waiting for a set of tasks.

### 2018-08-17 Presentations/Paper/paper_thread_pool.tex
Polish

@@
         up with the requirements of modern software systems. However, the
-        current implementation of parallel processing in GeckoDB, a graph database system developed in our group, requires
+        current implementation of parallel processing in \textsc{GeckoDB}, a graph database system developed in our group, requires
         to spawn many short-lived threads that execute a single task and then
@@
         being reused. To counter this effect, we implemented a thread pool to be
-        used in GeckoDB to reduce the additional time required
+        used in \textsc{GeckoDB} to reduce the additional time required
         to create and close thousands of threads. In this paper, we show our
@@
 For this paper, we focus on managing a far lower level: general parallel data processing.
-In particular, with this paper we provide insights into our thread managing strategy in GeckoDB/BOOLSTER, its design and its implementation.
+In particular, with this paper we provide insights into our thread managing strategy in \textsc{GeckoDB}/\textsc{Bolster}, its design and its implementation.
 
 Parallel data processing can be achieved by different approaches, like instruction and data parallelism or multi threading.
-We focus on multi threading by implementing a thread pool for the graph database system GeckoDB\footnote{GeckoDB source code repository: \url{https://github.com/geckodb}}.
-The thread pool will be integrated into BOLSTER, a high performance library for parallel execution of primitives like for or filter on large data sets.
-BOLSTER has similarities to Intel TBB~\cite{books/daglib/0018624}, a C++ template library for parallel programming, but BOLSTER is written from scratch to perfectly fit into GeckoDBs specialized storage engine and vectorized query engine, both written in C11~\cite{pinneckegridtables}.
+We focus on multi threading by implementing a thread pool for the graph database system \textsc{GeckoDB}\footnote{\textsc{GeckoDB} source code repository: \url{https://github.com/\textsc{GeckoDB}}}.
+The thread pool will be integrated into \textsc{Bolster}, a high performance library for parallel execution of primitives like for or filter on large data sets.
+\textsc{Bolster} has similarities to Intel \textsc{Tbb}~\cite{books/daglib/0018624}, a \textsc{C++} template library for parallel programming, but \textsc{Bolster} is written from scratch to perfectly fit into \textsc{GeckoDB}s specialized storage engine and vectorized query engine, both written in C11~\cite{pinneckegridtables}.
 
-In the current implementation, BOLSTER creates a fix number of threads for each call of a primitive.
+In the current implementation, \textsc{Bolster} creates a fix number of threads for each call of a primitive.
 This approach is called \emph{thread-per-request}.
@@
 
-To overcome these drawbacks, we integrate an optimized thread pool in BOLSTER.
+To overcome these drawbacks, we integrate an optimized thread pool in \textsc{Bolster}.
 Along with the implementation, we measure the performance of the primitives to determine the thread pool overhead.
@@
 	\item \textbf{Waiting Strategies} We evaluate the possibility to wait for a group of task in the calling thread
-	\item \textbf{Evaluation} We compare our thread pool against the existing implementation in BOLSTER
+	\item \textbf{Evaluation} We compare our thread pool against the existing implementation in \textsc{Bolster}
 	\item \textbf{Statistics per Configuration} We measure and evaluate additional statistics for different thread pool configurations
@@
 With higher generalization, the implementation can be used for more use cases, but the execution performance may decrease.
-Therefore, we provide a set of functions to process task in the thread pool to match different requirements of BOLSTER.
+Therefore, we provide a set of functions to process task in the thread pool to match different requirements of \textsc{Bolster}.
 In Table 1, we show the functions provided by our thread pool to process tasks.
-One of the requirements of BOLSTER is waiting until a set of tasks is finished.
+One of the requirements of \textsc{Bolster} is waiting until a set of tasks is finished.
 To achieve at least the same performance as the baseline implementation reaches, we let the main thread first participate in the execution of the tasks and then wait until all tasks with the same handle are finished.
@@
 Consequently, another thread can update the state between check of the current state and update.
-To avoid this, we use the function \emph{atomic\_compare\_exchange\_strong} \cite{atomicstrong} of the stdatomic library.
+To avoid this, we use the function \emph{atomic\_compare\_exchange\_strong} \cite{atomicstrong} of the \texttt{stdatomic} library.
 This function combines the check and update in one atomic procedure.
@@
 As last aspect of our implementation, we show how we test the thread pool and how we ensure to avoid memory leaks.
-Since GeckoDB already uses GTest \cite{sen2010quick} for unit testing, we also wrote our test using this framework.
-GTest provides a simple, powerful library to perform unit tests in C++.
+Since \textsc{GeckoDB} already uses \textsc{GTest} \cite{sen2010quick} for unit testing, we also wrote our test using this framework.
+\textsc{GTest} provides a simple, powerful library to perform unit tests in \textsc{C++}.
 We wrote unit tests for the priority queue, the thread pool itself and the performance monitoring.
-As mentioned before, we use functions from the stdatomic library of C.
-Using this library in GTest results in compiler errors with GCC, since the C atomic implementation differs from the C++ one of the standard library and the GCC developers will not fix this issue.
-Consequently, we wrote a workaround not using the C stdatomic library in the GTest compilation unit.
+As mentioned before, we use functions from the \texttt{stdatomic} library of C.
+Using this library in \textsc{GTest} results in compiler errors with \textsc{Gcc}, since the C atomic implementation differs from the \textsc{C++} one of the standard library and the \textsc{Gcc} developers will not fix this issue.
+Consequently, we wrote a workaround not using the C \texttt{stdatomic} library in the \textsc{GTest} compilation unit.
 
-Since C as low level language does not have an own memory management, memory leaks may occur using heap space allocations.
-To find and remove all memory leaks in our implementation, we use the tool \emph{valgrind}, which offers different functionalities to find race and jump condition and to count unfreed heap blocks.
+Since C as low level language does not have an own automatic memory management, memory leaks may occur using heap space allocations.
+To find and remove all memory leaks in our implementation, we use the tool \textsc{Valgrind}, which offers different functionalities to find race and jump condition and to count unfreed heap blocks.
 We found memory leaks in our priority queue tests and also in the deallocation of thread pool instances.
@@
 \subsection{Evaluation Setup}
-In our evaluation setup, we focus primarily on the comparison of our thread pool against the baseline implementation in BOLSTER, that creates new threads at runtime for each request.
-We simulate calls of BOLSTER primitives with different thread pool and task configurations and compare the results against the baseline.
+In our evaluation setup, we focus primarily on the comparison of our thread pool against the baseline implementation in \textsc{Bolster}, that creates new threads at runtime for each request.
+We simulate calls of \textsc{Bolster} primitives with different thread pool and task configurations and compare the results against the baseline.
 
-Additionally, we evaluate the metrics for threads, tasks and the thread pool mentioned in Section \ref{subsec:monitor} regarding the thread pool and task configuration used in the BOLSTER primitives.
+Additionally, we evaluate the metrics for threads, tasks and the thread pool mentioned in Section \ref{subsec:monitor} regarding the thread pool and task configuration used in the \textsc{Bolster} primitives.
 To evaluate optimal configurations for the thread pool, we consider the relation between waiting and completion time of thread pool instances with different sizes.
 
-%Furthermore, we compare the scope and usage of our thread pool against an existing implementation \emph{Threading Building Blocks (Intel TBB)} \cite{kim2011multicore} written in C++.
+%Furthermore, we compare the scope and usage of our thread pool against an existing implementation \emph{Threading Building Blocks (Intel TBB)} \cite{kim2011multicore} written in \textsc{C++}.
 Since Intel TBB uses completely different approaches to parallel the execution of tasks, a direct performance comparison may not lead to reliable results.
@@
 
-For testing purposes we used primarily Linux setups, for one Ubuntu 17.04
-(linux 4.13) in a virtual machine environment, second Arch Linux (linux
+For testing purposes we used primarily \textsc{Linux} setups, for one \textsc{Ubuntu} 17.04
+(\textsc{Linux} 4.13) in a virtual machine environment, second \textsc{Arch Linux} (\textsc{Linux}
 4.17.3-1) running native, both updated on stable branch. Tests have also been
-conducted on Windows 10 and macOS High Sierra machines.
+conducted on \textsc{Windows 10} and \textsc{macOS High Sierra} machines.
 
-In our setups Sky- and Kabylake as well as Ryzen processors have been used with memory
+In our setups \textsc{Skylake} and \textsc{Kabylake} as well as \textsc{Ryzen} processors have been used with memory
 ranging from 8 to 24 GB of both DDR3-SODIMM and DDR4-SODIMM. 
@@
 configuration, but in this work we present only the results measured on
-Arch Linux with kernel 4.17.3-1 and 24 GB DDR4-SODIMM. We decided to use this
+\textsc{Arch} \textsc{Linux} with kernel 4.17.3-1 and 24 GB DDR4-SODIMM. We decided to use this
 system, since it is running a current kernel version and not in a virtual
@@
 \subsubsection{Benchmark Implementation}
-Since BOLSTER is deeply integrated into GeckoDB and the workload and usage of multi threading depends on the graph data in the system, we can not directly use the BOLSTER implementation for our evaluation.
+Since \textsc{Bolster} is deeply integrated into \textsc{GeckoDB} and the workload and usage of multi threading depends on the graph data in the system, we can not directly use the \textsc{Bolster} implementation for our evaluation.
 Thus, we extracted the respecting code parts into an own implementation and use this for a direct comparison against one that uses our thread pool.
 
-In Figure \ref{fig4}, we show the approximated version of the BOLSTER multi threading code that we use as baseline implementation.
+In Figure \ref{fig4}, we show the approximated version of the \textsc{Bolster} multi threading code that we use as baseline implementation.
 In our benchmark.
@@
 	\end{lstlisting}
-	\caption{Baseline multi threading implementation extracted from BOLSTER}
+	\caption{Baseline multi threading implementation extracted from \textsc{Bolster}}
 	\label{fig4}
@@
 \subsection{Summary}
-As last step of the evaluation, we summarize our results and compare them to the requirements of GeckoDB.
+As last step of the evaluation, we summarize our results and compare them to the requirements of \textsc{GeckoDB}.
 As mentioned before, our implementation reaches better performance than the baseline regarding our specific tests.
-Consequently, we solved the task for implementing and integrating the thread pool into BOLSTER to reach at least the baseline performance.
+Consequently, we solved the task for implementing and integrating the thread pool into \textsc{Bolster} to reach at least the baseline performance.
 Additionally, using our thread pool implementation solves the problem, that debugging and profiling with a huge number of threads is not possible.
 
-Although a small and constant waiting time was not a requirement for integration into GeckoDB, we ensure that assigning many tasks works fast for different thread pool sizes.
-Regarding the performance and also the design requirements of GeckoDB and BOLSTER, we fulfil these requirements and successfully implemented and evaluated our thread pool implementation.
+Although a small and constant waiting time was not a requirement for integration into \textsc{GeckoDB}, we ensure that assigning many tasks works fast for different thread pool sizes.
+Regarding the performance and also the design requirements of \textsc{GeckoDB} and \textsc{Bolster}, we fulfil these requirements and successfully implemented and evaluated our thread pool implementation.
 
@@
 
-To sum up, our thread pool implementation fulfils all performance and design requirements of GeckoDB and BOLSTER and reaches even faster execution time than the existing implementation.
+To sum up, our thread pool implementation fulfils all performance and design requirements of \textsc{GeckoDB} and \textsc{Bolster} and reaches even faster execution time than the existing implementation.
 In the future, we plan to make the following adaptations to our thread pool system.

### 2018-08-21 Presentations/Paper/paper_thread_pool.tex
Added pool_queue.pdf
Add pdf figures to paper tex file, and correct mail descriptions

@@
 \definecolor{backgroundColour}{rgb}{1,1,1}
+\definecolor{grayColor}{rgb}{0.56,0.56,0.56}
 
@@
 \author{
-	\IEEEauthorblockN{Robert Jendersie, Johannes Wuensche, Johann Wagner, \\Marten Wallewein-Eising, Marcus Pinnecke, Gunter Saake}
+	\IEEEauthorblockN{Robert Jendersie{\footnotesize\color{grayColor}$^\text{1}$}, Johannes Wuensche{\footnotesize\color{grayColor}$^\text{2}$}, Johann Wagner{\footnotesize\color{grayColor}$^\text{1}$}, \\Marten Wallewein-Eising{\footnotesize\color{grayColor}$^\text{2}$}, Marcus Pinnecke{\footnotesize\color{grayColor}$^\text{1}$}, and Gunter Saake{\footnotesize\color{grayColor}$^\text{1}$}}
 	\IEEEauthorblockA{\textit{Database and Software Engineering Group, Otto-von-Guericke University} \\
 		Magdeburg, Germany \\
-		\emph{firstname}.\emph{lastname}@ovgu.de} \and
+		\emph{firstname}.\emph{lastname}@{\color{grayColor}(\footnotesize$^\text{1}$}ovgu {\color{grayColor}\footnotesize$\mid$ $^\text{2}$}st.ovgu{\color{grayColor})}.de}
 }
@@
 Parallel data processing can be achieved by different approaches, like instruction and data parallelism or multi threading.
-We focus on multi threading by implementing a thread pool for the graph database system \textsc{GeckoDB}\footnote{\textsc{GeckoDB} source code repository: \url{https://github.com/\textsc{GeckoDB}}}.
+We focus on multi threading by implementing a thread pool for the graph database system \textsc{GeckoDB}\footnote{\textsc{GeckoDB} source code repository: \url{https://github.com/geckodb}}.
 The thread pool will be integrated into \textsc{Bolster}, a high performance library for parallel execution of primitives like for or filter on large data sets.
@@
 \begin{figure*}[htbp]
-	\centering \includegraphics[width=0.7\textwidth]{img/pool_structure.png}
+	\centering \includegraphics[width=0.7\textwidth]{img/pool_structure.pdf}
 	\caption{Design of the thread pool system}
@@
 \begin{figure}
-	\centering\includegraphics[width=0.4\textwidth]{img/pool_queue.png}
+	\centering\includegraphics[width=0.4\textwidth]{img/pool_queue.pdf}
 	\caption{Workflow of task enqueueing and execution in the thread pool}
@@
 \begin{figure}
-\centering	\includegraphics[width=0.25\textwidth]{img/waitingconcept.png}
+\centering	\includegraphics[width=0.25\textwidth]{img/waitingconcept.pdf}
 	\caption{Concept of waiting for tasks using a slot-map}
@@
 \begin{figure}
-	\includegraphics[width=0.5\textwidth]{img/pool_baseline.png}
+	\includegraphics[width=0.5\textwidth]{img/pool_baseline.pdf}
 	\caption{Time measures in nanoseconds of thread pool (blue) and baseline (orange) implementation with different numbers of threads.  }
@@
 \begin{figure}
-	\includegraphics[width=0.5\textwidth]{img/pool_avg.png}
+	\includegraphics[width=0.5\textwidth]{img/pool_avg.pdf}
 	\caption{Relation between waiting time (orange) and completion time (blue) of thread pools with different sizes processing 10000 tasks}
@@
 \begin{figure}
-	\centering\includegraphics[width=0.4\textwidth]{img/slot_distr.png}
+	\centering\includegraphics[width=0.4\textwidth]{img/slot_distr.pdf}
 	\caption{Utilization \% with different slot distribution strategies}
@@
 \begin{figure}
-	\centering\includegraphics[width=0.4\textwidth]{img/lock_free.png}
+	\centering\includegraphics[width=0.4\textwidth]{img/lock_free.pdf}
 	\caption{Comparison of Utilization \% of the lock-free queue versus mutex based heap}

### 2018-08-27 Presentations/Paper/paper_thread_pool.tex
Small fix

@@
 We focus on multi threading by implementing a thread pool for the graph database system \textsc{GeckoDB}\footnote{\textsc{GeckoDB} source code repository: \url{https://github.com/geckodb}}.
-The thread pool will be integrated into \textsc{Bolster}, a high performance library for parallel execution of primitives like for or filter on large data sets.
+The thread pool will be integrated into \textsc{Bolster}, a high performance library for parallel execution of primitives like \textit{for}-loops, or \textit{filter}-operation on large data sets.
 \textsc{Bolster} has similarities to Intel \textsc{Tbb}~\cite{books/daglib/0018624}, a \textsc{C++} template library for parallel programming, but \textsc{Bolster} is written from scratch to perfectly fit into \textsc{GeckoDB}s specialized storage engine and vectorized query engine, both written in C11~\cite{pinneckegridtables}.
